---
title: "Condition Pipeline"
author: "Hollie Keating, Sarah Conway & Maeve Samali"
date: "2026-03-10"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document contains the pipeline which can be applied to any of the conditions and nodes. Included:

 - Smoothing 
 - Generate FPCA scores
 - Plot top 5 and bottom 5 of FPCA scores 1, 2, and 3 
 
```{r, libraries, include = FALSE, warning=FALSE, message=FALSE}
library(data.table)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(plotly)
library(fda)
library(gridExtra)
library(fdapace)
library(tidyr)
library(table1)
library(factoextra)
library(cluster)
library(sjPlot)
library(refund)
library(glmnet)
library(pROC)  
library(DescTools) 
library(CalibrationCurves)
library(rsample)


#setwd("C:/Users/holli/OneDrive - National University of Ireland, Galway/Modules/Research Project")
setwd("C:/Users/35387/Documents/College/MSc Health Data Science/HDS Project/Rscript/Data")
#setwd("C:/Users/Sarah Conway/OneDrive/part 2/OneDrive/MSc-Project")
```

```{r, pipeline parameters}
c = 1 # condition 1 
t = 10 # 10 trials 
t1 = 1 # selecting one trial
node <- "fp1"
```

 
```{r, read in files}
# Loop which reads in separate excel file for each patient and compiles all data into a dataframe
# Merges time data and demogrpahic data which contains whether patients are controls or cases


#setwd("C:/Users/holli/OneDrive - National University of Ireland, Galway/Modules/Research Project")
setwd("C:/Users/35387/Documents/College/MSc Health Data Science/HDS Project/Rscript/Data")

times = fread("archive/time.csv")
demographic <- fread("archive/demographic.csv")
  
filenames = list.files("archive/patients/")

p_list <- list()

for(i in 1:length(filenames)){
  
  pathtomyfile = paste0("archive/patients/", filenames[i])
  temp <- fread(pathtomyfile, sep = ",", header = "auto")
  temp <- temp[ ,1:12]
  colnames(temp)<- c("subject", "trial", "condition", "sample", "fp1", "af7", "af3", "f1", "f3", "f5", "f7", "ft7")
  temp <- merge(temp, times, by ="sample")
  temp <- merge(temp, demographic, by = "subject")
  temp<- temp%>%
    filter (time_ms > -100)%>%   # time from -100ms to remove unnecessary data while including some pre beep curve info (100ms is where the beep occurs
    # ifelse(subject %in% missing, trial==15, trial==1) %>% # Only one trial out of 100 choosen. Trial 1 so patient hasn't learnt the test/timing/ keep up concealment
    filter(trial<=t)
  temp<- rename(temp, schizophrenia = group)

  p_list[[i]] <- temp
  
}

p_df <- do.call(rbind, p_list)

p_df<- p_df%>%
  mutate(
    subject = as.factor(subject),
    schizophrenia = as.factor(schizophrenia)
  )
```

 
```{r, smoothing splines 1}
## Smooth the data 
## Create a B-spline basis (mybasis) - building blocks to represent the curves
K = 100 # basis functions. More = more flexible
d = 3 # basis degree e.g. 3 = cubic

mybasis = create.bspline.basis(rangeval = c(0,1638), # x-axis range for curves 0-1638 timepoints
                               nbasis = K, 
                               norder = d+1) # cubic is an order 4 polynomial 


## Smoothing parameter object

## create an fd object - functional data object

penalty_order = 2 # 2nd derivative ie penalizes wigglier
penalty = 0.05 #amount of smoothing to be applied. Larger penalty = smoother.
fdParObj <- fdPar(fd(matrix(0, K, 1), mybasis), # fd object with k=100 coeff, 1 curve
                  Lfdobj = penalty_order, 
                  lambda = penalty)
```

 
```{r, smoothing splines 2}
# Smooth the data 
cond <- p_df %>%
  filter(condition == c) %>%
  filter(trial == t1)

#the subjects are not ordered correctly to fix this
cond$subject <- as.numeric(as.character(cond$subject))
cond <- cond %>% arrange(subject)

r_len <- nrow(cond %>% filter(subject == 1))
c_len <- length(unique(cond$subject))

matrix_sm_all <- matrix(cond[[node]], nrow = r_len, ncol = c_len, byrow= FALSE)

## smooth the data - cond
# smooth.basis(x, y, fdParObj)
sm_cond <- smooth.basis(seq(1, r_len, length.out = r_len), # x
                            matrix_sm_all, # y
                            fdParObj)# $fd     # extracts the smoothed functional data (fd) object ie
plot(sm_cond,
     main = "Smoothed Curves")

```




```{r, df}
# Plots 
# dfresponse - raw eeg, smoothed eeg, merged demographic 
 
# sm_cond$y is raw eeg signals
# smoothed is predict(sm_cond) 
# fd is basis and coeff

dfresponse <- data.frame(ID = as.numeric(as.character(cond$subject)),
                         time = cond$time_ms,
                         eeg_raw = as.numeric(sm_cond$y),
                         eeg_smooth = as.numeric(predict(sm_cond))) %>%   # smoothed predictions
                          arrange(ID)


demographic <- demographic %>%
  rename(ID = subject) %>%
  mutate(group = as.character(group),
         gender = stringr::str_trim(as.character(gender)), # to handle space
    
    group = recode(group, `0` = "Control", `1` = "Case"),
    gender = recode(gender, "M" = "Male", "F" = "Female"),
    group = factor(group, levels = c("Control", "Case")),
    gender = factor(gender, levels = c("Male", "Female")))

# table(demographic$group) # 25 controls, 11 cases
# table(demographic$gender) # 30 males. 6 females

myIDS = unique((dfresponse$ID))
#test = as.numeric(predict(sm_cond)) # smoothed predictions
#dfresponse$eeg_smooth = test

dfresponse_merged <- dfresponse %>% left_join(demographic, by = "ID")


```

```{r, demographic plots, fig.width=9, fig.asp=1.5}

# plot color = cases vs controls 
# group = ID: draw one seperate line for each patient ID
group_plot <- dfresponse_merged %>%
  ggplot(aes(x = time, y = eeg_smooth, group = ID, color = group)) +
  geom_line(alpha = 0.6) +
  scale_color_manual(values = c("Control" = "orange", "Case" = "darkgreen")) +
  theme_bw()+
  labs(title = "Smoothed Curves Case vs Controls",
       x = "Time",
       y = "EEG Signals",
       color = "")

#group_plot


gender_plot <- dfresponse_merged %>%
  ggplot(aes(x = time, y = eeg_smooth, group = ID, color = group)) +
  geom_line(alpha = 0.6) +
  scale_color_manual(values = c("Control" = "orange", "Case" = "darkgreen")) +
  theme_bw()+
  labs(title = "Smoothed Curves Case vs Controls",
       x = "Time",
       y = "EEG Signals",
       color = "")+facet_wrap(~gender)

#gender_plot


# plot by age > 40 or < 40     its not the age they werre diagnosed at its just the age they are now.................
age_plot <- dfresponse_merged %>% 
  mutate(age_grouped = ifelse(age < 40, "<40", ">=40")) %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = ID, color = group)) +
  geom_line(alpha = 0.6) +
  scale_color_manual(values = c("Control" = "orange", "Case" = "darkgreen")) +
  theme_bw()+
  labs(title = "Smoothed Curves by Age",
       x = "Time",
       y = "EEG Signals",
       color = "")+facet_wrap(~age_grouped)

#age_plot

# years in education 
# primary plus secondary is ~14 years 
edu_plot <- dfresponse_merged %>% 
  mutate(education_grouped = ifelse(education < 15, "<15", ">=15")) %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = ID, color = group)) +
  geom_line(alpha = 0.6) +
  scale_color_manual(values = c("Control" = "orange", "Case" = "darkgreen")) +
  theme_bw()+
  labs(title = "Smoothed Curves by Years in Education",
       x = "Time",
       y = "EEG Signals",
       color = "")+facet_wrap(~education_grouped)

#edu_plot

grid.arrange(group_plot, gender_plot, age_plot, edu_plot, ncol = 1)

# table_data <- dfresponse_merged %>%
#   distinct(ID, group, gender)      # one per patient
# 
# table1(~ gender | group,
#        data = table_data)


# demographic only has up to patient 77


dfresponse %>%
  filter(ID==1) %>%
  ggplot(aes(x=time, y = eeg_raw)) + geom_line() + geom_line(lwd = 1.5, aes(y=eeg_smooth, color = "smooth"))


```

 
 
## FPCA Scores
 
```{r, fpca scores}
## FPCA through conditional expectation (PACE) - irregular data allowed
# FPCA via PACE 
# applied to sm_cond: condition c smoothed data 

# make a list containing dataframes with rows grouped by ID
list_format <- split(dfresponse, dfresponse$ID)

# create lists of time and response (eeg)
list_t <- lapply(list_format,
                 dplyr::select, time)
list_eeg <- lapply(list_format,
                   dplyr::select, eeg_raw)

# unlist
list_t <- lapply(list_t,
                 unlist, use.names = F)
list_eeg <- lapply(list_eeg,
                   unlist, use.names = F)

# apply PACE (with fdapace)
resMEAN <- FPCA(Ly = list_eeg, Lt = list_t,
                optns = list(dataType = 'Sparse')) #Dense or p >> n (time pts > n patients)

##is the data dense?
plot(FPCA(Ly = list_eeg, Lt = list_t,
                optns = list(dataType = 'Dense')))
# FPC results
plot(resMEAN)
#resMEAN$xiEst #FPCA scores
```
 
 
## Highest & Lowest Scores 
 
```{r, fpc df}
# FPCA scores

# FPC result
# resMEAN$xiEst #FPCA scores

fpca_scores <- as.data.frame(resMEAN$xiEst) %>% 
  mutate(ID =myIDS, .before = 1)             # add an ID row, first column

colnames(fpca_scores) <- c("ID", "fpca_1", "fpca_2", "fpca_3")

dfresponse <- dfresponse %>% mutate(ID = as.numeric(ID))

# Plots - merge FCP scores with df response 
fpca_scores_merged <- dfresponse_merged %>% left_join(fpca_scores, by = "ID")
```

```{r, fpc plots function }
# Extract top 5 fpc scores
# Top 5: 1638*5= 8190

fpca_plots <- function(df = fpca_scores_merged, fpc){
  
  fpca_name <- as.character(substitute(fpc))
  
  top_5_fpc <- fpca_scores_merged %>%
  arrange(desc({{fpc}})) %>%
  slice(1:8190) %>%
  pull(ID)

  # Extract bottom 5 fpca 1 scores
  bottom_5_fpc <- fpca_scores_merged %>%
    arrange({{fpc}}) %>%              # default ascending
    slice(1:8190) %>%
    pull(ID)
  
  # df of top & bottom 
  group_ids_fpc <- bind_rows(
    data.frame(ID = unique(top_5_fpc), top_bot = "Top 5"),
    data.frame(ID = unique(bottom_5_fpc), top_bot = "Bottom 5")
  )
  
  plot_data_fpc <- fpca_scores_merged %>%
    semi_join(group_ids_fpc, by = "ID") %>%      # keeps rows that IDs are in both dfs
    left_join(group_ids_fpc, by = "ID")          # adds the group label
  
  plot_data_fpc %>% ggplot(aes(x = time, y = eeg_raw, group = ID, colour = top_bot)) +
    geom_line(linewidth = 0.5) +
    theme_bw() +
    labs(title = paste("Top vs Bottom 5 Curves Scores:", fpca_name),
         colour = "Group")
}
```
 
```{r, fpca plots}
fpca_plots(fpc = fpca_1)
fpca_plots(fpc = fpca_2)
fpca_plots(fpc = fpca_3)
```
 
 
```{r, kmeans, include = F}
##KMEANS CLUSTERING
set.seed(1)

#scores <- na.omit(scores)
rownames(fpca_scores)

fpca_scores_demo <- merge(fpca_scores, demographic, by = "ID")

target_y <- fpca_scores_demo$group
features_x <- fpca_scores_demo[c(2,3,4)]

kmeans_c1 <- kmeans(features_x, centers = 4, nstart = 10)

#CONFUSION MATRIX
table(fpca_scores_demo$group, kmeans_c1$cluster)

#mean(scores$group == kmeans_c1$cluster)
fviz_cluster(kmeans_c1, data = features_x, 
             main = "K-Means Clusters")
```

```{r, plot kmeans, include=FALSE}
##PLOTTING THE CLUSTERS
km <- data.frame(ID = fpca_scores_demo$ID, 
                 cluster = kmeans_c1$cluster)

#smoothed_all <- merge(smooth_long, cond1,  by = "subject")

df_cluster <- merge(km, fpca_scores_merged, by= "ID") #%>% 
  #arrange(subject, time_ms)

#plotting the cluster by schizophrenia
df_cluster %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = ID, colour = group)) +
  geom_line(lwd = 0.5)+
  facet_wrap(~ cluster) +
  labs(title = "EEG signmals by cluster, coloured by schizophrenia")
```


```{r, glms}
subject_df <- fpca_scores_merged %>%
  distinct(ID, group, gender, age, education, fpca_1, fpca_2, fpca_3)

# subject_df$group <- factor(subject_df$group)
# subject_df$gender <- factor(subject_df$gender)


#=================
# Gender Model 
# ================

log_gender <- glm(group ~ age + education + fpca_1 + fpca_2 + fpca_3,
             data = subject_df,
             family = binomial)

summary(log_gender)


## converting coefficients into OR

exp(coef(log_gender))


# from summary, FPCA1 shows most significance at p = 0.06.., genderM significance = 0.2131
# gender coef = 0.1485062


# ====================================
# Education Model
# ====================================

model_edu <- glm(group ~ education,
                 data = subject_df,
                 family = binomial)

summary(model_edu)
exp(coef(model_edu))

## for each additional year of education, the odds of being in the schizophrenia group decrease by about 57%

## creating a boxplot 

ggplot(subject_df, aes(x = factor(group), y = education, color = factor(group))) +
  geom_boxplot() +
  labs(x = "Group", y = "Education")

# =====================
# FPCA Model 
# ====================

model_eeg <- glm(group ~ fpca_1 + fpca_2 + fpca_3,
                 data = subject_df,
                 family = binomial)

summary(model_eeg)

exp(coef(model_eeg))

# 1st PC most significant, OR = 0.99.. so only marginally reduces odds of being in schizophrenia group with increase of PC1

ggplot(subject_df, aes(x = factor(group), y = fpca_1, color = factor(group))) +
  geom_boxplot() +
  labs(x = "Group", y = "FPCA 1")

```

### Education Model

Education was significantly associated with schizophrenia diagnosis, with higher education linked to lower log-odds (OR = 0.57, p = 0.018), indicating a decrease in odds of approximately 43% per additional year of education.

By analysing the grouped bar chart for education, individuals with schizophrenia tended to have lower levels of education compared to controls - however this association may reflect the impact of earlier illness onset or other cofounding factors.


### Gender Model

Gender was not significant within association with schizophrenia presence (OR = 0.15, p = 0.213). 
The distribution of gender was similar across the control and schizophrenia groups, with a higher proportion of males in both groups which supports our above result. However, the sample was predominantly male which may limit our ability to detect gender differences. 

### FPCA Model

FPCA 1 was significantly associated with schizophrenia diagnosis (p = 0.039), although the the odds ratio was quite small (OR \approx 1.00). FPCA 2 showed no association, and FPCA 3 showed a weak non-significant trend.  
 
# MFPCA Model 

So now we want to take into account that we have more than 1 function per person. People completed each sensory condition up to 100 times. Including more of these trials will improve our ability to predict whether an EEG is typical of schizophrenia or not.
 
## Smoothing all 10 trials

As before we need to smooth the data, but this time for all the trials.

```{r, mfpca smooth}
##smoothed data for 10 trials:
cond_trials <- p_df %>% 
  filter(condition == c) %>% 
  arrange(trial)

#the subjects are not ordered correctly to fix this
cond_trials$subject <- as.numeric(as.character(cond_trials$subject))
cond_trials <- cond_trials %>% arrange(subject)

##number of time points for each curve
r_lent <- nrow(cond_trials %>% filter(subject == 1) %>% filter(trial == 1))
#number of individual curves
c_lent <- nrow(cond_trials %>% distinct(subject, trial))

##all trials for all people for fp1
matrix_sm_trial <- matrix(cond_trials[[node]], nrow = r_lent, ncol = c_lent, byrow= FALSE)

## smooth the data - cond
# smooth.basis(x, y, fdParObj)
sm_trials <- smooth.basis(seq(1, r_lent, length.out = r_lent), # x
                            matrix_sm_trial, # y
                            fdParObj)# $fd     # extracts the smoothed functional data (fd) object ie
plot(sm_trials,
     main = "Smoothed Curves")
```

### Data in long format & plots

We create a dataframe of all the relevant measurements and we have plotted the smoothed data for each trial for each person below.

```{r, mfpca plots}
##create a response df in long format
dfresponse10 <- data.frame(ID = as.numeric(as.character(cond_trials$subject)),
                         time = cond_trials$time_ms,
                         trial = cond_trials$trial,
                         eeg_raw = as.numeric(sm_trials$y),
                         eeg_smooth = as.numeric(predict(sm_trials)),
                         schizophrenia = cond_trials$schizophrenia) %>%   # smoothed predictions
                         arrange(ID)

##Plot all trials for each subject 
dfresponse10  %>% filter(ID <= 10) %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = factor(trial), color = factor(schizophrenia))) +
  geom_line() +
  facet_wrap(~ ID, nrow = 2) +
  ggtitle("Multilevel functional observations")

dfresponse10  %>% filter(ID > 10 & ID <=20) %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = factor(trial), color = factor(schizophrenia))) +
  geom_line() +
  facet_wrap(~ ID, nrow = 2) +
  ggtitle("Multilevel functional observations")

dfresponse10  %>% filter(ID > 20 & ID <= 71) %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = factor(trial), color = factor(schizophrenia))) +
  geom_line() +
  facet_wrap(~ ID, nrow = 2) +
  ggtitle("Multilevel functional observations")

dfresponse10  %>% filter(ID > 71) %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = factor(trial), color = factor(schizophrenia))) +
  geom_line() +
  facet_wrap(~ ID, nrow = 2) +
  ggtitle("Multilevel functional observations")

```


## mFPCA

Now we perform multi-level modelling, where the first level is the individual subject level and the second level is the individual trial level. So we get a score for each person and for each curve (trial).

```{r, mfpca}
# df: Data in long format with columns, subject Id = Id, stride_id (wave_id) = visit, arg = time (seq(0, 1, length = 101), obs = y values
#Output: each row is a curve with subject Id and wave id
df_wide <- pivot_wider(dfresponse10[,-4], names_from = time, values_from = eeg_smooth)
Y <- as.matrix(df_wide[, c(4:1641)]) # excluding column id, trial, schizophrenia

#mfpca
mlfpc <- mfpca.face(Y = Y, # matrix- each row is a curves
                    id = df_wide$ID, # subject id: c(1,1,2,2,2,3,3 ...)
                    visit = df_wide$trial, # trial id c(1,2,3,.., 1, 2,3, 4, ...)
                    twoway = TRUE,
                    #pve2 = 0.75, # pve for level 2
                    pve = 0.95
                    )

```

## Multlevel FPCs

-   Level 1

```{r, mfpca level 1}
t <- unique(dfresponse10$time)
egf1 <- data.frame(mlfpc$efunctions$level1)
ggplot(egf1) +
  geom_line(lwd =0.55, aes(x = t, y = egf1[, 1], color = " 1")) +
  geom_line(lwd =0.55,aes(x = t, y = egf1[, 2], color = " 2")) +
  geom_line(lwd =0.55,aes(x = t, y = egf1[, 3], color = " 3")) +
  geom_line(lwd =0.55,aes(x = t, y = egf1[, 4], color = " 4")) +
  scale_color_manual(values = c(" 1" = "blue", " 2" = "green", " 3" = "red", " 4" = "purple")) +
  labs(x = "Time", y = "value", color = "mFPC") +
  theme_bw()

```


-   Level 2

```{r, mfpca level 2}
t <- unique(dfresponse10$time)
egf2 <- data.frame(mlfpc$efunctions$level2)
ggplot(egf2) +
  geom_line(aes(x = t, y = egf2[, 1], color = " 1")) +
  geom_line(aes(x = t, y = egf2[, 2], color = " 2")) +
  geom_line(aes(x = t, y = egf2[, 3], color = " 3")) +
  geom_line(aes(x = t, y = egf2[, 4], color = " 4")) +
  scale_color_manual(values = c(" 1" = "blue", " 2" = "green", " 3" = "red", " 4" = "purple")) +
  labs(x = "Time (proportion)", y = "value", color = "mFPC") +
  theme_bw()

```


## Nature of the variability

Level 1

```{r, eigenfns 1}
# add and subtract eigenfunction to the overall mean
egf_l1 <- egf1 %>%  mutate(mu = mlfpc$mu,
                     p1 = mlfpc$mu  + X1*sqrt(mlfpc$evalues$level1[1]),
                     n1 = mlfpc$mu - X1*sqrt(mlfpc$evalues$level1[1]),
                     
                     p2 = mlfpc$mu  + X2*sqrt(mlfpc$evalues$level1[2]),
                     n2 = mlfpc$mu - X2*sqrt(mlfpc$evalues$level1[2]),
                     
                     p3 = mlfpc$mu  + X3*sqrt(mlfpc$evalues$level1[3]),
                     n3 = mlfpc$mu - X3*sqrt(mlfpc$evalues$level1[3]),
                     
                     p4 = mlfpc$mu  + X4*sqrt(mlfpc$evalues$level1[4]),
                     n4 = mlfpc$mu - X4*sqrt(mlfpc$evalues$level1[4]))

```



# First and second modes
```{r, level 1 plot}
# first mode
ggplot(egf_l1, aes(x = t, y = mu)) + geom_line(color = "red", lwd = 0.8) +
geom_point(aes(x = t, y = p1), shape = "+", size = 3) + 
geom_point(aes(x = t, y = n1), shape = "-", size = 3)  +
  labs(y ="Y", x = "Time (proportion)") +ggtitle("PC 1 (..%)") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),plot.title.position = "panel"
  )

# Second mode
ggplot(egf_l1, aes(x = t, y = mu)) + geom_line(color = "red", lwd = 0.8) +
geom_point(aes(x = t, y = p2), shape = "+", size = 3) + 
geom_point(aes(x = t, y = n2), shape = "-", size = 3)  +
  labs(y ="Y", x = "Time (proportion)") + ggtitle("PC 2 (...%)") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),plot.title.position = "panel"
  )

# third mode
ggplot(egf_l1, aes(x = t, y = mu)) + geom_line(color = "red", lwd = 0.8) +
geom_point(aes(x = t, y = p3), shape = "+", size = 3) + 
geom_point(aes(x = t, y = n3), shape = "-", size = 3)  +
  labs(y ="Y", x = "Time (proportion)") + ggtitle("PC 3 (...%)") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),plot.title.position = "panel"
  )
```


 level 2

First and second modes

```{r, fpc level 2}
egf2 <- egf2 %>%  mutate(mu = mlfpc$mu,
                       p1 = mlfpc$mu  + X1*sqrt(mlfpc$evalues$level2[1]),
                       n1 = mlfpc$mu - X1*sqrt(mlfpc$evalues$level2[1]),
                       
                       p2 = mlfpc$mu  + X2*sqrt(mlfpc$evalues$level2[2]),
                       n2 = mlfpc$mu - X2*sqrt(mlfpc$evalues$level2[2]),
                       
                       p3 = mlfpc$mu  + X3*sqrt(mlfpc$evalues$level2[3]),
                       n3 = mlfpc$mu - X3*sqrt(mlfpc$evalues$level2[3]),
                       
                       p4 = mlfpc$mu  + X4*sqrt(mlfpc$evalues$level2[4]),
                       n4 = mlfpc$mu - X4*sqrt(mlfpc$evalues$level2[4]))
```


```{r, fpc level 2 plots}
# First mode
ggplot(egf2, aes(x = t, y = mu)) + geom_line(color = "red", lwd = 0.8) +
geom_point(aes(x = t, y = p1), shape = "+", size = 3) + 
geom_point(aes(x = t, y = n1), shape = "-", size = 3)  +
  labs(y ="Y", x = "Time (proportion)") +
   ggtitle("PC 1 (....%)") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),plot.title.position = "panel"
  )

# Second mode
ggplot(egf2, aes(x = t, y = mu)) + geom_line(color = "red", lwd = 0.8) +
geom_point(aes(x = t, y = p2), shape = "+", size = 3) + 
geom_point(aes(x = t, y = n2), shape = "-", size = 3)  +
  labs(y ="Y", x = "Time (proportion)") +
   ggtitle("PC 2 (....%)") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),plot.title.position = "panel"
  )

# Third mode
ggplot(egf2, aes(x = t, y = mu)) + geom_line(color = "red", lwd = 0.8) +
geom_point(aes(x = t, y = p3), shape = "+", size = 3) + 
geom_point(aes(x = t, y = n3), shape = "-", size = 3)  +
  labs(y ="Y", x = "Time (proportion)") +
   ggtitle("PC 3 (....%)") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),plot.title.position = "panel"
  )
```


### Multilevel FPC scores

#### Level 1

Patient Level scores - 1 per person

```{r, mfpc1}
level1_scores <- cbind(data.frame(mlfpc$scores$level1), subj_id = as.factor(unique(df_wide$ID)), schizophrenia = df_wide$schizophrenia[match(unique(df_wide$ID), df_wide$ID)])# subject level scores

ggplot(level1_scores,aes(x = X1, y = X2, color = schizophrenia)) + geom_point() +
  labs(x = "mFPC1", y = "mFPC2", title = "..") + theme_bw()
```

#### Level 2

Individual Trial scores (i.e. 1 score per function/trial) -multiple per person

```{r, mfpc2}
level2_scores <- cbind(data.frame(mlfpc$scores$level2), subj_id = factor(df_wide$ID), trial_id = df_wide$trial, schizophrenia = df_wide$schizophrenia)

ggplot(level2_scores, aes(x = trial_id, y = X1, group = subj_id, color = schizophrenia)) +
  geom_line() + geom_point() + theme_bw() +
  labs(y = "mFPC1", x = "trials order - longitudinal", title = "") 
 
ggplot(level2_scores, aes(x = subj_id, y = X1, group = subj_id, colour = schizophrenia)) + geom_boxplot() + theme_bw() +
  labs(y = "mFPC1", x = "Subject Id", title = "Score distribution") 
 

```




# edited all trials

- Creating sample of IDs, one row per subject
- Using split / test / train
- Sensors and conditions separate, all trials used
- Level 1 and level 2 scores as predictors


condition 1, fp1

```{r cond1fp1}
# Set up dataset: 9 scores per person, gender, 
# 3 conditions x 3 fpca scores for each condition

# Include nodes fp1, fp2 (prefrontal), f3, f4, fz (frontal)

# all three conditions. one at a time and merge
nodes_en <- c("fp1") # pre-frontal and frontal nodes


# Smooth the data 
# en: elastic net 
cond_en <- p_df %>%
  filter(condition == 1) %>%
  dplyr::select(-sample, -condition, -af7, -af3, , -f3, -f1, -f5, -f7, -ft7)

cond_en <- as.data.frame(cond_en)

#the subjects are not ordered correctly to fix this
cond_en$subject <- as.numeric(as.character(cond_en$subject))
cond_en <- cond_en %>% arrange(subject)

# dimensions

# changing r_len as subject 1 now has all trials together
# rows = number of time points per curve 
# columns, number of curves = subject x trial combinations

r_len <- length(unique(cond_en$time_ms))
c_len <- n_distinct(cond_en$subject, cond_en$trial)

# make matrix
matrix_sm_all_en <- matrix(as.matrix(cond_en[, nodes_en]), 
                        nrow = r_len,
                        ncol = c_len * length(nodes_en),
                        byrow = FALSE)

## smooth the data - cond
sm_cond_en <- smooth.basis(seq(1, r_len, length.out = r_len), # x
                            matrix_sm_all_en, # y
                            fdParObj)# $fd     # extracts the smoothed functional data (fd) object ie

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                         time = cond_en$time_ms,
                         eeg_raw = as.numeric(sm_cond_en$y), #raw eeg signals
                         eeg_smooth = as.numeric(predict(sm_cond_en))) %>%   # smoothed predictions
                          arrange(ID, trial)

# table(demographic$group) # 25 controls, 11 cases
# table(demographic$gender) # 30 males. 6 females

myIDS = unique((dfresponse_en$ID))
#test = as.numeric(predict(sm_cond)) # smoothed predictions
#dfresponse$eeg_smooth = test

dfresponse_merged_en <- dfresponse_en %>% left_join(demographic, by = "ID")


# mFPCA scores condition 1 fp1


# Now we perform multi-level modelling, where the first level is the individual subject level and the second level is the individual trial level. So we get a score for each person and for each curve (trial).

# df: Data in long format with columns, subject Id = Id, stride_id (wave_id) = visit, arg = time (seq(0, 1, length = 101), obs = y values
#Output: each row is a curve with subject Id and wave id

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)

Y_en <- as.matrix(df_wide_en[, -(1:2)]) # excluding column id, trial

#mfpca

mlfpc_en <- mfpca.face(Y = Y_en, # matrix- each row is a curves
                    id = df_wide_en$ID, # subject id: c(1,1,2,2,2,3,3 ...)
                    visit = df_wide_en$trial, # trial id c(1,2,3,.., 1, 2,3, 4, ...)
                    twoway = TRUE,
                    pve = 0.95)

# extracting level 1 and level 2 scores

## level 1 (subject level)

# mlfpc_en$score$level1 = mfpca level 1 scores


level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 ) # add an ID row, first column

level1_en <- level1_en[, 1:5]

colnames(level1_en) <- c("ID", "cond1_fp1_L1_1", "cond1_fp1_L1_2", "cond1_fp1_L1_3", "cond1_fp1_L1_4")


## level 2 (trial level)

level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 ) # add an ID row, first column

level2_en <- level2_en[, 1:3]

colnames(level2_en) <- c("ID", "trial", "cond1_fp1_L2_1") # only one score for level 2?


dfresponse_en <- dfresponse_en %>% mutate(ID = as.numeric(ID))


# Plots - merge level scores 


# merge level 1 and level 2 scores

fpca_fp1_en1 <- level2_en %>%
  left_join(level1_en, by = "ID")


fpca_fp1_merged_en1 <- demographic %>%
  left_join(fpca_fp1_en1, by = "ID")

```


condition 2, fp1
```{r cond2fp1}

cond_en <- p_df %>%
  filter(condition == 2) %>%
  dplyr::select(-sample, -condition, -af7, -af3, -f3, -f1, -f5, -f7, -ft7)

cond_en <- as.data.frame(cond_en)

cond_en$subject <- as.numeric(as.character(cond_en$subject))
cond_en <- cond_en %>% arrange(subject)

r_len <- length(unique(cond_en$time_ms))

c_len <- n_distinct(cond_en$subject, cond_en$trial)

matrix_sm_all_en <- matrix(as.matrix(cond_en[, nodes_en]), 
                        nrow = r_len,
                        ncol = c_len * length(nodes_en),
                        byrow = FALSE)

sm_cond_en <- smooth.basis(seq(1, r_len, length.out = r_len),
                            matrix_sm_all_en,
                            fdParObj)

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                         time = cond_en$time_ms,
                         eeg_raw = as.numeric(sm_cond_en$y),
                         eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                          arrange(ID, trial)

myIDS = unique((dfresponse_en$ID))

dfresponse_merged_en <- dfresponse_en %>% left_join(demographic, by = "ID")

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)

Y_en <- as.matrix(df_wide_en[, -(1:2)])

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )

myIDS = unique((dfresponse_en$ID))

level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )

level1_en <- level1_en[, 1:5]

colnames(level1_en) <- c("ID", "cond2_fp1_L1_1", "cond2_fp1_L1_2", "cond2_fp1_L1_3", "cond2_fp1_L1_4")

level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

level2_en <- level2_en[, 1:3]

colnames(level2_en) <- c("ID", "trial", "cond2_fp1_L2_1")

dfresponse_en <- dfresponse_en %>% mutate(ID = as.numeric(ID))

fpca_fp1_en2 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_fp1_merged_en2 <- fpca_fp1_merged_en1 %>%
  left_join(fpca_fp1_en2, by = c("ID", "trial")) # joining here by ID and trial as we are given "trial.x" and "trial.y"



```


condition 3, fp1

```{r cond3fp1}

nodes_en <- c("fp1")

cond_en <- p_df %>%
  filter(condition == 3) %>%
  dplyr::select(-sample, -condition, -af7, -af3, -f3, -f1, -f5, -f7, -ft7)

cond_en <- as.data.frame(cond_en)

cond_en$subject <- as.numeric(as.character(cond_en$subject))
cond_en <- cond_en %>% arrange(subject)

r_len <- length(unique(cond_en$time_ms))

c_len <- n_distinct(cond_en$subject, cond_en$trial)

matrix_sm_all_en <- matrix(as.matrix(cond_en[, nodes_en]), 
                        nrow = r_len,
                        ncol = c_len * length(nodes_en),
                        byrow = FALSE)

sm_cond_en <- smooth.basis(seq(1, r_len, length.out = r_len),
                            matrix_sm_all_en,
                            fdParObj)

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                         time = cond_en$time_ms,
                         eeg_raw = as.numeric(sm_cond_en$y),
                         eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                          arrange(ID, trial)

myIDS = unique((dfresponse_en$ID))

dfresponse_merged_en <- dfresponse_en %>% left_join(demographic, by = "ID")

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)

Y_en <- as.matrix(df_wide_en[, -(1:2)])

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )

myIDS = unique((dfresponse_en$ID))

level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )

level1_en <- level1_en[, 1:5]

colnames(level1_en) <- c("ID", "cond3_fp1_L1_1", "cond3_fp1_L1_2", "cond3_fp1_L1_3", "cond3_fp1_L1_4")

level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

level2_en <- level2_en[, 1:3]

colnames(level2_en) <- c("ID", "trial", "cond3_fp1_L2_1")

dfresponse_en <- dfresponse_en %>% mutate(ID = as.numeric(ID))

fpca_fp1_en3 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_fp1_merged_en3 <- fpca_fp1_merged_en2 %>%
  left_join(fpca_fp1_en3, by = c("ID", "trial"))


```


-- f3 --

condition 1, f3

```{r cond1f3}

nodes_en <- c("f3")

cond_en <- p_df %>%
  filter(condition == 1) %>%
  dplyr::select(-sample, -condition, -af7, -af3, -f1, -f5, -f7, -ft7, -fp1)

cond_en <- as.data.frame(cond_en)

cond_en$subject <- as.numeric(as.character(cond_en$subject))
cond_en <- cond_en %>% arrange(subject)

r_len <- length(unique(cond_en$time_ms))

c_len <- n_distinct(cond_en$subject, cond_en$trial)

matrix_sm_all_en <- matrix(as.matrix(cond_en[, nodes_en]), 
                        nrow = r_len,
                        ncol = c_len * length(nodes_en),
                        byrow = FALSE)

sm_cond_en <- smooth.basis(seq(1, r_len, length.out = r_len),
                            matrix_sm_all_en,
                            fdParObj)

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                         time = cond_en$time_ms,
                         eeg_raw = as.numeric(sm_cond_en$y),
                         eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                          arrange(ID, trial)

myIDS = unique((dfresponse_en$ID))

dfresponse_merged_en <- dfresponse_en %>% left_join(demographic, by = "ID")

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)

Y_en <- as.matrix(df_wide_en[, -(1:2)])

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )

myIDS = unique((dfresponse_en$ID))

level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )

level1_en <- level1_en[, 1:5]

colnames(level1_en) <- c("ID", "cond1_f3_L1_1", "cond1_f3_L1_2", "cond1_f3_L1_3", "cond1_f3_L1_4")

level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

level2_en <- level2_en[, 1:3]

colnames(level2_en) <- c("ID", "trial", "cond1_f3_L2_1")

dfresponse_en <- dfresponse_en %>% mutate(ID = as.numeric(ID))


fpca_f3_en1 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_f3_merged_en1 <- fpca_fp1_merged_en3 %>%
  left_join(fpca_f3_en1, by = c("ID", "trial"))

```


condition 2, f3

```{r cond2f3}

nodes_en <- c("f3")

cond_en <- p_df %>%
  filter(condition == 2) %>%
  dplyr::select(-sample, -condition, -af7, -af3, -f1, -f5, -f7, -ft7, -fp1)

cond_en <- as.data.frame(cond_en)

cond_en$subject <- as.numeric(as.character(cond_en$subject))
cond_en <- cond_en %>% arrange(subject)

r_len <- length(unique(cond_en$time_ms))

c_len <- n_distinct(cond_en$subject, cond_en$trial)

matrix_sm_all_en <- matrix(as.matrix(cond_en[, nodes_en]), 
                        nrow = r_len,
                        ncol = c_len * length(nodes_en),
                        byrow = FALSE)

sm_cond_en <- smooth.basis(seq(1, r_len, length.out = r_len),
                            matrix_sm_all_en,
                            fdParObj)

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                         time = cond_en$time_ms,
                         eeg_raw = as.numeric(sm_cond_en$y),
                         eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                          arrange(ID, trial)

myIDS = unique((dfresponse_en$ID))

dfresponse_merged_en <- dfresponse_en %>% left_join(demographic, by = "ID")

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)

Y_en <- as.matrix(df_wide_en[, -(1:2)])

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )

myIDS = unique((dfresponse_en$ID))

level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )

level1_en <- level1_en[, 1:5]

colnames(level1_en) <- c("ID", "cond2_f3_L1_1", "cond2_f3_L1_2", "cond2_f3_L1_3", "cond2_f3_L1_4")

level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

level2_en <- level2_en[, 1:3]

colnames(level2_en) <- c("ID", "trial", "cond2_f3_L2_1")

dfresponse_en <- dfresponse_en %>% mutate(ID = as.numeric(ID))


fpca_f3_en2 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_f3_merged_en2 <- fpca_f3_merged_en1 %>%
  left_join(fpca_f3_en2, by = c("ID", "trial"))


```


condition3, f3

```{r cond3f3}
nodes_en <- c("f3")

cond_en <- p_df %>%
  filter(condition == 1) %>%
  dplyr::select(-sample, -condition, -af7, -af3, -f1, -f5, -f7, -ft7, -fp1)

cond_en <- as.data.frame(cond_en)

cond_en$subject <- as.numeric(as.character(cond_en$subject))
cond_en <- cond_en %>% arrange(subject)

r_len <- length(unique(cond_en$time_ms))

c_len <- n_distinct(cond_en$subject, cond_en$trial)

matrix_sm_all_en <- matrix(as.matrix(cond_en[, nodes_en]), 
                        nrow = r_len,
                        ncol = c_len * length(nodes_en),
                        byrow = FALSE)

sm_cond_en <- smooth.basis(seq(1, r_len, length.out = r_len),
                            matrix_sm_all_en,
                            fdParObj)

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                         time = cond_en$time_ms,
                         eeg_raw = as.numeric(sm_cond_en$y),
                         eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                          arrange(ID, trial)

myIDS = unique((dfresponse_en$ID))

dfresponse_merged_en <- dfresponse_en %>% left_join(demographic, by = "ID")

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)

Y_en <- as.matrix(df_wide_en[, -(1:2)])

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )

myIDS = unique((dfresponse_en$ID))

level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )

level1_en <- level1_en[, 1:5]

colnames(level1_en) <- c("ID", "cond3_f3_L1_1", "cond3_f3_L1_2", "cond3_f3_L1_3", "cond3_f3_L1_4")

level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

level2_en <- level2_en[, 1:3]

colnames(level2_en) <- c("ID", "trial", "cond3_f3_L2_1")

dfresponse_en <- dfresponse_en %>% mutate(ID = as.numeric(ID))


fpca_f3_en3 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_f3_merged_en3 <- fpca_f3_merged_en2 %>%
  left_join(fpca_f3_en3, by = c("ID", "trial"))

```


-- f7 --


condition 1, f7

```{r cond1f7}


nodes_en <- c("f7")

cond_en <- p_df %>%
  filter(condition == 1) %>%
  dplyr::select(-sample, -condition, -af7, -af3, -f3, -fp1, -f1, -f5, -ft7)

cond_en <- as.data.frame(cond_en)

cond_en$subject <- as.numeric(as.character(cond_en$subject))
cond_en <- cond_en %>% arrange(subject)

r_len <- length(unique(cond_en$time_ms))

c_len <- n_distinct(cond_en$subject, cond_en$trial)

matrix_sm_all_en <- matrix(as.matrix(cond_en[, nodes_en]), 
                        nrow = r_len,
                        ncol = c_len * length(nodes_en),
                        byrow = FALSE)

sm_cond_en <- smooth.basis(seq(1, r_len, length.out = r_len),
                            matrix_sm_all_en,
                            fdParObj)

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                         time = cond_en$time_ms,
                         eeg_raw = as.numeric(sm_cond_en$y),
                         eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                          arrange(ID, trial)

myIDS = unique((dfresponse_en$ID))

dfresponse_merged_en <- dfresponse_en %>% left_join(demographic, by = "ID")

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)

Y_en <- as.matrix(df_wide_en[, -(1:2)])

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )

myIDS = unique((dfresponse_en$ID))

level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )

level1_en <- level1_en[, 1:5]

colnames(level1_en) <- c("ID", "cond1_f7_L1_1", "cond1_f7_L1_2", "cond1_f7_L1_3", "cond1_f7_L1_4")

level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

level2_en <- level2_en[, 1:3]

colnames(level2_en) <- c("ID", "trial", "cond1_f7_L2_1")

dfresponse_en <- dfresponse_en %>% mutate(ID = as.numeric(ID))


fpca_f7_en1 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_f7_merged_en1 <- fpca_f3_merged_en3 %>%
  left_join(fpca_f7_en1, by = c("ID", "trial"))

```


condition 2, f7


```{r cond2f7}

nodes_en <- c("f7")

cond_en <- p_df %>%
  filter(condition == 2) %>%
  dplyr::select(-sample, -condition, -af7, -af3, -f3, -fp1, -f1, -f5, -ft7)

cond_en <- as.data.frame(cond_en)

cond_en$subject <- as.numeric(as.character(cond_en$subject))
cond_en <- cond_en %>% arrange(subject)

r_len <- length(unique(cond_en$time_ms))

c_len <- n_distinct(cond_en$subject, cond_en$trial)

matrix_sm_all_en <- matrix(as.matrix(cond_en[, nodes_en]), 
                        nrow = r_len,
                        ncol = c_len * length(nodes_en),
                        byrow = FALSE)

sm_cond_en <- smooth.basis(seq(1, r_len, length.out = r_len),
                            matrix_sm_all_en,
                            fdParObj)

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                         time = cond_en$time_ms,
                         eeg_raw = as.numeric(sm_cond_en$y),
                         eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                          arrange(ID, trial)

myIDS = unique((dfresponse_en$ID))

dfresponse_merged_en <- dfresponse_en %>% left_join(demographic, by = "ID")

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)

Y_en <- as.matrix(df_wide_en[, -(1:2)])

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )

level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )

level1_en <- level1_en[, 1:5]

colnames(level1_en) <- c("ID", "cond2_f7_L1_1", "cond2_f7_L1_2", "cond2_f7_L1_3", "cond2_f7_L1_4")

level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

level2_en <- level2_en[, 1:3]

colnames(level2_en) <- c("ID", "trial", "cond2_f7_L2_1")

dfresponse_en <- dfresponse_en %>% mutate(ID = as.numeric(ID))


fpca_f7_en2 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_f7_merged_en2 <- fpca_f7_merged_en1 %>%
  left_join(fpca_f7_en2, by = c("ID", "trial"))


```


condition 3, f7

```{r cond3f7}


nodes_en <- c("f7")

cond_en <- p_df %>%
  filter(condition == 3) %>%
  dplyr::select(-sample, -condition, -af7, -af3, -f3, -fp1, -f1, -f5, -ft7)

cond_en <- as.data.frame(cond_en)

cond_en$subject <- as.numeric(as.character(cond_en$subject))
cond_en <- cond_en %>% arrange(subject)

r_len <- length(unique(cond_en$time_ms))

c_len <- n_distinct(cond_en$subject, cond_en$trial)

matrix_sm_all_en <- matrix(as.matrix(cond_en[, nodes_en]), 
                        nrow = r_len,
                        ncol = c_len * length(nodes_en),
                        byrow = FALSE)

sm_cond_en <- smooth.basis(seq(1, r_len, length.out = r_len),
                            matrix_sm_all_en,
                            fdParObj)

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                         time = cond_en$time_ms,
                         eeg_raw = as.numeric(sm_cond_en$y),
                         eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                          arrange(ID, trial)

myIDS = unique((dfresponse_en$ID))

dfresponse_merged_en <- dfresponse_en %>% left_join(demographic, by = "ID")

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)

Y_en <- as.matrix(df_wide_en[, -(1:2)])

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )

myIDS = unique((dfresponse_en$ID))

level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )

level1_en <- level1_en[, 1:5]

colnames(level1_en) <- c("ID", "cond3_f7_L1_1", "cond3_f7_L1_2", "cond3_f7_L1_3", "cond3_f7_L1_4")

level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

level2_en <- level2_en[, 1:3]

colnames(level2_en) <- c("ID", "trial", "cond3_f7_L2_1")

dfresponse_en <- dfresponse_en %>% mutate(ID = as.numeric(ID))


fpca_f7_en3 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_f7_merged_en3 <- fpca_f7_merged_en2 %>%
  left_join(fpca_f7_en3, by = c("ID", "trial"))


```

-- af7 --

condition 1, af7

```{r cond1af7}
nodes_en <- c("af7")

cond_en <- p_df %>%
  filter(condition == 1) %>%
  dplyr::select(-sample, -condition, -af3, -f3, -f7, -fp1, -f1, -f5, -ft7)


cond_en <- as.data.frame(cond_en)

cond_en$subject <- as.numeric(as.character(cond_en$subject))
cond_en <- cond_en %>% arrange(subject)

r_len <- length(unique(cond_en$time_ms))

c_len <- n_distinct(cond_en$subject, cond_en$trial)

matrix_sm_all_en <- matrix(as.matrix(cond_en[, nodes_en]), 
                        nrow = r_len,
                        ncol = c_len * length(nodes_en),
                        byrow = FALSE)

sm_cond_en <- smooth.basis(seq(1, r_len, length.out = r_len),
                            matrix_sm_all_en,
                            fdParObj)

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                         time = cond_en$time_ms,
                         eeg_raw = as.numeric(sm_cond_en$y),
                         eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                          arrange(ID, trial)

myIDS = unique((dfresponse_en$ID))

dfresponse_merged_en <- dfresponse_en %>% left_join(demographic, by = "ID")

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)

Y_en <- as.matrix(df_wide_en[, -(1:2)])

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )

myIDS = unique((dfresponse_en$ID))

level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 ) 
level1_en <- level1_en[, 1:5]

colnames(level1_en) <- c("ID", "cond1_af7_L1_1", "cond1_af7_L1_2", "cond1_af7_L1_3", "cond1_af7_L1_4")

level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

level2_en <- level2_en[, 1:3]

colnames(level2_en) <- c("ID", "trial", "cond1_af7_L2_1")

dfresponse_en <- dfresponse_en %>% mutate(ID = as.numeric(ID))


fpca_af7_en1 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_af7_merged_en1 <- fpca_f7_merged_en3 %>%
  left_join(fpca_af7_en1, by = c("ID", "trial"))


```


condition 2, af7

```{r cond2af7}


nodes_en <- c("af7")

cond_en <- p_df %>%
  filter(condition == 2) %>%
  dplyr::select(-sample, -condition, -af3, -f3, -f7, -fp1, -f1, -f5, -ft7)

cond_en$subject <- as.numeric(as.character(cond_en$subject))
cond_en <- cond_en %>% arrange(subject)

r_len <- length(unique(cond_en$time_ms))

c_len <- n_distinct(cond_en$subject, cond_en$trial)

matrix_sm_all_en <- matrix(as.matrix(cond_en[, nodes_en]), 
                        nrow = r_len,
                        ncol = c_len * length(nodes_en),
                        byrow = FALSE)

sm_cond_en <- smooth.basis(seq(1, r_len, length.out = r_len),
                            matrix_sm_all_en,
                            fdParObj)

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                         time = cond_en$time_ms,
                         eeg_raw = as.numeric(sm_cond_en$y),
                         eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                          arrange(ID, trial)

myIDS = unique((dfresponse_en$ID))

dfresponse_merged_en <- dfresponse_en %>% left_join(demographic, by = "ID")

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)

Y_en <- as.matrix(df_wide_en[, -(1:2)])

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )

myIDS = unique((dfresponse_en$ID))

level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )

level1_en <- level1_en[, 1:5]

colnames(level1_en) <- c("ID", "cond2_af7_L1_1", "cond2_af7_L1_2", "cond2_af7_L1_3", "cond2_af7_L1_4")

level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

level2_en <- level2_en[, 1:3]

colnames(level2_en) <- c("ID", "trial", "cond2_af7_L2_1")

dfresponse_en <- dfresponse_en %>% mutate(ID = as.numeric(ID))


fpca_af7_en2 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_af7_merged_en2 <- fpca_af7_merged_en1 %>%
  left_join(fpca_af7_en2, by = c("ID", "trial"))

```



condition 3, af7

```{r cond3af7}


nodes_en <- c("af7")
cond_en <- p_df %>%
  filter(condition == 3) %>%
  dplyr::select(-sample, -condition, -af3, -f3, -f7, -fp1, -f1, -f5, -ft7)

cond_en <- as.data.frame(cond_en)

cond_en$subject <- as.numeric(as.character(cond_en$subject))
cond_en <- cond_en %>% arrange(subject)

r_len <- length(unique(cond_en$time_ms))

c_len <- n_distinct(cond_en$subject, cond_en$trial)

matrix_sm_all_en <- matrix(as.matrix(cond_en[, nodes_en]), 
                        nrow = r_len,
                        ncol = c_len * length(nodes_en),
                        byrow = FALSE)

sm_cond_en <- smooth.basis(seq(1, r_len, length.out = r_len),
                            matrix_sm_all_en,
                            fdParObj)

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                         time = cond_en$time_ms,
                         eeg_raw = as.numeric(sm_cond_en$y),
                         eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                          arrange(ID, trial)

myIDS = unique((dfresponse_en$ID))

dfresponse_merged_en <- dfresponse_en %>% left_join(demographic, by = "ID")

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)

Y_en <- as.matrix(df_wide_en[, -(1:2)])

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )

myIDS = unique((dfresponse_en$ID))

level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )

level1_en <- level1_en[, 1:5]

colnames(level1_en) <- c("ID", "cond3_af7_L1_1", "cond3_af7_L1_2", "cond3_af7_L1_3", "cond3_af7_L1_4")

level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

level2_en <- level2_en[, 1:3]

colnames(level2_en) <- c("ID", "trial", "cond3_af7_L2_1")

dfresponse_en <- dfresponse_en %>% mutate(ID = as.numeric(ID))


fpca_af7_en3 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_af7_merged_en3 <- fpca_af7_merged_en2 %>%
  left_join(fpca_af7_en3, by = c("ID", "trial"))

mfpca_matrix <- fpca_af7_merged_en3

```


## Variable Selection - Elastic Net 

Set up dataset to have three fpca scores per condition per node per person.
eg cond1_fp1_fpca_1, cond2_fp1_fpca_1 etc


Apply Elastic Net 


```{r}

mfpca_allcond <- mfpca_matrix %>%
  distinct(ID, .keep_all = TRUE) %>% 
  mutate(group = ifelse(group == "Case", 1, 0),
         gender = ifelse(gender == "Female", 1, 0)) %>%
  drop_na() # TEMPORARY

set.seed(1020304050)

data_split <- group_initial_split(
  mfpca_allcond,
  group = ID,
  prop = 0.7
)

  
sch_train <- training(data_split)
sch_test <- testing(data_split)

# 10 fold CV in training to choose lambda (default). To be explicit: cv.glmnet(nfolds=10)
# glmnet() x:matrix. y: vector
x <- model.matrix(group ~ .-ID, data = mfpca_allcond)[, -1] # gets rid of intercept term
y <- mfpca_allcond$group


train_ids <- unique(sch_train$ID)
test_ids  <- unique(sch_test$ID)

train_index <- mfpca_allcond$ID %in% train_ids
test_index  <- mfpca_allcond$ID %in% test_ids

x_train <- x[train_index, ]
x_test  <- x[test_index, ]

y_train <- y[train_index]
y_test  <- y[test_index]


# table(sch_train$group)
# table(sch_test$group)



# elastic net: 0 < alpha < 1
# 0 ridge, 1 lasso
# cv to find best value of lambda and alpha 
alpha_grid <- seq(0.1, 0.9, by = 0.1)

cv_list <- lapply(alpha_grid, function(a) {
  cv.glmnet(x_train, y_train, 
            alpha = a, 
            family = "binomial")})

# extract CV error for eachg alpha 
cv_error <- sapply(cv_list, function(m) min(m$cvm))

# best model
best_index <- which.min(cv_error)
best_cv <- cv_list[[best_index]]
best_alpha <- alpha_grid[best_index]

# best lambda corresponding to that alpha
bestlam <- best_cv$lambda.min
selam <- best_cv$lambda.1se


# fit elastic net model
enet <- glmnet(x_train, y_train,
               alpha = best_alpha,
               family = "binomial")      # standardize = TRUE is the default 

# predicted probabilities with best lam from test set, type = response for pred prob (not log odds)
enet_prob <- predict(enet, s = bestlam, newx = x_test, type = "response")
# 1se lam
enet_prob_se <- predict(enet, s = selam, newx = x_test, type = "response")

# convert to vector
enet_prob_v <- as.vector(enet_prob)
enet_prob_se_v <- as.vector(enet_prob_se)


# calibration plots
valProbggplot(enet_prob_v, y_test)
valProbggplot(enet_prob_se_v, y_test)


# Brier score on test set - 1se lam
enet_brier <- BrierScore(y_test, enet_prob_se_v)

# ROC + AUC on test set
enet_roc_obj <- roc(y_test, enet_prob_se_v)
enet_auc <- auc(enet_roc_obj)
plot(enet_roc_obj, print.auc = TRUE)
coords(enet_roc_obj, "best")   # 0.33 gives a good balance between sensitivity and specificity

# Plots 
plot(enet, xvar = "lambda")
abline(v = -log(bestlam), col = "darkgreen", lty = 2, lwd = 2)
abline(v = -log(selam), col = "orange", lty = 2, lwd = 2)
legend("topleft",
       legend = c("Lambda.min", "Lambda.1se"),
       col = c("darkgreen", "orange"),
       lty = 2,
       lwd = 2,
       bty = "n")
title("Coefficients - Elastic Net", line = 3)

# CV plot
plot(best_cv, pch = 19, main = "")   # remove default title so no overlap
title("Tuning Lambda - Elastic Net", line = 3)

# plot - best alpha 
plot(alpha_grid, cv_error, type = "b", pch = 19,
     xlab = expression(alpha),
     ylab = "Cross-validated binomial deviance",
     main = "Tuning Alpha - Elastic Net")

# highlight best alpha
points(best_alpha, min(cv_error), col = "darkgreen", pch = 19, cex = 1.5)
abline(v = best_alpha, col = "darkgreen", lty = 2)


# misclassification error
enet_class <- ifelse(enet_prob_se_v > 0.33, 1, 0)
# confusion matrix (predicted, outcome)
table(enet_class, y_test)
enet_error <- mean(enet_class != y_test)


enet_error # closer to zero is better. 0.5 = random guessing
enet_brier # closer to zero is better. >0.25 is poor
enet_auc # 1 is perfect discrimination. 0.5 = random guessing

# To see which predictor was shrunk to zero (selam)
rownames(coef(enet, s = selam))[as.vector(coef(enet, s = selam)) == 0]

coef_selam <- coef(enet, s = selam)

coef_df <- data.frame(predictor = rownames(coef_selam),
                      coefficient = as.vector(coef_selam))

non_zero_coeffs <- coef_df %>% filter(coefficient != 0)
non_zero_coeffs

zero_coeffs <- coef_df %>% filter(coefficient == 0)

```

