title: "Condition Pipeline"
author: "Hollie Keating, Sarah Conway & Maeve Samali"
date: "2026-03-10"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document contains the pipeline which can be applied to any of the conditions and nodes. Included:

 - Smoothing 
 - Generate FPCA scores
 - Plot top 5 and bottom 5 of FPCA scores 1, 2, and 3 
 
 
```{r include = FALSE, warning=FALSE, message=FALSE}
library(data.table)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(plotly)
library(fda)
library(gridExtra)
library(fdapace)
library(cluster)
library(factoextra)

#setwd("C:/Users/holli/OneDrive - National University of Ireland, Galway/Modules/Research Project")
 setwd("C:/Users/35387/Documents/College/MSc Health Data Science/HDS Project/Rscript/Data")
```

```{r}
c = 1 # condition 1 
t = 1 # trial 1 
r_len = 1638 # timepoints in condition c
c_len = 32 # subjects 
node <- "fp1"

```
 
 
 
```{r}
# Loop which reads in separate excel file for each patient and compiles all data into a dataframe
# Merges time data and demogrpahic data which contains whether patients are controls or cases
missing = c(5, 8, 9, 26, 31, 35, 38)

#setwd("C:/Users/holli/OneDrive - National University of Ireland, Galway/Modules/Research Project")
setwd("C:/Users/35387/Documents/College/MSc Health Data Science/HDS Project/Rscript/Data")

times = fread("archive/time.csv")
demographic <- fread("archive/demographic.csv")
  
filenames = list.files("archive/patients/")

p_list <- list()

for(i in 1:length(filenames)){
  
  pathtomyfile = paste0("archive/patients/", filenames[i])
  temp <- fread(pathtomyfile, sep = ",", header = "auto")
  temp <- temp[ ,1:12]
  colnames(temp)<- c("subject", "trial", "condition", "sample", "fp1", "af7", "af3", "f1", "f3", "f5", "f7", "ft7")
  temp <- merge(temp, times, by ="sample")
  temp <- merge(temp, demographic, by = "subject")
  temp<- temp%>%
    filter (time_ms > -100)%>%   # time from -100ms to remove unnecessary data while including some pre beep curve info (100ms is where the beep occurs
    # ifelse(subject %in% missing, trial==15, trial==1) %>% # Only one trial out of 100 choosen. Trial 1 so patient hasn't learnt the test/timing/ keep up concealment
    filter(trial==t)
  temp<- rename(temp, schizophrenia = group)
  
  
  
  p_list[[i]] <- temp
  
}

p_df <- do.call(rbind, p_list)




p_df<- p_df%>%
  mutate(
    subject = as.factor(subject),
    schizophrenia = as.factor(schizophrenia)
  )

```
 
```{r}
## Smooth the data 
## Create a B-spline basis (mybasis) - building blocks to represent the curves
K = 100 # basis functions. More = more flexible
#?create.bspline.basis

d = 3 # basis degree e.g. 3 = cubic
mybasis = create.bspline.basis(rangeval = c(0,1638), # x-axis range for curves 0-1638 timepoints
                               nbasis = K, 
                               norder = d+1) # cubic is an order 4 polynomial 



## Smoothing parameter object

## create an fd object - functional data object

penalty_order = 2 # 2nd derivative ie penalizes wigglier
penalty = 0.05 #amount of smoothing to be applied. Larger penalty = smoother.
fdParObj <- fdPar(fd(matrix(0, K, 1), mybasis), # fd object with k=100 coeff, 1 curve
                  Lfdobj = penalty_order, 
                  lambda = penalty)
```

 
```{r}
# Smooth the data 
cond <- p_df %>%
  filter(condition == c)
str(cond$subject)

matrix_sm_all <- matrix(cond[[node]], nrow = r_len, ncol = c_len, byrow= FALSE)



## smooth the data - cond
# smooth.basis(x, y, fdParObj)
sm_cond <- smooth.basis(seq(1, r_len, length.out = r_len), # x
                            matrix_sm_all, # y
                            fdParObj)# $fd     # extracts the smoothed functional data (fd) object ie

plot(sm_cond,
     main = "Smoothed Curves")

# Colour by case vs control 
# one case/control value per curve
value_curve <- cond$schizophrenia[seq(1, nrow(cond), by = r_len)]

colours <- ifelse(value_curve == 1, "darkgreen", "orange")


plot(sm_cond$fd, col = colours, lwd = 2,
     main = "Smoothed Curves Case vs Controls")

legend("topleft",
       legend = c("Control (0)", "Case (1)"),
       col = c("orange", "darkgreen"),
       lwd = 2)


#length(value_curve)
# ncol(matrix_sm_all)


```
 
## FPCA Scores
 
```{r}
## FPCA through conditional expectation (PACE) - irregular data allowed
# FPCA via PACE 
# applied to sm_cond: condition c smoothed data 


dfresponse <- data.frame(ID = as.numeric(as.character(cond$subject)),
                         time = cond$time_ms,
                         eeg = as.numeric(sm_cond$y)) %>% 
                          arrange(ID)

####################
myIDS = unique((dfresponse$ID))
test = as.numeric(predict(sm_cond))
dfresponse$eeg_smooth = test

dfresponse %>%
  filter(ID==1) %>%
  ggplot(aes(x=time, y = eeg)) + geom_line() + geom_line(aes(y=eeg_smooth, color = "smooth"))
########################


dfresponse


# make a list containing dataframes with rows grouped by ID
list_format <- split(dfresponse, dfresponse$ID)

# create lists of time and response (eeg)
list_t <- lapply(list_format,
                 dplyr::select, time)
list_eeg <- lapply(list_format,
                   dplyr::select, eeg)

# unlist
list_t <- lapply(list_t,
                 unlist, use.names = F)
list_eeg <- lapply(list_eeg,
                   unlist, use.names = F)

# apply PACE (with fdapace)
resMEAN <- FPCA(Ly = list_eeg, Lt = list_t,
                optns = list(dataType = 'Sparse')) 

# FPC results
plot(resMEAN)
# resMEAN$xiEst #FPCA scores
```
#Smoothed FPCA Scores

```{r}
library(fdapace)

time_index <- seq_len(1638)
smooth_mat <- predict(sm_cond, time_index)

smooth_df <- as.data.frame(smooth_mat)
smooth_df$time_index <- time_index

colnames(smooth_df)[1:32] <- myIDS

smooth_long <- smooth_df %>%
  pivot_longer(
    cols = -time_index,
    names_to = "subject",
    values_to = "smooth"
  ) %>% 
  arrange(subject, time_index)



# df_long <- dfresponse %>%
#   group_by(ID) %>%
#   mutate(time_index = row_number()) %>%
#   ungroup() %>%
#   left_join(smooth_long, by = c("time_index"))

# make a list containing dataframes with rows grouped by ID
list_format <- split(smooth_long, smooth_long$subject)

# create lists of time and response (eeg)
list_t <- lapply(list_format,
                 dplyr::select, time_index)
list_eeg <- lapply(list_format,
                   dplyr::select, smooth)

# unlist
list_t <- lapply(list_t,
                 unlist, use.names = F)
list_eeg <- lapply(list_eeg,
                   unlist, use.names = F)

# apply PACE (with fdapace)
resMEAN <- FPCA(Ly = list_eeg, Lt = list_t,
                optns = list(dataType = 'Sparse')) 

# FPC results
plot(resMEAN)
scores <- resMEAN$xiEst #FPCA scores
```

 
# FPCA 1
 
```{r}
# FPCA scores

# FPC result
# resMEAN$xiEst #FPCA scores

fpca_scores <- as.data.frame(scores) %>% 
  mutate(ID =myIDS, .before = 1)             # add an ID row, first column

colnames(fpca_scores) <- c("ID", "fpca_1", "fpca_2", "fpca_3")


dfresponse <- dfresponse %>% mutate(ID = as.numeric(ID))


# Plots - merge FCP scores with df response 
fpca_scores_merged <- dfresponse %>% left_join(fpca_scores, by = "ID")


# Extract top 5 fpca 1 scores
# Top 5: 1638*5= 8190
top_5_fpca1 <- fpca_scores_merged %>%
  arrange(desc(fpca_1)) %>%
  slice(1:8190) %>%
  pull(ID)


# Extract bottom 5 fpca 1 scores
bottom_5_fpca1 <- fpca_scores_merged %>%
  arrange(fpca_1) %>%              # default ascending
  slice(1:8190) %>%
  pull(ID)

# df of top & bottom 
group_ids_fpca1 <- bind_rows(
  data.frame(ID = top_5_fpca1, group = "Top 5"),
  data.frame(ID = bottom_5_fpca1, group = "Bottom 5")
)


plot_data_fpca1 <- fpca_scores_merged %>%
  semi_join(group_ids_fpca1, by = "ID") %>%      # keeps rows that IDs are in both dfs
  left_join(group_ids_fpca1, by = "ID")          # adds the group label

plot_data_fpca1 %>% ggplot(aes(x = time, y = eeg, group = ID, colour = group)) +
  geom_line(linewidth = 0.5) +
  theme_bw() +
  labs(title = "Top vs Bottom 5 Curves FPCA 1 Scores",
       colour = "Group")
```
 
# FPCA 2 

```{r}
# Extract top 5 fpca 1 scores
# Top 5: 1638*5= 8190
top_5_fpca2 <- fpca_scores_merged %>%
  arrange(desc(fpca_2)) %>%
  slice(1:8190) %>%
  pull(ID)


# Extract bottom 5 fpca 1 scores
bottom_5_fpca2 <- fpca_scores_merged %>%
  arrange(fpca_2) %>%              # default ascending
  slice(1:8190) %>%
  pull(ID)

# df of top & bottom 
group_ids_fpca2 <- bind_rows(
  data.frame(ID = top_5_fpca2, group = "Top 5"),
  data.frame(ID = bottom_5_fpca2, group = "Bottom 5")
)


plot_data_fpca2 <- fpca_scores_merged %>%
  semi_join(group_ids_fpca2, by = "ID") %>%      # keeps rows that IDs are in both dfs
  left_join(group_ids_fpca2, by = "ID")          # adds the group label

plot_data_fpca2 %>% ggplot(aes(x = time, y = eeg, group = ID, colour = group)) +
  geom_line(linewidth = 0.5) +
  theme_bw() +
  labs(title = "Top vs Bottom 5 Curves FPCA 2 Scores",
       colour = "Group")


```


# FPCA 3 

```{r}
# Extract top 5 fpca 1 scores
# Top 5: 1638*5= 8190
top_5_fpca3 <- fpca_scores_merged %>%
  arrange(desc(fpca_3)) %>%
  slice(1:8190) %>%
  pull(ID)


# Extract bottom 5 fpca 1 scores
bottom_5_fpca3 <- fpca_scores_merged %>%
  arrange(fpca_3) %>%              # default ascending
  slice(1:8190) %>%
  pull(ID)

# df of top & bottom 
group_ids_fpca3 <- bind_rows(
  data.frame(ID = top_5_fpca3, group = "Top 5"),
  data.frame(ID = bottom_5_fpca3, group = "Bottom 5")
)


plot_data_fpca3 <- fpca_scores_merged %>%
  semi_join(group_ids_fpca3, by = "ID") %>%      # keeps rows that IDs are in both dfs
  left_join(group_ids_fpca3, by = "ID")          # adds the group label

plot_data_fpca3 %>% ggplot(aes(x = time, y = eeg, group = ID, colour = group)) +
  geom_line(linewidth = 0.5) +
  theme_bw() +
  labs(title = "Top vs Bottom 5 Curves FPCA 3 Scores",
       colour = "Group")
  





```

 
 
 - this plot shows xxx
 
 
```{r}
##KMEANS CLUSTERING

set.seed(1)
#change to df
scores <- data.frame(scores)
#scores <- na.omit(scores)


scores <- scores %>% 
  mutate(subject = myIDS)

scores <- merge(scores, demographic, by = "subject")
#scores <- na.omit(scores)

target_y <- scores$group

features_x <- scores[c(2,3,4)]

kmeans <- kmeans(features_x, centers = 4, nstart = 10)

#CONFUSION MATRIX
table(scores$group, kmeans$cluster)

#mean(scores$group == kmeans_c1$cluster)
fviz_cluster(kmeans, data = features_x, 
             main = "K-Means Clusters")
```

```{r}
##PLOTTING THE CLUSTERS
km <- data.frame(subject = myIDS, 
                 cluster = kmeans$cluster)

#smoothed_all <- merge(smooth_long, cond1,  by = "subject")

df_cluster <- merge(km, smooth_df, by= "subject") #%>% 
  #arrange(subject, time_ms)

#plotting the cluster by schizophrenia
df_cluster %>% 
  ggplot(aes(x = time_ms, y = fp1, group = subject, colour = schizophrenia)) +
  geom_line(lwd = 0.5)+
  facet_wrap(~ cluster) +
  labs(title = "EEG signmals by cluster, coloured by schizophrenia")


```


```{r}
#Reshape data for multi node glm 


nodes <- c("fp1", "af7", "af3", "f1", "f3", "f5", "f7", "ft7")

nodes_df <- p_df %>% pivot_longer(
  cols = all_of(nodes),
  names_to = "node",
  values_to = "EEG"
) %>% dplyr::select(subject, time_ms, schizophrenia, node, EEG) %>% 
  arrange(subject, node, time_ms)

```

```{r  glm}

subject_df <- full_merged_df %>%
  distinct(ID, group, gender, age, education, fpca_1, fpca_2, fpca_3)

subject_df$group <- factor(subject_df$group)
subject_df$gender <- factor(subject_df$gender)




#=================
# Gender Model 
# ================

log_gender <- glm(group ~ gender + age + education + fpca_1 + fpca_2 + fpca_3,
             data = subject_df,
             family = binomial)

summary(log_gender)

# summary table
tab_model( log_gender, transform = "exp")

# Sound checking gender in dataset

ggplot(subject_df, aes(x = gender, fill = group)) +
  geom_bar(position = "dodge") +
  labs(x = "Gender", y = "Count", fill = "Group")


# large discrepency between males and females within condition 1 group (more males, majority in control group)




# ====================================
# Education Model
# ====================================

model_edu <- glm(group ~ education,
                 data = subject_df,
                 family = binomial)

summary(model_edu)

# summary table
tab_model(model_edu, transform = "exp")


# creating a boxplot 

ggplot(subject_df, aes(x = factor(group), y = education)) +
  geom_boxplot() +
  labs(x = "Group", y = "Education")



# =====================
# FPCA Model 
# =====================


model_eeg <- glm(group ~ fpca_1 + fpca_2 + fpca_3,
                 data = subject_df,
                 family = binomial)

summary(model_eeg)

# summary table

tab_model(model_eeg, transform = "exp")


ggplot(subject_df, aes(x = factor(group), y = fpca_1)) +
  geom_boxplot() +
  labs(x = "Group", y = "FPCA 1")

plot_model(model_eeg, type = "pred", terms = "fpca_1")
plot_model(model_eeg, type = "pred", terms = "fpca_2")
plot_model(model_eeg, type = "pred", terms = "fpca_3")

# ===============
# Checking Education in Dataset
# ===============

ggplot(subject_df, aes(x = education)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  labs(x = "Years of Education", y = "Count") +
  facet_wrap(subject_df$group)

table(subject_df$group, subject_df$education)


```

### Education Model

Education was significantly associated with schizophrenia diagnosis, with higher education linked to lower log-odds (OR = 0.57, p = 0.018), indicating a decrease in odds of approximately 43% per additional year of education.

By analysing the grouped bar chart for education, individuals with schizophrenia tended to have lower levels of education compared to controls - however this association may reflect the impact of earlier illness onset or other cofounding factors.


### Gender Model

Gender was not significant within association with schizophrenia presence (OR = 0.15, p = 0.213). 
The distribution of gender was similar across the control and schizophrenia groups, with a higher proportion of males in both groups which supports our above result. However, the sample was predominantly male which may limit our ability to detect gender differences. 


### FPCA Model

FPCA 1 was significantly associated with schizophrenia diagnosis (p = 0.039), although the the odds ratio was quite small (OR \approx 1.00). FPCA 2 showed no association, and FPCA 3 showed a weak non-significant trend.  
 
