---
title: "Condition Pipeline"
author: "Hollie Keating, Sarah Conway & Maeve Samali"
date: "2026-03-10"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document contains the pipeline which can be applied to any of the conditions and nodes. Included:

 - Smoothing 
 - Generate FPCA scores
 - Plot top 5 and bottom 5 of FPCA scores 1, 2, and 3 
 
```{r, libraries, include = FALSE, warning=FALSE, message=FALSE}
library(data.table)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(plotly)
library(fda)
library(gridExtra)
library(fdapace)
library(tidyr)
library(table1)
library(factoextra)
library(cluster)
library(sjPlot)
library(refund)
library(glmnet)
library(pROC)  
library(DescTools) 
library(CalibrationCurves)
library(rsample)

#setwd("C:/Users/holli/OneDrive - National University of Ireland, Galway/Modules/Research Project")
#setwd("C:/Users/35387/Documents/College/MSc Health Data Science/HDS Project/Rscript/Data")
#setwd("C:/Users/Sarah Conway/OneDrive/part 2/OneDrive/MSc-Project")
```

```{r, pipeline parameters}
c = 1 # condition 1 
t = 50 # multiple trials 
t1 = 1 # selecting one trial
node <- "fp1"
```

 
```{r, read in files}
# Loop which reads in separate excel file for each patient and compiles all data into a dataframe
# Merges time data and demogrpahic data which contains whether patients are controls or cases

#setwd("C:/Users/holli/OneDrive - National University of Ireland, Galway/Modules/Research Project")
setwd("C:/Users/35387/Documents/College/MSc Health Data Science/HDS Project/Rscript/Data")

times = fread("archive/time.csv")
demographic <- fread("archive/demographic.csv")
  
filenames = list.files("archive/patients/")

p_list <- list()

for(i in 1:length(filenames)){
  
  pathtomyfile = paste0("archive/patients/", filenames[i])
  temp <- fread(pathtomyfile, sep = ",", header = "auto")
  temp <- temp[ ,1:12]
  colnames(temp)<- c("subject", "trial", "condition", "sample", "fp1", "af7", "af3", "f1", "f3", "f5", "f7", "ft7")
  temp <- merge(temp, times, by ="sample")
  temp <- merge(temp, demographic, by = "subject")
  temp<- temp%>%
    filter (time_ms > -100)%>%   # time from -100ms to remove unnecessary data while including some pre beep curve info (100ms is where the beep occurs
    # ifelse(subject %in% missing, trial==15, trial==1) %>% # Only one trial out of 100 choosen. Trial 1 so patient hasn't learnt the test/timing/ keep up concealment
    filter(trial<=t)
  temp<- rename(temp, schizophrenia = group)

  p_list[[i]] <- temp
  
}

p_df <- do.call(rbind, p_list)

p_df<- p_df%>%
  mutate(
    subject = as.factor(subject),
    schizophrenia = as.factor(schizophrenia)
  )

```

```{r splitting dataset}
set.seed(1)

p_df_full <- p_df

p_df_data_split <- group_initial_split(
  p_df_full,
  group = subject,
  strata = schizophrenia,
  prop = 0.7
)

pdf_train <- training(p_df_data_split)
pdf_test <- testing(p_df_data_split)

train_ids <- as.numeric(as.character(unique(pdf_train$subject)))
test_ids  <- as.numeric(as.character(unique(pdf_test$subject)))

# choose either train or test dataset
p_df<- pdf_train

demographic <- demographic %>%
  rename(ID = subject) %>%
  mutate(group = as.character(group),
         gender = stringr::str_trim(as.character(gender)), # to handle space
        group = recode(group, `0` = "Control", `1` = "Case"),
        gender = recode(gender, "M" = "Male", "F" = "Female"),
        group = factor(group, levels = c("Control", "Case")),
        gender = factor(gender, levels = c("Male", "Female")))


train_index <- demographic$ID %in% train_ids
test_index <- demographic$ID %in% test_ids

demo_train <- demographic[train_index, ]
demo_test <- demographic[test_index, ]

```

 
```{r, smoothing splines 1}
## Smooth the data 
## Create a B-spline basis (mybasis) - building blocks to represent the curves
K = 100 # basis functions. More = more flexible
d = 3 # basis degree e.g. 3 = cubic

mybasis = create.bspline.basis(rangeval = c(0,1638), # x-axis range for curves 0-1638 timepoints
                               nbasis = K, 
                               norder = d+1) # cubic is an order 4 polynomial 


## Smoothing parameter object

## create an fd object - functional data object

penalty_order = 2 # 2nd derivative ie penalizes wigglier
penalty = 0.05 #amount of smoothing to be applied. Larger penalty = smoother.
fdParObj <- fdPar(fd(matrix(0, K, 1), mybasis), # fd object with k=100 coeff, 1 curve
                  Lfdobj = penalty_order, 
                  lambda = penalty)
```

 
```{r, smoothing splines 2}
# Smooth the data 
cond <- p_df %>%
  filter(condition == c) %>%
  filter(trial == t1)

#the subjects are not ordered correctly to fix this
cond$subject <- as.numeric(as.character(cond$subject))
cond <- cond %>% arrange(subject)

r_len <- nrow(cond %>% filter(subject == 1))
c_len <- length(unique(cond$subject))

matrix_sm_all <- matrix(cond[[node]], nrow = r_len, ncol = c_len, byrow= FALSE)

## smooth the data - cond
# smooth.basis(x, y, fdParObj)
sm_cond <- smooth.basis(seq(1, r_len, length.out = r_len), # x
                            matrix_sm_all, # y
                            fdParObj)# $fd     # extracts the smoothed functional data (fd) object ie
plot(sm_cond,
     main = "Smoothed Curves")

```




```{r, df}
# Plots 
# dfresponse - raw eeg, smoothed eeg, merged demographic 
 
# sm_cond$y is raw eeg signals
# smoothed is predict(sm_cond) 
# fd is basis and coeff

dfresponse <- data.frame(ID = as.numeric(as.character(cond$subject)),
                         time = cond$time_ms,
                         eeg_raw = as.numeric(sm_cond$y),
                         eeg_smooth = as.numeric(predict(sm_cond))) %>%   # smoothed predictions
                          arrange(ID)


# table(demographic$group) # 25 controls, 11 cases
# table(demographic$gender) # 30 males. 6 females

myIDS = unique((dfresponse$ID))
#test = as.numeric(predict(sm_cond)) # smoothed predictions
#dfresponse$eeg_smooth = test

dfresponse_merged <- dfresponse %>% left_join(demographic, by = "ID")


```

```{r, demographic plots, fig.width=9, fig.asp=1.5}

# plot color = cases vs controls 
# group = ID: draw one seperate line for each patient ID
group_plot <- dfresponse_merged %>%
  ggplot(aes(x = time, y = eeg_smooth, group = ID, color = group)) +
  geom_line(alpha = 0.6) +
  scale_color_manual(values = c("Control" = "orange", "Case" = "darkgreen")) +
  theme_bw()+
  labs(title = "Smoothed Curves Case vs Controls",
       x = "Time",
       y = "EEG Signals",
       color = "")

#group_plot


gender_plot <- dfresponse_merged %>%
  ggplot(aes(x = time, y = eeg_smooth, group = ID, color = group)) +
  geom_line(alpha = 0.6) +
  scale_color_manual(values = c("Control" = "orange", "Case" = "darkgreen")) +
  theme_bw()+
  labs(title = "Smoothed Curves Case vs Controls",
       x = "Time",
       y = "EEG Signals",
       color = "")+facet_wrap(~gender)

#gender_plot


# plot by age > 40 or < 40     its not the age they werre diagnosed at its just the age they are now.................
age_plot <- dfresponse_merged %>% 
  mutate(age_grouped = ifelse(age < 40, "<40", ">=40")) %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = ID, color = group)) +
  geom_line(alpha = 0.6) +
  scale_color_manual(values = c("Control" = "orange", "Case" = "darkgreen")) +
  theme_bw()+
  labs(title = "Smoothed Curves by Age",
       x = "Time",
       y = "EEG Signals",
       color = "")+facet_wrap(~age_grouped)

#age_plot

# years in education 
# primary plus secondary is ~14 years 
edu_plot <- dfresponse_merged %>% 
  mutate(education_grouped = ifelse(education < 15, "<15", ">=15")) %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = ID, color = group)) +
  geom_line(alpha = 0.6) +
  scale_color_manual(values = c("Control" = "orange", "Case" = "darkgreen")) +
  theme_bw()+
  labs(title = "Smoothed Curves by Years in Education",
       x = "Time",
       y = "EEG Signals",
       color = "")+facet_wrap(~education_grouped)

#edu_plot

grid.arrange(group_plot, gender_plot, age_plot, edu_plot, ncol = 1)

# table_data <- dfresponse_merged %>%
#   distinct(ID, group, gender)      # one per patient
# 
# table1(~ gender | group,
#        data = table_data)


# demographic only has up to patient 77


dfresponse %>%
  filter(ID==1) %>%
  ggplot(aes(x=time, y = eeg_raw)) + geom_line() + geom_line(lwd = 1.5, aes(y=eeg_smooth, color = "smooth"))


```

 
 
## FPCA Scores
 
```{r, fpca scores}
## FPCA through conditional expectation (PACE) - irregular data allowed
# FPCA via PACE 
# applied to sm_cond: condition c smoothed data 

# make a list containing dataframes with rows grouped by ID
list_format <- split(dfresponse, dfresponse$ID)

# create lists of time and response (eeg)
list_t <- lapply(list_format,
                 dplyr::select, time)
list_eeg <- lapply(list_format,
                   dplyr::select, eeg_raw)

# unlist
list_t <- lapply(list_t,
                 unlist, use.names = F)
list_eeg <- lapply(list_eeg,
                   unlist, use.names = F)

# apply PACE (with fdapace)
resMEAN <- FPCA(Ly = list_eeg, Lt = list_t,
                optns = list(dataType = 'Sparse')) #Dense or p >> n (time pts > n patients)

# ##is the data dense?
# plot(FPCA(Ly = list_eeg, Lt = list_t,
#                 optns = list(dataType = 'Dense')))
# FPC results
plot(resMEAN)
#resMEAN$xiEst #FPCA scores
```
 
 
## Highest & Lowest Scores 
 
```{r, fpc df}
# FPCA scores

# FPC result
# resMEAN$xiEst #FPCA scores

fpca_scores <- as.data.frame(resMEAN$xiEst) %>% 
  mutate(ID =myIDS, .before = 1)             # add an ID row, first column

colnames(fpca_scores) <- c("ID", "fpca_1", "fpca_2", "fpca_3")

dfresponse <- dfresponse %>% mutate(ID = as.numeric(ID))

# Plots - merge FCP scores with df response 
fpca_scores_merged <- dfresponse_merged %>% left_join(fpca_scores, by = "ID")
```

```{r, fpc plots function }
# Extract top 5 fpc scores
# Top 5: 1638*5= 8190

fpca_plots <- function(df = fpca_scores_merged, fpc){
  
  fpca_name <- as.character(substitute(fpc))
  
  top_5_fpc <- fpca_scores_merged %>%
  arrange(desc({{fpc}})) %>%
  slice(1:8190) %>%
  pull(ID)

  # Extract bottom 5 fpca 1 scores
  bottom_5_fpc <- fpca_scores_merged %>%
    arrange({{fpc}}) %>%              # default ascending
    slice(1:8190) %>%
    pull(ID)
  
  # df of top & bottom 
  group_ids_fpc <- bind_rows(
    data.frame(ID = unique(top_5_fpc), top_bot = "Top 5"),
    data.frame(ID = unique(bottom_5_fpc), top_bot = "Bottom 5")
  )
  
  plot_data_fpc <- fpca_scores_merged %>%
    semi_join(group_ids_fpc, by = "ID") %>%      # keeps rows that IDs are in both dfs
    left_join(group_ids_fpc, by = "ID")          # adds the group label
  
  plot_data_fpc %>% ggplot(aes(x = time, y = eeg_raw, group = ID, colour = top_bot)) +
    geom_line(linewidth = 0.5) +
    theme_bw() +
    labs(title = paste("Top vs Bottom 5 Curves Scores:", fpca_name),
         colour = "Group")
}
```
 
```{r, fpca plots}
fpca_plots(fpc = fpca_1)
fpca_plots(fpc = fpca_2)
fpca_plots(fpc = fpca_3)
```
 
 
```{r, kmeans}
##KMEANS CLUSTERING
set.seed(1)

#scores <- na.omit(scores)
rownames(fpca_scores)

fpca_scores_demo <- merge(fpca_scores, demographic, by = "ID")

target_y <- fpca_scores_demo$group
features_x <- fpca_scores_demo[c(2,3,4)]

kmeans_c1 <- kmeans(features_x, centers = 4, nstart = 10)

#CONFUSION MATRIX
table(fpca_scores_demo$group, kmeans_c1$cluster)

#mean(scores$group == kmeans_c1$cluster)
fviz_cluster(kmeans_c1, data = features_x, 
             main = "K-Means Clusters")
```

```{r, plot kmeans}
##PLOTTING THE CLUSTERS
km <- data.frame(ID = fpca_scores_demo$ID, 
                 cluster = kmeans_c1$cluster)

#smoothed_all <- merge(smooth_long, cond1,  by = "subject")

df_cluster <- merge(km, fpca_scores_merged, by= "ID") #%>% 
  #arrange(subject, time_ms)

#plotting the cluster by schizophrenia
df_cluster %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = ID, colour = group)) +
  geom_line(lwd = 0.5)+
  facet_wrap(~ cluster) +
  labs(title = "EEG signals by cluster, coloured by schizophrenia") %>%
  show()
```


```{r, glms}
subject_df <- fpca_scores_merged %>%
  distinct(ID, group, gender, age, education, fpca_1, fpca_2, fpca_3)

# subject_df$group <- factor(subject_df$group)
# subject_df$gender <- factor(subject_df$gender)


#=================
# Gender Model 
# ================

log_gender <- glm(group ~ age + education + fpca_1 + fpca_2 + fpca_3,
             data = subject_df,
             family = binomial)

summary(log_gender)


## converting coefficients into OR

exp(coef(log_gender))


# from summary, FPCA1 shows most significance at p = 0.06.., genderM significance = 0.2131
# gender coef = 0.1485062


# ====================================
# Education Model
# ====================================

model_edu <- glm(group ~ education,
                 data = subject_df,
                 family = binomial)

summary(model_edu)
exp(coef(model_edu))

## for each additional year of education, the odds of being in the schizophrenia group decrease by about 57%

## creating a boxplot 

ggplot(subject_df, aes(x = factor(group), y = education, color = factor(group))) +
  geom_boxplot() +
  labs(x = "Group", y = "Education")

# =====================
# FPCA Model 
# ====================

model_eeg <- glm(group ~ fpca_1 + fpca_2 + fpca_3,
                 data = subject_df,
                 family = binomial)

summary(model_eeg)

exp(coef(model_eeg))

# 1st PC most significant, OR = 0.99.. so only marginally reduces odds of being in schizophrenia group with increase of PC1

ggplot(subject_df, aes(x = factor(group), y = fpca_1, color = factor(group))) +
  geom_boxplot() +
  labs(x = "Group", y = "FPCA 1")

```

### Education Model

Education was significantly associated with schizophrenia diagnosis, with higher education linked to lower log-odds (OR = 0.57, p = 0.018), indicating a decrease in odds of approximately 43% per additional year of education.

By analysing the grouped bar chart for education, individuals with schizophrenia tended to have lower levels of education compared to controls - however this association may reflect the impact of earlier illness onset or other cofounding factors.


### Gender Model

Gender was not significant within association with schizophrenia presence (OR = 0.15, p = 0.213). 
The distribution of gender was similar across the control and schizophrenia groups, with a higher proportion of males in both groups which supports our above result. However, the sample was predominantly male which may limit our ability to detect gender differences. 

### FPCA Model

FPCA 1 was significantly associated with schizophrenia diagnosis (p = 0.039), although the the odds ratio was quite small (OR \approx 1.00). FPCA 2 showed no association, and FPCA 3 showed a weak non-significant trend.  
 
# MFPCA Model 

So now we want to take into account that we have more than 1 function per person. People completed each sensory condition up to 100 times. Including more of these trials will improve our ability to predict whether an EEG is typical of schizophrenia or not.
 
## Smoothing multiple trials

As before we need to smooth the data, but this time for all the trials.

```{r, smoothing function}
smooth_trial_data <- function(df, cond_num, node){
  
  # Smooth the data 
  cond <- df %>%
    filter(condition == cond_num)  %>% 
    arrange(trial)
  
  #the subjects are not ordered correctly to fix this
  cond$subject <- as.numeric(as.character(cond$subject))
  cond <- cond %>% arrange(subject)
  
  # sub <- unique(cond$subject)[1] #1st subject in df
  # ##r_len = number of time points for each curve
  # r_len <- nrow(cond %>% filter(subject == sub) %>% filter(trial == 1))
  r_len <- length(unique(cond$time_ms))
  #number of curves per person
  #number of individual curves
  c_len <- nrow(cond %>% distinct(subject, trial))
  
  matrix_sm_all <- matrix(cond[[node]], nrow = r_len, ncol = c_len, byrow= FALSE)
  
  ## smooth the data - cond
  # smooth.basis(x, y, fdParObj)
  sm_trials <- smooth.basis(seq(1, r_len, length.out = r_len), # x
                              matrix_sm_all, # y
                              fdParObj)# $fd     # extracts the smoothed functional data (fd) object ie
 x <- list(cond, sm_trials)
  return(x)
}

```


```{r, plot mfpca smoothed}
sm_trials <- smooth_trial_data(df = p_df, cond_num = 1, node = "fp1")[[2]]
cond_trials <- smooth_trial_data(df = p_df, cond_num = 1, node = "fp1")[[1]]
plot(sm_trials,
      main = "Smoothed Curves")
```

### Data in long format & plots

We create a dataframe of all the relevant measurements and we have plotted the smoothed data for each trial for each person below.

```{r, mfpca plots}
##create a response df in long format
dfresponse_trial <- data.frame(ID = as.numeric(as.character(cond_trials$subject)),
                         time = cond_trials$time_ms,
                         trial = cond_trials$trial,
                         eeg_raw = as.numeric(sm_trials$y),
                         eeg_smooth = as.numeric(predict(sm_trials)),
                         schizophrenia = cond_trials$schizophrenia) %>%   # smoothed predictions
                         arrange(ID)

##Plot all trials for each subject 
dfresponse_trial  %>% filter(ID <= 10) %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = factor(trial), color = factor(schizophrenia))) +
  geom_line() +
  facet_wrap(~ ID, nrow = 2) +
  ggtitle("Multilevel functional observations")

dfresponse_trial  %>% filter(ID > 10 & ID <=20) %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = factor(trial), color = factor(schizophrenia))) +
  geom_line() +
  facet_wrap(~ ID, nrow = 2) +
  ggtitle("Multilevel functional observations")

dfresponse_trial  %>% filter(ID > 20 & ID <= 71) %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = factor(trial), color = factor(schizophrenia))) +
  geom_line() +
  facet_wrap(~ ID, nrow = 2) +
  ggtitle("Multilevel functional observations")

dfresponse_trial  %>% filter(ID > 71) %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = factor(trial), color = factor(schizophrenia))) +
  geom_line() +
  scale_color_manual(values = c("#00BFC4"))+
  facet_wrap(~ ID, nrow = 2) +
  ggtitle("Multilevel functional observations")
  
```


## mFPCA

Now we perform multi-level modelling, where the first level is the individual subject level and the second level is the individual trial level. So we get a score for each person and for each curve (trial).

```{r, mfpca}
# df: Data in long format with columns, subject Id = Id, stride_id (wave_id) = visit, arg = time (seq(0, 1, length = 101), obs = y values
#Output: each row is a curve with subject Id and wave id
df_wide <- pivot_wider(dfresponse_trial[,-4], names_from = time, values_from = eeg_smooth)
Y <- as.matrix(df_wide[, c(4:1641)]) # excluding column id, trial, schizophrenia

#mfpca
mlfpc <- mfpca.face(Y = Y, # matrix- each row is a curves
                    id = df_wide$ID, # subject id: c(1,1,2,2,2,3,3 ...)
                    visit = df_wide$trial, # trial id c(1,2,3,.., 1, 2,3, 4, ...)
                    twoway = TRUE,
                    #pve2 = 0.75, # pve for level 2
                    pve = 0.95
                    )

```

## Multlevel FPCs

-   Level 1

```{r, mfpca level 1}
t <- unique(dfresponse_trial$time)
egf1 <- data.frame(mlfpc$efunctions$level1)
ggplot(egf1) +
  geom_line(lwd =0.55, aes(x = t, y = egf1[, 1], color = " 1")) +
  geom_line(lwd =0.55,aes(x = t, y = egf1[, 2], color = " 2")) +
  geom_line(lwd =0.55,aes(x = t, y = egf1[, 3], color = " 3")) +
  geom_line(lwd =0.55,aes(x = t, y = egf1[, 4], color = " 4")) +
  scale_color_manual(values = c(" 1" = "blue", " 2" = "green", " 3" = "red", " 4" = "purple")) +
  labs(x = "Time", y = "value", color = "mFPC") +
  theme_bw()

```


-   Level 2

```{r, mfpca level 2}
t <- unique(dfresponse_trial$time)
egf2 <- data.frame(mlfpc$efunctions$level2)
ggplot(egf2) +
  geom_line(aes(x = t, y = egf2[, 1], color = " 1")) +
  geom_line(aes(x = t, y = egf2[, 2], color = " 2")) +
  geom_line(aes(x = t, y = egf2[, 3], color = " 3")) +
  geom_line(aes(x = t, y = egf2[, 4], color = " 4")) +
  scale_color_manual(values = c(" 1" = "blue", " 2" = "green", " 3" = "red", " 4" = "purple")) +
  labs(x = "Time (proportion)", y = "value", color = "mFPC") +
  theme_bw()

```


## Nature of the variability

Level 1

```{r, eigenfns 1}
# add and subtract eigenfunction to the overall mean
egf_l1 <- egf1 %>%  mutate(mu = mlfpc$mu,
                     p1 = mlfpc$mu  + X1*sqrt(mlfpc$evalues$level1[1]),
                     n1 = mlfpc$mu - X1*sqrt(mlfpc$evalues$level1[1]),
                     
                     p2 = mlfpc$mu  + X2*sqrt(mlfpc$evalues$level1[2]),
                     n2 = mlfpc$mu - X2*sqrt(mlfpc$evalues$level1[2]),
                     
                     p3 = mlfpc$mu  + X3*sqrt(mlfpc$evalues$level1[3]),
                     n3 = mlfpc$mu - X3*sqrt(mlfpc$evalues$level1[3]),
                     
                     p4 = mlfpc$mu  + X4*sqrt(mlfpc$evalues$level1[4]),
                     n4 = mlfpc$mu - X4*sqrt(mlfpc$evalues$level1[4]))

```



# First and second modes
```{r, level 1 plot}
# first mode
ggplot(egf_l1, aes(x = t, y = mu)) + geom_line(color = "red", lwd = 0.8) +
geom_point(aes(x = t, y = p1), shape = "+", size = 3) + 
geom_point(aes(x = t, y = n1), shape = "-", size = 3)  +
  labs(y ="Y", x = "Time (proportion)") +ggtitle("PC 1 (..%)") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),plot.title.position = "panel"
  )

# Second mode
ggplot(egf_l1, aes(x = t, y = mu)) + geom_line(color = "red", lwd = 0.8) +
geom_point(aes(x = t, y = p2), shape = "+", size = 3) + 
geom_point(aes(x = t, y = n2), shape = "-", size = 3)  +
  labs(y ="Y", x = "Time (proportion)") + ggtitle("PC 2 (...%)") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),plot.title.position = "panel"
  )

# third mode
ggplot(egf_l1, aes(x = t, y = mu)) + geom_line(color = "red", lwd = 0.8) +
geom_point(aes(x = t, y = p3), shape = "+", size = 3) + 
geom_point(aes(x = t, y = n3), shape = "-", size = 3)  +
  labs(y ="Y", x = "Time (proportion)") + ggtitle("PC 3 (...%)") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),plot.title.position = "panel"
  )
```


 level 2

First and second modes

```{r, fpc level 2}
egf2 <- egf2 %>%  mutate(mu = mlfpc$mu,
                       p1 = mlfpc$mu  + X1*sqrt(mlfpc$evalues$level2[1]),
                       n1 = mlfpc$mu - X1*sqrt(mlfpc$evalues$level2[1]),
                       
                       p2 = mlfpc$mu  + X2*sqrt(mlfpc$evalues$level2[2]),
                       n2 = mlfpc$mu - X2*sqrt(mlfpc$evalues$level2[2]),
                       
                       p3 = mlfpc$mu  + X3*sqrt(mlfpc$evalues$level2[3]),
                       n3 = mlfpc$mu - X3*sqrt(mlfpc$evalues$level2[3]),
                       
                       p4 = mlfpc$mu  + X4*sqrt(mlfpc$evalues$level2[4]),
                       n4 = mlfpc$mu - X4*sqrt(mlfpc$evalues$level2[4]))
```


```{r, fpc level 2 plots}
# First mode
ggplot(egf2, aes(x = t, y = mu)) + geom_line(color = "red", lwd = 0.8) +
geom_point(aes(x = t, y = p1), shape = "+", size = 3) + 
geom_point(aes(x = t, y = n1), shape = "-", size = 3)  +
  labs(y ="Y", x = "Time (proportion)") +
   ggtitle("PC 1 (....%)") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),plot.title.position = "panel"
  )

# Second mode
ggplot(egf2, aes(x = t, y = mu)) + geom_line(color = "red", lwd = 0.8) +
geom_point(aes(x = t, y = p2), shape = "+", size = 3) + 
geom_point(aes(x = t, y = n2), shape = "-", size = 3)  +
  labs(y ="Y", x = "Time (proportion)") +
   ggtitle("PC 2 (....%)") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),plot.title.position = "panel"
  )

# Third mode
ggplot(egf2, aes(x = t, y = mu)) + geom_line(color = "red", lwd = 0.8) +
geom_point(aes(x = t, y = p3), shape = "+", size = 3) + 
geom_point(aes(x = t, y = n3), shape = "-", size = 3)  +
  labs(y ="Y", x = "Time (proportion)") +
   ggtitle("PC 3 (....%)") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),plot.title.position = "panel"
  )
```


### Multilevel FPC scores

#### Level 1

Patient Level scores - 1 per person

```{r, mfpc1}
level1_scores <- cbind(data.frame(mlfpc$scores$level1), subj_id = as.factor(unique(df_wide$ID)), schizophrenia = df_wide$schizophrenia[match(unique(df_wide$ID), df_wide$ID)])# subject level scores

ggplot(level1_scores,aes(x = X1, y = X2, color = factor(schizophrenia))) + geom_point() +
  labs(x = "mFPC1", y = "mFPC2", title = "..") + theme_bw()
```

#### Level 2

Individual Trial scores (i.e. 1 score per function/trial) -multiple per person

```{r, mfpc2}
level2_scores <- cbind(data.frame(mlfpc$scores$level2), subj_id = factor(df_wide$ID), trial_id = df_wide$trial, schizophrenia = factor(df_wide$schizophrenia))

ggplot(level2_scores, aes(x = trial_id, y = X1, group = subj_id, color = schizophrenia)) +
  geom_line() + geom_point() + theme_bw() +
  labs(y = "mFPC1", x = "trials order - longitudinal", title = "") 
 
ggplot(level2_scores, aes(x = subj_id, y = X1, group = subj_id, colour = schizophrenia)) + geom_boxplot() + theme_bw() +
  labs(y = "mFPC1", x = "Subject Id", title = "Score distribution") 
 

```


## Predicting Test Subject's Scores

```{r, predicting mfpca scores function}
predict_scores <- function(Y, id, visit, mfpca_obj){
  # Input: Y -  Nxt matrix of new curves
  #        id - a vector of repeated subject id with length = number of new curves
  #       visit - a vector of F/NF with length = number of curves
  #       mfpca_obj - mfpca.face object
  
  visit <- as.factor(visit)
  id <- as.factor(id)
  df <- data.frame(id = id, visit = visit, Y = I(Y))
  rm(id, visit, Y)
  J <- length(levels(df$visit))
  L <- ncol(df$Y)
  nVisits <- data.frame(table(df$id))
  colnames(nVisits) = c("id", "numVisits")
  ID = sort(unique(df$id))
  I <- length(ID)
  argvals <- seq(0, 1, length.out = L)
  
  # Population and visit-specific mean functions - from the mfpca.face object
  mu <- mfpca_obj$mu
  eta <- mfpca_obj$eta
  # Ytilde = Y - mu - eta
  # mueta = mu + eta
  mueta <- mu + eta
  
  # Ytilde for the new curves
  Ytilde <- matrix(NA, nrow = nrow(df$Y), ncol = ncol(df$Y))
  for (j in 1:J) {
    ind_j <- which(df$visit == levels(df$visit)[j])
    Ytilde[ind_j, ] <- df$Y[ind_j, ] - matrix(mueta[, 
                                                    j], nrow = length(ind_j), ncol = L, byrow = TRUE)
  }
  
  df$Ytilde <- I(Ytilde)
  rm(Ytilde)
  
  efunctions <- mfpca_obj$efunctions
  npc <- mfpca_obj$npc
  sigma2 <- mfpca_obj$sigma2
  evalues <-mfpca_obj$evalues
  
  # Estimate principal component scores
  phi1 <- efunctions[[1]]
  phi2 <- efunctions[[2]]
  score1 <- matrix(0, I, npc[[1]])
  score2 <- matrix(0, nrow(df$Y), npc[[2]])
  unVisits <- unique(nVisits$numVisits)
  
  #........... for I number of subjects - here I = 1 (one subject)
  for (m in 1:I) {
    Jm <- nVisits[m, 2]
    A <- Jm * (t(phi1) %*% phi1)/sigma2 + diag(1/evalues[[1]])
    B = matrix(rep(t(phi1) %*% phi2/sigma2, Jm), nrow = npc[[1]])
    temp = ginv(t(phi2) %*% phi2/sigma2 + diag(1/evalues[[2]]))
    
    C <- t(B)
    invD <- kronecker(diag(1, Jm), temp)
    MatE <- ginv(A - B %*% invD %*% C)
    MatF <- -invD %*% C %*% MatE
    MatG <- -MatE %*% B %*% invD
    MatH <- invD - invD %*% C %*% MatG
    Mat1 <- cbind(MatE, MatG)
    Mat2 <- cbind(MatF, MatH)
    
    # 
    int1 <- colSums(matrix(df$Ytilde[df$id == ID[m], ], ncol = L) %*% phi1) 
    int2 <- matrix(df$Ytilde[df$id == ID[m], ], ncol = L) %*% phi2
    
    int <- c(int1, as.vector(t(int2)))/sigma2
    
    score1[m, ] <- Mat1 %*% int
    score2[which(df$id == ID[m]), ] <- matrix(Mat2 %*% int, ncol = npc[[2]], byrow = TRUE)
    
  }
  
  scores <- list(level1 = score1, level2 = score2)
  
  return(scores)
  
}

```

```{r, predict scores}
node_test <- c("fp1")

sm_test <- smooth_trial_data(df = pdf_test, cond_num = 1, node = "fp1")[[2]]
cond_test <- smooth_trial_data(df = pdf_test, cond_num = 1, node = "fp1")[[1]]

dfresponse_test <- data.frame(ID = as.numeric(as.character(cond_test$subject)),
                         time = cond_test$time_ms,
                         trial = cond_test$trial,
                         eeg_raw = as.numeric(sm_test$y),
                         eeg_smooth = as.numeric(predict(sm_test)),
                         schizophrenia = cond_test$schizophrenia) %>% 
                         arrange(ID)

df_wide_test <- pivot_wider(dfresponse_test[,-4], #exclude the raw eeg column
                       names_from = time,
                       values_from = eeg_smooth)
n <- ncol(df_wide_test)
Y_test <- as.matrix(df_wide_test[, c(4:n)]) # excluding column id, trial, schizophrenia

myIDS_test = unique((dfresponse_test$ID))

predicted_scores <- predict_scores(Y_test, df_wide_test$ID, df_wide_test$trial, mlfpc)

#df of scores level1 & 2
level1_test <- as.data.frame(predicted_scores[[1]]) %>%
  mutate(ID = myIDS_test, .before =1 )
level2_test <- as.data.frame(predicted_scores[[2]]) %>%
  mutate(ID = df_wide_test$ID, trial = df_wide_test$trial, .before = 1 )

l1 <- ncol(as.data.frame(predicted_scores[[1]])) # number of fpcs
colnames(level1_test) <- c("ID", c(paste("cond1_", node_test, "_L1_", 1:l1, sep = "")))

l2 <- ncol(as.data.frame(predicted_scores[[2]])) 
colnames(level2_test) <- c("ID", "trial", c(paste("cond1_", node_test, "_L2_", 1:l2, sep = "")))
```


# edited all trials

- Creating sample of IDs, one row per subject
- Using split / test / train
- Sensors and conditions separate, all trials used
- Level 1 and level 2 scores as predictors


condition 1, fp1

```{r cond1fp1}
# Set up data set of score for each trial for each person and overall person score
# 3 conditions x 3 fpca scores for each condition

# Include nodes fp1, fp2 (prefrontal), f3, f4, fz (frontal)

# all three conditions. one at a time and merge
nodes_en <- c("fp1") # pre-frontal and frontal nodes
 
sm_cond_en <- smooth_trial_data(df = p_df, cond_num = 1, node = nodes_en)[[2]]
cond_en <- smooth_trial_data(df = p_df, cond_num = 1, node = nodes_en)[[1]]

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                             time = cond_en$time_ms,
                             eeg_raw = as.numeric(sm_cond_en$y), #raw eeg signals
                             eeg_smooth = as.numeric(predict(sm_cond_en))) %>%   # smoothed predictions
                              arrange(ID, trial)

# mFPCA scores condition 1 fp1

# Now we perform multi-level modelling, where the first level is the individual subject level and the second level is the individual trial level. So we get a score for each person and for each curve (trial).

# df: Data in long format with columns, subject Id = Id, stride_id (wave_id) = visit, arg = time (seq(0, 1, length = 101), obs = y values
#Output: each row is a curve with subject Id and wave id

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)
n <- ncol(df_wide_en)
Y_en <- as.matrix(df_wide_en[, c(3:n)]) #not including trial or id columns

#mfpca

mlfpc_en <- mfpca.face(Y = Y_en, # matrix- each row is a curves
                    id = df_wide_en$ID, # subject id: c(1,1,2,2,2,3,3 ...)
                    visit = df_wide_en$trial, # trial id c(1,2,3,.., 1, 2,3, 4, ...)
                    twoway = TRUE,
                    pve = 0.95) # retain this % of the variability

myIDS = unique((dfresponse_en$ID))

#df of scores level1 & 2
level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )
level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

l1 <- ncol(as.data.frame(mlfpc_en$score$level1)) # number of fpcs
colnames(level1_en) <- c("ID", c(paste("cond1_", nodes_en, "_L1_", 1:l1, sep = "")))

l2 <- ncol(as.data.frame(mlfpc_en$score$level2)) 
colnames(level2_en) <- c("ID", "trial", c(paste("cond1_", nodes_en, "_L2_", 1:l2, sep = "")))

# merge level 1 and level 2 scores

fpca_fp1_en1 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_fp1_merged_en1 <- demo_train %>%
  left_join(fpca_fp1_en1, by = "ID")
```

```{r, predict scores cond1fp1}
node_test <- c("fp1")

sm_test <- smooth_trial_data(df = pdf_test, cond_num = 1, node = "fp1")[[2]]
cond_test <- smooth_trial_data(df = pdf_test, cond_num = 1, node = "fp1")[[1]]

dfresponse_test <- data.frame(ID = as.numeric(as.character(cond_test$subject)),
                              trial = as.numeric(as.character(cond_test$trial)),
                              time = cond_test$time_ms,
                              eeg_raw = as.numeric(sm_test$y),
                              eeg_smooth = as.numeric(predict(sm_test)),
                              schizophrenia = cond_test$schizophrenia) %>% 
  arrange(ID, trial)

df_wide_test <- pivot_wider(dfresponse_test[,-4], #exclude the raw eeg column
                       names_from = time,
                       values_from = eeg_smooth)
n <- ncol(df_wide_test)
Y_test <- as.matrix(df_wide_test %>% dplyr::select(-ID, -trial, -schizophrenia)) # excluding column id, trial, schizophrenia

myIDS_test = unique((dfresponse_test$ID))

predicted_scores <- predict_scores(Y_test, df_wide_test$ID, df_wide_test$trial, mlfpc_en)

#df of scores level1 & 2
level1_test <- as.data.frame(predicted_scores[[1]]) %>%
  mutate(ID = myIDS_test, .before =1 )
level2_test <- as.data.frame(predicted_scores[[2]]) %>%
  mutate(ID = df_wide_test$ID, trial = df_wide_test$trial, .before = 1 )

l1 <- ncol(as.data.frame(predicted_scores[[1]])) # number of fpcs
colnames(level1_test) <- c("ID", c(paste("cond1_", node_test, "_L1_", 1:l1, sep = "")))

l2 <- ncol(as.data.frame(predicted_scores[[2]])) 
colnames(level2_test) <- c("ID", "trial", c(paste("cond1_", node_test, "_L2_", 1:l2, sep = "")))

#put in a single df
pred_fp1_c1 <- level2_test %>%
  left_join(level1_test, by = "ID")

pred_fp1_c1_merged <- demo_test %>%
  left_join(pred_fp1_c1, by = "ID")
```

condition 2, fp1
```{r cond2fp1}
nodes_en <- c("fp1")
sm_cond_en <- smooth_trial_data(df = p_df, cond_num = 2, node = nodes_en)[[2]]
cond_en <- smooth_trial_data(df = p_df, cond_num = 2, node = nodes_en)[[1]]

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                             time = cond_en$time_ms,
                             eeg_raw = as.numeric(sm_cond_en$y),
                             eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                              arrange(ID, trial)

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)
n <- ncol(df_wide_en)
Y_en <- as.matrix(df_wide_en[, c(3:n)]) #not including trial or id columns

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )

myIDS = unique((dfresponse_en$ID))

#df of scores level1 & 2
level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )
level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

l1 <- ncol(as.data.frame(mlfpc_en$score$level1)) # number of fpcs
colnames(level1_en) <- c("ID", c(paste("cond2_", nodes_en, "_L1_", 1:l1, sep = "")))

l2 <- ncol(as.data.frame(mlfpc_en$score$level2)) 
colnames(level2_en) <- c("ID", "trial", c(paste("cond2_", nodes_en, "_L2_", 1:l2, sep = "")))

#add scores into dataframe
fpca_fp1_en2 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_fp1_merged_en2 <- fpca_fp1_merged_en1 %>%
  left_join(fpca_fp1_en2, by = c("ID", "trial")) # joining here by ID and trial as we are given "trial.x" and "trial.y"
```

```{r, predict scores cond2fp1}
node_test <- c("fp1")

sm_test <- smooth_trial_data(df = pdf_test, cond_num = 2, node = "fp1")[[2]]
cond_test <- smooth_trial_data(df = pdf_test, cond_num = 2, node = "fp1")[[1]]

dfresponse_test <- data.frame(ID = as.numeric(as.character(cond_test$subject)),
                              trial = as.numeric(as.character(cond_test$trial)),
                         time = cond_test$time_ms,
                         eeg_raw = as.numeric(sm_test$y),
                         eeg_smooth = as.numeric(predict(sm_test)),
                         schizophrenia = cond_test$schizophrenia) %>% 
                         arrange(ID, trial)

df_wide_test <- pivot_wider(dfresponse_test[,-4], #exclude the raw eeg column
                       names_from = time,
                       values_from = eeg_smooth)
n <- ncol(df_wide_test)
Y_test <- as.matrix(df_wide_test %>% dplyr::select(-ID, -trial, -schizophrenia)) # excluding column id, trial, schizophrenia

myIDS_test = unique((dfresponse_test$ID))
predicted_scores <- predict_scores(Y_test, df_wide_test$ID, df_wide_test$trial, mlfpc_en)

#df of scores level1 & 2
level1_test <- as.data.frame(predicted_scores[[1]]) %>%
  mutate(ID = myIDS_test, .before =1 )
level2_test <- as.data.frame(predicted_scores[[2]]) %>%
  mutate(ID = df_wide_test$ID, trial = df_wide_test$trial, .before = 1 )

l1 <- ncol(as.data.frame(predicted_scores[[1]])) # number of fpcs
colnames(level1_test) <- c("ID", c(paste("cond2_", node_test, "_L1_", 1:l1, sep = "")))

l2 <- ncol(as.data.frame(predicted_scores[[2]])) 
colnames(level2_test) <- c("ID", "trial", c(paste("cond2_", node_test, "_L2_", 1:l2, sep = "")))

#put in a single df
pred_fp1_c2 <- level2_test %>%
  left_join(level1_test, by = "ID")

pred_fp1_c2_merged <- pred_fp1_c1_merged %>%
  left_join(pred_fp1_c2, by = c("ID", "trial"))
```

condition 3, fp1

```{r cond3fp1}
nodes_en <- c("fp1")

sm_cond_en <- smooth_trial_data(df = p_df, cond_num = 3, node = nodes_en)[[2]]
cond_en <- smooth_trial_data(df = p_df, cond_num = 3, node = nodes_en)[[1]]

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                             time = cond_en$time_ms,
                             eeg_raw = as.numeric(sm_cond_en$y),
                             eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                              arrange(ID, trial)

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)
n <- ncol(df_wide_en)
Y_en <- as.matrix(df_wide_en[, c(3:n)]) #not including trial or id columns

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )
myIDS = unique((dfresponse_en$ID))

#df of scores level1 & 2
level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )
level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

l1 <- ncol(as.data.frame(mlfpc_en$score$level1)) # number of fpcs
colnames(level1_en) <- c("ID", c(paste("cond3_", nodes_en, "_L1_", 1:l1, sep = "")))

l2 <- ncol(as.data.frame(mlfpc_en$score$level2)) 
colnames(level2_en) <- c("ID", "trial", c(paste("cond3_", nodes_en, "_L2_", 1:l2, sep = "")))

#add to dataframe
fpca_fp1_en3 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_fp1_merged_en3 <- fpca_fp1_merged_en2 %>%
  left_join(fpca_fp1_en3, by = c("ID", "trial"))
```


```{r, predict scores cond3fp1}
node_test <- c("fp1")

sm_test <- smooth_trial_data(df = pdf_test, cond_num = 3, node = "fp1")[[2]]
cond_test <- smooth_trial_data(df = pdf_test, cond_num = 3, node = "fp1")[[1]]

dfresponse_test <- data.frame(ID = as.numeric(as.character(cond_test$subject)),
                              trial = as.numeric(as.character(cond_test$trial)),
                         time = cond_test$time_ms,
                         eeg_raw = as.numeric(sm_test$y),
                         eeg_smooth = as.numeric(predict(sm_test)),
                         schizophrenia = cond_test$schizophrenia) %>% 
                         arrange(ID, trial)

df_wide_test <- pivot_wider(dfresponse_test[,-4], #exclude the raw eeg column
                       names_from = time,
                       values_from = eeg_smooth)
n <- ncol(df_wide_test)
Y_test <- as.matrix(df_wide_test %>% dplyr::select(-ID, -trial, -schizophrenia)) # excluding column id, trial, schizophrenia

myIDS_test = unique((dfresponse_test$ID))
predicted_scores <- predict_scores(Y_test, df_wide_test$ID, df_wide_test$trial, mlfpc_en)

#df of scores level1 & 2
level1_test <- as.data.frame(predicted_scores[[1]]) %>%
  mutate(ID = myIDS_test, .before =1 )
level2_test <- as.data.frame(predicted_scores[[2]]) %>%
  mutate(ID = df_wide_test$ID, trial = df_wide_test$trial, .before = 1 )

l1 <- ncol(as.data.frame(predicted_scores[[1]])) # number of fpcs
colnames(level1_test) <- c("ID", c(paste("cond3_", node_test, "_L1_", 1:l1, sep = "")))

l2 <- ncol(as.data.frame(predicted_scores[[2]])) 
colnames(level2_test) <- c("ID", "trial", c(paste("cond3_", node_test, "_L2_", 1:l2, sep = "")))

#put in a single df
pred_fp1_c3 <- level2_test %>%
  left_join(level1_test, by = "ID")

pred_fp1_c3_merged <- pred_fp1_c2_merged %>%
  left_join(pred_fp1_c3, by = c("ID", "trial"))
```

-- f3 --

condition 1, f3

```{r cond1f3}
nodes_en <- c("f3")

sm_cond_en <- smooth_trial_data(df = p_df, cond_num = 1, node = nodes_en)[[2]]
cond_en <- smooth_trial_data(df = p_df, cond_num = 1, node = nodes_en)[[1]]

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                             time = cond_en$time_ms,
                             eeg_raw = as.numeric(sm_cond_en$y),
                             eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                              arrange(ID, trial)

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)
n <- ncol(df_wide_en)
Y_en <- as.matrix(df_wide_en[, c(3:n)]) #not including trial or id columns

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )

myIDS = unique((dfresponse_en$ID))

#df of scores level1 & 2
level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )
level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

l1 <- ncol(as.data.frame(mlfpc_en$score$level1)) # number of fpcs
colnames(level1_en) <- c("ID", c(paste("cond1_", nodes_en, "_L1_", 1:l1, sep = "")))

l2 <- ncol(as.data.frame(mlfpc_en$score$level2)) 
colnames(level2_en) <- c("ID", "trial", c(paste("cond1_", nodes_en, "_L2_", 1:l2, sep = "")))

#add scores to data
fpca_f3_en1 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_f3_merged_en1 <- fpca_fp1_merged_en3 %>%
  left_join(fpca_f3_en1, by = c("ID", "trial"))

```

```{r, predict scores cond1f3}
node_test <- c("f3")

sm_test <- smooth_trial_data(df = pdf_test, cond_num = 1, node = "f3")[[2]]
cond_test <- smooth_trial_data(df = pdf_test, cond_num = 1, node = "f3")[[1]]

dfresponse_test <- data.frame(ID = as.numeric(as.character(cond_test$subject)),
                              trial = as.numeric(as.character(cond_test$trial)),
                         time = cond_test$time_ms,
                         eeg_raw = as.numeric(sm_test$y),
                         eeg_smooth = as.numeric(predict(sm_test)),
                         schizophrenia = cond_test$schizophrenia) %>% 
                         arrange(ID, trial)

df_wide_test <- pivot_wider(dfresponse_test[,-4], #exclude the raw eeg column
                       names_from = time,
                       values_from = eeg_smooth)
n <- ncol(df_wide_test)
Y_test <- as.matrix(df_wide_test %>% dplyr::select(-ID, -trial, -schizophrenia)) # excluding column id, trial, schizophrenia

myIDS_test = unique((dfresponse_test$ID))
predicted_scores <- predict_scores(Y_test, df_wide_test$ID, df_wide_test$trial, mlfpc_en)

#df of scores level1 & 2
level1_test <- as.data.frame(predicted_scores[[1]]) %>%
  mutate(ID = myIDS_test, .before =1 )
level2_test <- as.data.frame(predicted_scores[[2]]) %>%
  mutate(ID = df_wide_test$ID, trial = df_wide_test$trial, .before = 1 )

l1 <- ncol(as.data.frame(predicted_scores[[1]])) # number of fpcs
colnames(level1_test) <- c("ID", c(paste("cond1_", node_test, "_L1_", 1:l1, sep = "")))

l2 <- ncol(as.data.frame(predicted_scores[[2]])) 
colnames(level2_test) <- c("ID", "trial", c(paste("cond1_", node_test, "_L2_", 1:l2, sep = "")))

#put in a single df
pred_f3_c1 <- level2_test %>%
  left_join(level1_test, by = "ID")

pred_f3_c1_merged <- pred_fp1_c3_merged %>%
  left_join(pred_f3_c1, by = c("ID", "trial"))
```


condition 2, f3

```{r cond2f3}
nodes_en <- c("f3")

sm_cond_en <- smooth_trial_data(df = p_df, cond_num = 2, node = nodes_en)[[2]]
cond_en <- smooth_trial_data(df = p_df, cond_num = 2, node = nodes_en)[[1]]

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                             time = cond_en$time_ms,
                             eeg_raw = as.numeric(sm_cond_en$y),
                             eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                              arrange(ID, trial)

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)
n <- ncol(df_wide_en)
Y_en <- as.matrix(df_wide_en[, c(3:n)]) #not including trial or id columns

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )

myIDS = unique((dfresponse_en$ID))

#df of scores level1 & 2
level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )
level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

l1 <- ncol(as.data.frame(mlfpc_en$score$level1)) # number of fpcs
colnames(level1_en) <- c("ID", c(paste("cond2_", nodes_en, "_L1_", 1:l1, sep = "")))

l2 <- ncol(as.data.frame(mlfpc_en$score$level2)) 
colnames(level2_en) <- c("ID", "trial", c(paste("cond2_", nodes_en, "_L2_", 1:l2, sep = "")))

#add scores to dataframe
fpca_f3_en2 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_f3_merged_en2 <- fpca_f3_merged_en1 %>%
  left_join(fpca_f3_en2, by = c("ID", "trial"))
```

```{r, predict scores cond2f3}
node_test <- c("f3")

sm_test <- smooth_trial_data(df = pdf_test, cond_num = 2, node = "f3")[[2]]
cond_test <- smooth_trial_data(df = pdf_test, cond_num = 2, node = "f3")[[1]]

dfresponse_test <- data.frame(ID = as.numeric(as.character(cond_test$subject)),
                              trial = as.numeric(as.character(cond_test$trial)),
                         time = cond_test$time_ms,
                         eeg_raw = as.numeric(sm_test$y),
                         eeg_smooth = as.numeric(predict(sm_test)),
                         schizophrenia = cond_test$schizophrenia) %>% 
                         arrange(ID, trial)

df_wide_test <- pivot_wider(dfresponse_test[,-4], #exclude the raw eeg column
                       names_from = time,
                       values_from = eeg_smooth)
n <- ncol(df_wide_test)
Y_test <- as.matrix(df_wide_test %>% dplyr::select(-ID, -trial, -schizophrenia)) # excluding column id, trial, schizophrenia

myIDS_test = unique((dfresponse_test$ID))
predicted_scores <- predict_scores(Y_test, df_wide_test$ID, df_wide_test$trial, mlfpc_en)

#df of scores level1 & 2
level1_test <- as.data.frame(predicted_scores[[1]]) %>%
  mutate(ID = myIDS_test, .before =1 )
level2_test <- as.data.frame(predicted_scores[[2]]) %>%
  mutate(ID = df_wide_test$ID, trial = df_wide_test$trial, .before = 1 )

l1 <- ncol(as.data.frame(predicted_scores[[1]])) # number of fpcs
colnames(level1_test) <- c("ID", c(paste("cond2_", node_test, "_L1_", 1:l1, sep = "")))

l2 <- ncol(as.data.frame(predicted_scores[[2]])) 
colnames(level2_test) <- c("ID", "trial", c(paste("cond2_", node_test, "_L2_", 1:l2, sep = "")))

#put in a single df
pred_f3_c2 <- level2_test %>%
  left_join(level1_test, by = "ID")

pred_f3_c2_merged <- pred_f3_c1_merged %>%
  left_join(pred_f3_c2, by = c("ID", "trial"))
```

condition3, f3

```{r cond3f3}
nodes_en <- c("f3")

sm_cond_en <- smooth_trial_data(df = p_df, cond_num = 3, node = nodes_en)[[2]]
cond_en <- smooth_trial_data(df = p_df, cond_num = 3, node = nodes_en)[[1]]

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                             time = cond_en$time_ms,
                             eeg_raw = as.numeric(sm_cond_en$y),
                             eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                              arrange(ID, trial)

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)
n <- ncol(df_wide_en)
Y_en <- as.matrix(df_wide_en[, c(3:n)]) #not including trial or id columns

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )

myIDS = unique((dfresponse_en$ID))

#df of scores level1 & 2
level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )
level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

l1 <- ncol(as.data.frame(mlfpc_en$score$level1)) # number of fpcs
colnames(level1_en) <- c("ID", c(paste("cond3_", nodes_en, "_L1_", 1:l1, sep = "")))

l2 <- ncol(as.data.frame(mlfpc_en$score$level2)) 
colnames(level2_en) <- c("ID", "trial", c(paste("cond3_", nodes_en, "_L2_", 1:l2, sep = "")))

#add scores to dataframe
fpca_f3_en3 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_f3_merged_en3 <- fpca_f3_merged_en2 %>%
  left_join(fpca_f3_en3, by = c("ID", "trial"))

```


```{r, predict scores cond3f3}
node_test <- c("f3")

sm_test <- smooth_trial_data(df = pdf_test, cond_num = 3, node = "f3")[[2]]
cond_test <- smooth_trial_data(df = pdf_test, cond_num = 3, node = "f3")[[1]]

dfresponse_test <- data.frame(ID = as.numeric(as.character(cond_test$subject)),
                              trial = as.numeric(as.character(cond_test$trial)),
                         time = cond_test$time_ms,
                         eeg_raw = as.numeric(sm_test$y),
                         eeg_smooth = as.numeric(predict(sm_test)),
                         schizophrenia = cond_test$schizophrenia) %>% 
                         arrange(ID, trial)

df_wide_test <- pivot_wider(dfresponse_test[,-4], #exclude the raw eeg column
                       names_from = time,
                       values_from = eeg_smooth)
n <- ncol(df_wide_test)
Y_test <- as.matrix(df_wide_test %>% dplyr::select(-ID, -trial, -schizophrenia)) # excluding column id, trial, schizophrenia

myIDS_test = unique((dfresponse_test$ID))
predicted_scores <- predict_scores(Y_test, df_wide_test$ID, df_wide_test$trial, mlfpc_en)

#df of scores level1 & 2
level1_test <- as.data.frame(predicted_scores[[1]]) %>%
  mutate(ID = myIDS_test, .before =1 )
level2_test <- as.data.frame(predicted_scores[[2]]) %>%
  mutate(ID = df_wide_test$ID, trial = df_wide_test$trial, .before = 1 )

l1 <- ncol(as.data.frame(predicted_scores[[1]])) # number of fpcs
colnames(level1_test) <- c("ID", c(paste("cond3_", node_test, "_L1_", 1:l1, sep = "")))

l2 <- ncol(as.data.frame(predicted_scores[[2]])) 
colnames(level2_test) <- c("ID", "trial", c(paste("cond3_", node_test, "_L2_", 1:l2, sep = "")))

#put in a single df
pred_f3_c3 <- level2_test %>%
  left_join(level1_test, by = "ID")

pred_f3_c3_merged <- pred_f3_c2_merged %>%
  left_join(pred_f3_c3, by = c("ID", "trial"))
```

-- f7 --


condition 1, f7

```{r cond1f7}
nodes_en <- c("f7")

sm_cond_en <- smooth_trial_data(df = p_df, cond_num = 1, node = nodes_en)[[2]]
cond_en <- smooth_trial_data(df = p_df, cond_num = 1, node = nodes_en)[[1]]

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                             time = cond_en$time_ms,
                             eeg_raw = as.numeric(sm_cond_en$y),
                             eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                              arrange(ID, trial)

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)
n <- ncol(df_wide_en)
Y_en <- as.matrix(df_wide_en[, c(3:n)]) #not including trial or id columns

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )
myIDS = unique((dfresponse_en$ID))

#df of scores level1 & 2
level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )
level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

l1 <- ncol(as.data.frame(mlfpc_en$score$level1)) # number of fpcs
colnames(level1_en) <- c("ID", c(paste("cond1_", nodes_en, "_L1_", 1:l1, sep = "")))

l2 <- ncol(as.data.frame(mlfpc_en$score$level2)) 
colnames(level2_en) <- c("ID", "trial", c(paste("cond1_", nodes_en, "_L2_", 1:l2, sep = "")))


#add scores to dataframe
fpca_f7_en1 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_f7_merged_en1 <- fpca_f3_merged_en3 %>%
  left_join(fpca_f7_en1, by = c("ID", "trial"))

```


```{r, predict scores cond1f7}
node_test <- c("f7")

sm_test <- smooth_trial_data(df = pdf_test, cond_num = 1, node = "f7")[[2]]
cond_test <- smooth_trial_data(df = pdf_test, cond_num = 1, node = "f7")[[1]]

dfresponse_test <- data.frame(ID = as.numeric(as.character(cond_test$subject)),
                              trial = as.numeric(as.character(cond_test$trial)),
                         time = cond_test$time_ms,
                         eeg_raw = as.numeric(sm_test$y),
                         eeg_smooth = as.numeric(predict(sm_test)),
                         schizophrenia = cond_test$schizophrenia) %>% 
                         arrange(ID, trial)

df_wide_test <- pivot_wider(dfresponse_test[,-4], #exclude the raw eeg column
                       names_from = time,
                       values_from = eeg_smooth)
n <- ncol(df_wide_test)
Y_test <- as.matrix(df_wide_test %>% dplyr::select(-ID, -trial, -schizophrenia)) # excluding column id, trial, schizophrenia

myIDS_test = unique((dfresponse_test$ID))
predicted_scores <- predict_scores(Y_test, df_wide_test$ID, df_wide_test$trial, mlfpc_en)

#df of scores level1 & 2
level1_test <- as.data.frame(predicted_scores[[1]]) %>%
  mutate(ID = myIDS_test, .before =1 )
level2_test <- as.data.frame(predicted_scores[[2]]) %>%
  mutate(ID = df_wide_test$ID, trial = df_wide_test$trial, .before = 1 )

l1 <- ncol(as.data.frame(predicted_scores[[1]])) # number of fpcs
colnames(level1_test) <- c("ID", c(paste("cond1_", node_test, "_L1_", 1:l1, sep = "")))
l2 <- ncol(as.data.frame(predicted_scores[[2]])) 
colnames(level2_test) <- c("ID", "trial", c(paste("cond1_", node_test, "_L2_", 1:l2, sep = "")))

#put in a single df
pred_f7_c1 <- level2_test %>%
  left_join(level1_test, by = "ID")

pred_f7_c1_merged <- pred_f3_c3_merged %>%
  left_join(pred_f7_c1, by = c("ID", "trial"))
```

condition 2, f7


```{r cond2f7}

nodes_en <- c("f7")

sm_cond_en <- smooth_trial_data(df = p_df, cond_num = 2, node = nodes_en)[[2]]
cond_en <- smooth_trial_data(df = p_df, cond_num = 2, node = nodes_en)[[1]]

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                             time = cond_en$time_ms,
                             eeg_raw = as.numeric(sm_cond_en$y),
                             eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                              arrange(ID, trial)

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)
n <- ncol(df_wide_en)
Y_en <- as.matrix(df_wide_en[, c(3:n)]) #not including trial or id columns

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )
myIDS = unique((dfresponse_en$ID))

#df of scores level1 & 2
level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )
level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

l1 <- ncol(as.data.frame(mlfpc_en$score$level1)) # number of fpcs
colnames(level1_en) <- c("ID", c(paste("cond2_", nodes_en, "_L1_", 1:l1, sep = "")))

l2 <- ncol(as.data.frame(mlfpc_en$score$level2)) 
colnames(level2_en) <- c("ID", "trial", c(paste("cond2_", nodes_en, "_L2_", 1:l2, sep = "")))

#add new scores to df
fpca_f7_en2 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_f7_merged_en2 <- fpca_f7_merged_en1 %>%
  left_join(fpca_f7_en2, by = c("ID", "trial"))


```


```{r, predict scores cond2f7}
node_test <- c("f7")

sm_test <- smooth_trial_data(df = pdf_test, cond_num = 2, node = "f7")[[2]]
cond_test <- smooth_trial_data(df = pdf_test, cond_num = 2, node = "f7")[[1]]

dfresponse_test <- data.frame(ID = as.numeric(as.character(cond_test$subject)),
                              trial = as.numeric(as.character(cond_test$trial)),
                         time = cond_test$time_ms,
                         eeg_raw = as.numeric(sm_test$y),
                         eeg_smooth = as.numeric(predict(sm_test)),
                         schizophrenia = cond_test$schizophrenia) %>% 
                         arrange(ID, trial)

df_wide_test <- pivot_wider(dfresponse_test[,-4], #exclude the raw eeg column
                       names_from = time,
                       values_from = eeg_smooth)
n <- ncol(df_wide_test)
Y_test <- as.matrix(df_wide_test %>% dplyr::select(-ID, -trial, -schizophrenia)) # excluding column id, trial, schizophrenia

myIDS_test = unique((dfresponse_test$ID))
predicted_scores <- predict_scores(Y_test, df_wide_test$ID, df_wide_test$trial, mlfpc_en)

#df of scores level1 & 2
level1_test <- as.data.frame(predicted_scores[[1]]) %>%
  mutate(ID = myIDS_test, .before =1 )
level2_test <- as.data.frame(predicted_scores[[2]]) %>%
  mutate(ID = df_wide_test$ID, trial = df_wide_test$trial, .before = 1 )

l1 <- ncol(as.data.frame(predicted_scores[[1]])) # number of fpcs
colnames(level1_test) <- c("ID", c(paste("cond2_", node_test, "_L1_", 1:l1, sep = "")))
l2 <- ncol(as.data.frame(predicted_scores[[2]])) 
colnames(level2_test) <- c("ID", "trial", c(paste("cond2_", node_test, "_L2_", 1:l2, sep = "")))

#put in a single df
pred_f7_c2 <- level2_test %>%
  left_join(level1_test, by = "ID")

pred_f7_c2_merged <- pred_f7_c1_merged %>%
  left_join(pred_f7_c2, by = c("ID", "trial"))
```

condition 3, f7

```{r cond3f7}
nodes_en <- c("f7")

sm_cond_en <- smooth_trial_data(df = p_df, cond_num = 3, node = nodes_en)[[2]]
cond_en <- smooth_trial_data(df = p_df, cond_num = 3, node = nodes_en)[[1]]

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                         time = cond_en$time_ms,
                         eeg_raw = as.numeric(sm_cond_en$y),
                         eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                          arrange(ID, trial)


df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)
Y_en <- as.matrix(df_wide_en[, -(1:2)]) # remove trial and id columns

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )
myIDS = unique((dfresponse_en$ID))

#df of scores level1 & 2
level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )
level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

l1 <- ncol(as.data.frame(mlfpc_en$score$level1)) # number of fpcs
colnames(level1_en) <- c("ID", c(paste("cond3_", nodes_en, "_L1_", 1:l1, sep = "")))

l2 <- ncol(as.data.frame(mlfpc_en$score$level2)) 
colnames(level2_en) <- c("ID", "trial", c(paste("cond3_", nodes_en, "_L2_", 1:l2, sep = "")))

# add to full scores df
fpca_f7_en3 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_f7_merged_en3 <- fpca_f7_merged_en2 %>%
  left_join(fpca_f7_en3, by = c("ID", "trial"))


```

```{r, predict scores cond3f7}
node_test <- c("f7")

sm_test <- smooth_trial_data(df = pdf_test, cond_num = 3, node = "f7")[[2]]
cond_test <- smooth_trial_data(df = pdf_test, cond_num = 3, node = "f7")[[1]]

dfresponse_test <- data.frame(ID = as.numeric(as.character(cond_test$subject)),
                              trial = as.numeric(as.character(cond_test$trial)),
                         time = cond_test$time_ms,
                         eeg_raw = as.numeric(sm_test$y),
                         eeg_smooth = as.numeric(predict(sm_test)),
                         schizophrenia = cond_test$schizophrenia) %>% 
                         arrange(ID, trial)

df_wide_test <- pivot_wider(dfresponse_test[,-4], #exclude the raw eeg column
                       names_from = time,
                       values_from = eeg_smooth)
n <- ncol(df_wide_test)
Y_test <- as.matrix(df_wide_test %>% dplyr::select(-ID, -trial, -schizophrenia)) # excluding column id, trial, schizophrenia

myIDS_test = unique((dfresponse_test$ID))
predicted_scores <- predict_scores(Y_test, df_wide_test$ID, df_wide_test$trial, mlfpc_en)

#df of scores level1 & 2
level1_test <- as.data.frame(predicted_scores[[1]]) %>%
  mutate(ID = myIDS_test, .before =1 )
level2_test <- as.data.frame(predicted_scores[[2]]) %>%
  mutate(ID = df_wide_test$ID, trial = df_wide_test$trial, .before = 1 )

l1 <- ncol(as.data.frame(predicted_scores[[1]])) # number of fpcs
colnames(level1_test) <- c("ID", c(paste("cond3_", node_test, "_L1_", 1:l1, sep = "")))
l2 <- ncol(as.data.frame(predicted_scores[[2]])) 
colnames(level2_test) <- c("ID", "trial", c(paste("cond3_", node_test, "_L2_", 1:l2, sep = "")))

#put in a single df
pred_f7_c3 <- level2_test %>%
  left_join(level1_test, by = "ID")

pred_f7_c3_merged <- pred_f7_c2_merged %>%
  left_join(pred_f7_c3, by = c("ID", "trial"))
```


-- af7 --

condition 1, af7

```{r cond1af7}
nodes_en <- c("af7")

sm_cond_en <- smooth_trial_data(df = p_df, cond_num = 1, node = nodes_en)[[2]]
cond_en <- smooth_trial_data(df = p_df, cond_num = 1, node = nodes_en)[[1]]

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                             time = cond_en$time_ms,
                             eeg_raw = as.numeric(sm_cond_en$y),
                             eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                              arrange(ID, trial)

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)
Y_en <- as.matrix(df_wide_en[, -(1:2)]) # remove id & trial 

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )
myIDS = unique((dfresponse_en$ID))

#df of scores level1 & 2
level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )
level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

l1 <- ncol(as.data.frame(mlfpc_en$score$level1)) # number of fpcs
colnames(level1_en) <- c("ID", c(paste("cond1_", nodes_en, "_L1_", 1:l1, sep = "")))

l2 <- ncol(as.data.frame(mlfpc_en$score$level2)) 
colnames(level2_en) <- c("ID", "trial", c(paste("cond1_", nodes_en, "_L2_", 1:l2, sep = "")))

# add scores to full scores df
fpca_af7_en1 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_af7_merged_en1 <- fpca_f7_merged_en3 %>%
  left_join(fpca_af7_en1, by = c("ID", "trial"))
```

```{r, predict scores cond1af7}
node_test <- c("af7")

sm_test <- smooth_trial_data(df = pdf_test, cond_num = 1, node = "af7")[[2]]
cond_test <- smooth_trial_data(df = pdf_test, cond_num = 1, node = "af7")[[1]]

dfresponse_test <- data.frame(ID = as.numeric(as.character(cond_test$subject)),
                              trial = as.numeric(as.character(cond_test$trial)),
                         time = cond_test$time_ms,
                         eeg_raw = as.numeric(sm_test$y),
                         eeg_smooth = as.numeric(predict(sm_test)),
                         schizophrenia = cond_test$schizophrenia) %>% 
                         arrange(ID, trial)

df_wide_test <- pivot_wider(dfresponse_test[,-4], #exclude the raw eeg column
                       names_from = time,
                       values_from = eeg_smooth)
n <- ncol(df_wide_test)
Y_test <- as.matrix(df_wide_test %>% dplyr::select(-ID, -trial, -schizophrenia)) # excluding column id, trial, schizophrenia

myIDS_test = unique((dfresponse_test$ID))
predicted_scores <- predict_scores(Y_test, df_wide_test$ID, df_wide_test$trial, mlfpc_en)

#df of scores level1 & 2
level1_test <- as.data.frame(predicted_scores[[1]]) %>%
  mutate(ID = myIDS_test, .before =1 )
level2_test <- as.data.frame(predicted_scores[[2]]) %>%
  mutate(ID = df_wide_test$ID, trial = df_wide_test$trial, .before = 1 )

l1 <- ncol(as.data.frame(predicted_scores[[1]])) # number of fpcs
colnames(level1_test) <- c("ID", c(paste("cond1_", node_test, "_L1_", 1:l1, sep = "")))
l2 <- ncol(as.data.frame(predicted_scores[[2]])) 
colnames(level2_test) <- c("ID", "trial", c(paste("cond1_", node_test, "_L2_", 1:l2, sep = "")))

#put in a single df
pred_af7_c1 <- level2_test %>%
  left_join(level1_test, by = "ID")

pred_af7_c1_merged <- pred_f7_c3_merged %>%
  left_join(pred_af7_c1, by = c("ID", "trial"))
```


condition 2, af7

```{r cond2af7}
nodes_en <- c("af7")

sm_cond_en <- smooth_trial_data(df = p_df, cond_num = 2, node = nodes_en)[[2]]
cond_en <- smooth_trial_data(df = p_df, cond_num = 2, node = nodes_en)[[1]]

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                         time = cond_en$time_ms,
                         eeg_raw = as.numeric(sm_cond_en$y),
                         eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                          arrange(ID, trial)


df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)
Y_en <- as.matrix(df_wide_en[, -(1:2)]) # remove id & trial columns

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )
myIDS = unique((dfresponse_en$ID))

#df of scores level1 & 2
level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )
level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

l1 <- ncol(as.data.frame(mlfpc_en$score$level1)) # number of fpcs
colnames(level1_en) <- c("ID", c(paste("cond2_", nodes_en, "L1_", 1:l1, sep = "")))

l2 <- ncol(as.data.frame(mlfpc_en$score$level2)) 
colnames(level2_en) <- c("ID", "trial", c(paste("cond2_", nodes_en, "L2_", 1:l2, sep = "")))

# add scores to df
fpca_af7_en2 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_af7_merged_en2 <- fpca_af7_merged_en1 %>%
  left_join(fpca_af7_en2, by = c("ID", "trial"))

```


```{r, predict scores cond2af7}
node_test <- c("af7")

sm_test <- smooth_trial_data(df = pdf_test, cond_num = 2, node = "af7")[[2]]
cond_test <- smooth_trial_data(df = pdf_test, cond_num = 2, node = "af7")[[1]]

dfresponse_test <- data.frame(ID = as.numeric(as.character(cond_test$subject)),
                              trial = as.numeric(as.character(cond_test$trial)),
                         time = cond_test$time_ms,
                         eeg_raw = as.numeric(sm_test$y),
                         eeg_smooth = as.numeric(predict(sm_test)),
                         schizophrenia = cond_test$schizophrenia) %>% 
                         arrange(ID, trial)

df_wide_test <- pivot_wider(dfresponse_test[,-4], #exclude the raw eeg column
                       names_from = time,
                       values_from = eeg_smooth)
n <- ncol(df_wide_test)
Y_test <- as.matrix(df_wide_test %>% dplyr::select(-ID, -trial, -schizophrenia)) # excluding column id, trial, schizophrenia

myIDS_test = unique((dfresponse_test$ID))
predicted_scores <- predict_scores(Y_test, df_wide_test$ID, df_wide_test$trial, mlfpc_en)

#df of scores level1 & 2
level1_test <- as.data.frame(predicted_scores[[1]]) %>%
  mutate(ID = myIDS_test, .before =1 )
level2_test <- as.data.frame(predicted_scores[[2]]) %>%
  mutate(ID = df_wide_test$ID, trial = df_wide_test$trial, .before = 1 )

l1 <- ncol(as.data.frame(predicted_scores[[1]])) # number of fpcs
colnames(level1_test) <- c("ID", c(paste("cond2_", node_test, "_L1_", 1:l1, sep = "")))
l2 <- ncol(as.data.frame(predicted_scores[[2]])) 
colnames(level2_test) <- c("ID", "trial", c(paste("cond2_", node_test, "_L2_", 1:l2, sep = "")))

#put in a single df
pred_af7_c2 <- level2_test %>%
  left_join(level1_test, by = "ID")

pred_af7_c2_merged <- pred_af7_c1_merged %>%
  left_join(pred_af7_c2, by = c("ID", "trial"))
```

condition 3, af7

```{r cond3af7}
nodes_en <- c("af7")
sm_cond_en <- smooth_trial_data(df = p_df, cond_num = 3, node = nodes_en)[[2]]
cond_en <- smooth_trial_data(df = p_df, cond_num = 3, node = nodes_en)[[1]]

dfresponse_en <- data.frame(ID = as.numeric(as.character(cond_en$subject)),
                             trial = as.numeric(as.character(cond_en$trial)),
                             time = cond_en$time_ms,
                             eeg_raw = as.numeric(sm_cond_en$y),
                             eeg_smooth = as.numeric(predict(sm_cond_en))) %>%
                              arrange(ID, trial)

df_wide_en <- pivot_wider(dfresponse_en[,-4], names_from = time, values_from = eeg_smooth)
Y_en <- as.matrix(df_wide_en[, -(1:2)])

mlfpc_en <- mfpca.face(Y = Y_en,
                    id = df_wide_en$ID,
                    visit = df_wide_en$trial,
                    twoway = TRUE,
                    pve = 0.95
                    )
myIDS = unique((dfresponse_en$ID))

#df of scores level1 & 2
level1_en <- as.data.frame(mlfpc_en$score$level1) %>%
  mutate(ID = myIDS, .before =1 )
level2_en <- as.data.frame(mlfpc_en$score$level2) %>%
  mutate(ID = df_wide_en$ID, trial = df_wide_en$trial, .before = 1 )

l1 <- ncol(as.data.frame(mlfpc_en$score$level1)) # number of fpcs
colnames(level1_en) <- c("ID", c(paste("cond3_", nodes_en, "_L1_", 1:l1, sep = "")))

l2 <- ncol(as.data.frame(mlfpc_en$score$level2)) 
colnames(level2_en) <- c("ID", "trial", c(paste("cond3_", nodes_en, "_L2_", 1:l2, sep = "")))

# add to full score df
fpca_af7_en3 <- level2_en %>%
  left_join(level1_en, by = "ID")

fpca_af7_merged_en3 <- fpca_af7_merged_en2 %>%
  left_join(fpca_af7_en3, by = c("ID", "trial"))

mfpca_matrix <- fpca_af7_merged_en3

```

```{r, predict scores cond3af7}
node_test <- c("af7")

sm_test <- smooth_trial_data(df = pdf_test, cond_num = 3, node = "af7")[[2]]
cond_test <- smooth_trial_data(df = pdf_test, cond_num = 3, node = "af7")[[1]]

dfresponse_test <- data.frame(ID = as.numeric(as.character(cond_test$subject)),
                              trial = as.numeric(as.character(cond_test$trial)),
                         time = cond_test$time_ms,
                         eeg_raw = as.numeric(sm_test$y),
                         eeg_smooth = as.numeric(predict(sm_test)),
                         schizophrenia = cond_test$schizophrenia) %>% 
                         arrange(ID, trial)

df_wide_test <- pivot_wider(dfresponse_test[,-4], #exclude the raw eeg column
                       names_from = time,
                       values_from = eeg_smooth)
n <- ncol(df_wide_test)
Y_test <- as.matrix(df_wide_test %>% dplyr::select(-ID, -trial, -schizophrenia)) # excluding column id, trial, schizophrenia

myIDS_test = unique((dfresponse_test$ID))
predicted_scores <- predict_scores(Y_test, df_wide_test$ID, df_wide_test$trial, mlfpc_en)

#df of scores level1 & 2
level1_test <- as.data.frame(predicted_scores[[1]]) %>%
  mutate(ID = myIDS_test, .before =1 )
level2_test <- as.data.frame(predicted_scores[[2]]) %>%
  mutate(ID = df_wide_test$ID, trial = df_wide_test$trial, .before = 1 )

l1 <- ncol(as.data.frame(predicted_scores[[1]])) # number of fpcs
colnames(level1_test) <- c("ID", c(paste("cond3_", node_test, "_L1_", 1:l1, sep = "")))
l2 <- ncol(as.data.frame(predicted_scores[[2]])) 
colnames(level2_test) <- c("ID", "trial", c(paste("cond3_", node_test, "_L2_", 1:l2, sep = "")))

#put in a single df
pred_af7_c3 <- level2_test %>%
  left_join(level1_test, by = "ID")

pred_af7_c3_merged <- pred_af7_c2_merged %>%
  left_join(pred_af7_c3, by = c("ID", "trial"))

mfpca_pred_scores <- pred_af7_c3_merged
```


## Variable Selection - Elastic Net 

Set up dataset to have three fpca scores per condition per node per person.
eg cond1_fp1_fpca_1, cond2_fp1_fpca_1 etc


### Apply Elastic Net 

All 3 conditions

```{r}
set.seed(0)

mfpca_allcond <- mfpca_matrix %>%
  mutate(group = ifelse(group == "Case", 1, 0),
         gender = ifelse(gender == "Female", 1, 0)) %>%
  drop_na() # TEMPORARY

mfpca_test_all_cond <- mfpca_pred_scores %>%
  mutate(group = ifelse(group == "Case", 1, 0),
         gender = ifelse(gender == "Female", 1, 0)) %>%
  drop_na() # TEMPORARY

# 10 fold CV in training to choose lambda (default). To be explicit: cv.glmnet(nfolds=10)
# glmnet() x:matrix. y: vector
##this is train data
x_train <- model.matrix(group ~ .-ID - trial, data = mfpca_allcond)[, -1] # gets rid of intercept term
y_train <- mfpca_allcond$group

x_test <- model.matrix(group ~ .-ID - trial, data = mfpca_test_all_cond)[, -1] # gets rid of intercept term
y_test <- mfpca_test_all_cond$group

# elastic net: 0 < alpha < 1
# 0 ridge, 1 lasso
# cv to find best value of lambda and alpha 
alpha_grid <- seq(0.01, 0.99, by = 0.1)

cv_list <- lapply(alpha_grid, function(a) {
  cv.glmnet(x_train, y_train, 
            alpha = a, 
            family = "binomial")})

# extract CV error for eachg alpha 
cv_error <- sapply(cv_list, function(m) min(m$cvm))

# best model
best_index <- which.min(cv_error)
best_cv <- cv_list[[best_index]]
best_alpha <- alpha_grid[best_index]

# best lambda corresponding to that alpha
bestlam <- best_cv$lambda.min
selam <- best_cv$lambda.1se

# fit elastic net model
enet <- glmnet(x_train, y_train,
               alpha = best_alpha,
               family = "binomial")      # standardize = TRUE is the default 

# predicted probabilities with best lam from test set, type = response for pred prob (not log odds)
enet_prob <- predict(enet, s = bestlam, newx = x_test, type = "response")
# 1se lam
enet_prob_se <- predict(enet, s = selam, newx = x_test, type = "response")

# convert to vector
enet_prob_v <- as.vector(enet_prob)
enet_prob_se_v <- as.vector(enet_prob_se)

df_pred_prob <- data.frame(enet_prob_v) %>% mutate(ID = mfpca_test_all_cond$ID, .before = 1) %>% mutate(schizophrenia = as.factor(mfpca_test_all_cond$group)) %>% mutate(trial = mfpca_test_all_cond$trial)

df_pred_prob %>% 
  ggplot(aes(x = as.factor(ID), y = enet_prob_v, group = as.factor(ID)))+
  geom_boxplot(aes(fill = schizophrenia), outlier.shape = NA)+
  geom_jitter(aes(colour = schizophrenia, alpha = 0.5))+
  geom_hline(
        yintercept = 0.33,
        linetype = 'dashed', 
        color = "red"
    )

df_pred_prob %>% 
  filter(ID == 24) %>% 
  ggplot(aes(x = schizophrenia, y = enet_prob_v))+
  geom_boxplot(aes(fill = schizophrenia, alpha = 0.5), outlier.shape = NA)+
  geom_jitter(alpha = 0.5)+
  geom_hline(
        yintercept = 0.33, 
        linetype = 'dashed', 
        color = "red"
    )

df_pred_prob %>% 
  filter(ID == 24) %>% 
  ggplot(aes(x = trial, y = enet_prob_v))+
  geom_point(aes(colour = schizophrenia))+
  geom_hline(
        yintercept = 0.33,
        linetype = 'dashed', 
        color = "red"
    )


# calibration plots
valProbggplot(enet_prob_v, y_test)
valProbggplot(enet_prob_se_v, y_test)

# Brier score on test set - 1se lam
enet_brier <- BrierScore(y_test, enet_prob_se_v)

# ROC + AUC on test set
enet_roc_obj <- roc(y_test, enet_prob_se_v)
enet_auc <- auc(enet_roc_obj)
plot(enet_roc_obj, print.auc = TRUE)
coords(enet_roc_obj, "best")   # 0.33 gives a good balance between sensitivity and specificity

# Plots 
plot(enet, xvar = "lambda")
abline(v = -log(bestlam), col = "darkgreen", lty = 2, lwd = 2)
abline(v = -log(selam), col = "orange", lty = 2, lwd = 2)
legend("topleft",
       legend = c("Lambda.min", "Lambda.1se"),
       col = c("darkgreen", "orange"),
       lty = 2,
       lwd = 2,
       bty = "n")
title("Coefficients - Elastic Net", line = 3)

# CV plot
plot(best_cv, pch = 19, main = "")   # remove default title so no overlap
title("Tuning Lambda - Elastic Net", line = 3)

# plot - best alpha 
plot(alpha_grid, cv_error, type = "b", pch = 19,
     xlab = expression(alpha),
     ylab = "Cross-validated binomial deviance",
     main = "Tuning Alpha - Elastic Net")

# highlight best alpha
points(best_alpha, min(cv_error), col = "darkgreen", pch = 19, cex = 1.5)
abline(v = best_alpha, col = "darkgreen", lty = 2)

# misclassification error
enet_class <- ifelse(enet_prob_se_v > 0.33, 1, 0)
# confusion matrix (predicted, outcome)
table(enet_class, y_test)
enet_error <- mean(enet_class != y_test)

enet_error # closer to zero is better. 0.5 = random guessing
enet_brier # closer to zero is better. >0.25 is poor
enet_auc # 1 is perfect discrimination. 0.5 = random guessing

# To see which predictor was shrunk to zero (selam)
rownames(coef(enet, s = selam))[as.vector(coef(enet, s = selam)) == 0]

coef_selam <- coef(enet, s = selam)

coef_df <- data.frame(predictor = rownames(coef_selam),
                      coefficient = as.vector(coef_selam))

non_zero_coeffs <- coef_df %>% filter(coefficient != 0)
non_zero_coeffs

zero_coeffs <- coef_df %>% filter(coefficient == 0)


```


```{r one-curve-split}

set.seed(2210)

mfpca_allcond <- mfpca_matrix %>%
  mutate(group = ifelse(group == "Case", 1, 0),
         gender = ifelse(gender == "Female", 1, 0)) %>%
  drop_na() # TEMPORARY

# for each iteration, holding out 1 EEG curve and training elastic net on remaining curves, predict held-out curve


test_curve <- sample(1: nrow(mfpca_allcond), size = 1)

sch_test <- mfpca_allcond[test_curve, ]
sch_train <- mfpca_allcond[-test_curve, ]

x <- model.matrix(group ~ . - ID, data = mfpca_allcond)[, -1]
y <- mfpca_allcond$group

x_train <- x[-test_curve,, drop = FALSE ]
x_test <- x[test_curve, ]

y_train <- y[-test_curve]
y_test <- y[test_curve]


alpha_grid <- seq(0.1, 0.9, by = 0.1)

cv_list <- lapply(alpha_grid, function(a) { cv.glmnet(x_train, y_train, alpha = a, family = "binomial")}) 
# extract CV error for eachg alpha 
cv_error <- sapply(cv_list, function(m) min(m$cvm))


best_index <- which.min(cv_error)
best_cv <- cv_list[[best_index]] 
best_alpha <- alpha_grid[best_index]


# best lambda corresponding to that alpha 
bestlam <- best_cv$lambda.min 
selam <- best_cv$lambda.1se


# fit elastic net model 
enet <- glmnet(x_train, y_train, alpha = best_alpha, family = "binomial") # standardize = TRUE is the default

# predicted probabilities with best lam from test set, type = response for pred prob (not log odds) 
enet_prob <- predict(enet, s = bestlam, newx = x_test, type = "response") 
# 1se lam 
enet_prob_se <- predict(enet, s = selam, newx = x_test, type = "response")


# convert to vector 
enet_prob_v <- as.vector(enet_prob) 
enet_prob_se_v <- as.vector(enet_prob_se) 
enet_brier <- BrierScore(y_test, enet_prob_se_v)


enet_prob_se_v
y_test
enet_brier

###########################
## testing another curve ##
###########################

set.seed(1022)

mfpca_allcond <- mfpca_matrix %>%
  mutate(group = ifelse(group == "Case", 1, 0),
         gender = ifelse(gender == "Female", 1, 0)) %>%
  drop_na() # TEMPORARY

# for each iteration, holding out 1 EEG curve and training elastic net on remaining curves, predict held-out curve


test_curve <- sample(1: nrow(mfpca_allcond), size = 1)

sch_test <- mfpca_allcond[test_curve, ]
sch_train <- mfpca_allcond[-test_curve, ]

x <- model.matrix(group ~ . - ID, data = mfpca_allcond)[, -1]
y <- mfpca_allcond$group

x_train <- x[-test_curve,, drop = FALSE ]
x_test <- x[test_curve, ]

y_train <- y[-test_curve]
y_test <- y[test_curve]


alpha_grid <- seq(0.1, 0.9, by = 0.1)

cv_list <- lapply(alpha_grid, function(a) { cv.glmnet(x_train, y_train, alpha = a, family = "binomial")}) 
# extract CV error for eachg alpha 
cv_error <- sapply(cv_list, function(m) min(m$cvm))


best_index <- which.min(cv_error)
best_cv <- cv_list[[best_index]] 
best_alpha <- alpha_grid[best_index]


# best lambda corresponding to that alpha 
bestlam <- best_cv$lambda.min 
selam <- best_cv$lambda.1se


# fit elastic net model 
enet <- glmnet(x_train, y_train, alpha = best_alpha, family = "binomial") # standardize = TRUE is the default

# predicted probabilities with best lam from test set, type = response for pred prob (not log odds) 
enet_prob <- predict(enet, s = bestlam, newx = x_test, type = "response") 
# 1se lam 
enet_prob_se <- predict(enet, s = selam, newx = x_test, type = "response")


# convert to vector 
enet_prob_v <- as.vector(enet_prob) 
enet_prob_se_v <- as.vector(enet_prob_se) 
enet_brier <- BrierScore(y_test, enet_prob_se_v)


enet_prob_se_v
y_test
enet_brier


```
Above, held out two single EEG curves from model fitting. Elastic net correctly classified the curve here - and a brier score close to the prediction. 


# Creating condition-specific matrix



```{r cond matrix}

mfpca_allcond <- mfpca_allcond %>% mutate(ID = as.numeric(as.character(ID)), 
         group = as.numeric(as.character(group)))  # should be 0/1 for glmnet

mfpca_test_all_cond <- mfpca_test_all_cond %>% mutate(ID = as.numeric(as.character(ID)), 
         group = as.numeric(as.character(group)))  # should be 0/1 for glmnet

# Condition 1 matrix
mfpca_cond1 <- mfpca_allcond %>%
  dplyr::select(ID, trial, group, starts_with("cond1"))

mfpca_cond1_test <- mfpca_test_all_cond %>% 
    dplyr::select(ID, trial, group, starts_with("cond1"))

# Condition 2 matrix
mfpca_cond2 <- mfpca_allcond %>%
  dplyr::select(ID, trial, group, starts_with("cond2"))

mfpca_cond2_test <- mfpca_test_all_cond %>% 
    dplyr::select(ID, trial, group, starts_with("cond2"))

# Condition 3 matrix
mfpca_cond3 <- mfpca_allcond %>%
  dplyr::select(ID, trial, group, starts_with("cond3"))

mfpca_cond3_test <- mfpca_test_all_cond %>% 
    dplyr::select(ID, trial, group, starts_with("cond3"))


```

```{r Evaluating cond 1 performance}
set.seed(0)

x_train <- model.matrix(group ~ .-ID, data = mfpca_cond1)[, -1] # gets rid of intercept term
y_train <- mfpca_cond1$group

x_test <- model.matrix(group ~ .-ID, data = mfpca_cond1_test)[, -1] # gets rid of intercept term
y_test <- mfpca_cond1_test$group



# elastic net: 0 < alpha < 1
# 0 ridge, 1 lasso
# cv to find best value of lambda and alpha 
alpha_grid <- seq(0.01, 0.99, by = 0.1)

cv_list <- lapply(alpha_grid, function(a) {
  cv.glmnet(x_train, y_train, 
            alpha = a, 
            family = "binomial")})

# extract CV error for eachg alpha 
cv_error <- sapply(cv_list, function(m) min(m$cvm))

# best model
best_index <- which.min(cv_error)
best_cv <- cv_list[[best_index]]
best_alpha <- alpha_grid[best_index]

# best lambda corresponding to that alpha
bestlam <- best_cv$lambda.min
selam <- best_cv$lambda.1se


# fit elastic net model
enet <- glmnet(x_train, y_train,
               alpha = best_alpha,
               family = "binomial")      # standardize = TRUE is the default 

# predicted probabilities with best lam from test set, type = response for pred prob (not log odds)
enet_prob <- predict(enet, s = bestlam, newx = x_test, type = "response")
# 1se lam
enet_prob_se <- predict(enet, s = selam, newx = x_test, type = "response")

# convert to vector
enet_prob_v <- as.vector(enet_prob)
enet_prob_se_v <- as.vector(enet_prob_se)

df_pred_c1 <- data.frame(enet_prob_se_v) %>% mutate(ID = mfpca_cond1_test$ID, .before = 1) %>% mutate(schizophrenia = as.factor(mfpca_cond1_test$group)) %>% mutate(trial = mfpca_cond1_test$trial)

df_pred_c1 %>% 
  ggplot(aes(x = as.factor(ID), y = enet_prob_se_v, group = as.factor(ID)))+
  geom_boxplot(aes(fill = schizophrenia), outlier.shape = NA)+
  geom_jitter(aes(colour = schizophrenia, alpha = 0.5))+
  geom_hline(
        yintercept = 0.33,
        linetype = 'dashed', 
        color = "red"
    )

# calibration plots
valProbggplot(enet_prob_v, y_test)
valProbggplot(enet_prob_se_v, y_test)

# Brier score on test set - 1se lam
enet_brier <- BrierScore(y_test, enet_prob_se_v)

# ROC + AUC on test set
enet_roc_obj <- roc(y_test, enet_prob_se_v)
enet_auc <- auc(enet_roc_obj)
plot(enet_roc_obj, print.auc = TRUE)
coords(enet_roc_obj, "best")   # 0.33 gives a good balance between sensitivity and specificity

# Plots 
plot(enet, xvar = "lambda")
abline(v = -log(bestlam), col = "darkgreen", lty = 2, lwd = 2)
abline(v = -log(selam), col = "orange", lty = 2, lwd = 2)
legend("topleft",
       legend = c("Lambda.min", "Lambda.1se"),
       col = c("darkgreen", "orange"),
       lty = 2,
       lwd = 2,
       bty = "n")
title("Coefficients - Elastic Net", line = 3)

# CV plot
plot(best_cv, pch = 19, main = "")   # remove default title so no overlap
title("Tuning Lambda - Elastic Net", line = 3)

# plot - best alpha 
plot(alpha_grid, cv_error, type = "b", pch = 19,
     xlab = expression(alpha),
     ylab = "Cross-validated binomial deviance",
     main = "Tuning Alpha - Elastic Net")


as.matrix(coef(enet, s = selam))[as.matrix(coef(enet, s = selam)) != 0, , drop = FALSE]


# To see which predictor was shrunk to zero (selam)
rownames(coef(enet, s = selam))[as.vector(coef(enet, s = selam)) == 0]

coef_selam <- coef(enet, s = selam)

coef_df <- data.frame(predictor = rownames(coef_selam),
                      coefficient = as.vector(coef_selam))

non_zero_coeffs_cond1 <- coef_df %>% filter(coefficient != 0)
non_zero_coeffs_cond1

zero_coeffs_cond1 <- coef_df %>% filter(coefficient == 0)
```

```{r Evaluating cond 2 performance}

x_train <- model.matrix(group ~ .-ID, data = mfpca_cond2)[, -1] 
y_train <- mfpca_cond2$group

x_test <- model.matrix(group ~ .-ID, data = mfpca_cond2_test)[, -1] 
y_test <- mfpca_cond2_test$group

# elastic net: 0 < alpha < 1
# 0 ridge, 1 lasso
# cv to find best value of lambda and alpha 
alpha_grid <- seq(0.1, 0.9, by = 0.1)

cv_list <- lapply(alpha_grid, function(a) {
  cv.glmnet(x_train, y_train, 
            alpha = a, 
            family = "binomial")})

# extract CV error for eachg alpha 
cv_error <- sapply(cv_list, function(m) min(m$cvm))

# best model
best_index <- which.min(cv_error)
best_cv <- cv_list[[best_index]]
best_alpha <- alpha_grid[best_index]

# best lambda corresponding to that alpha
bestlam <- best_cv$lambda.min
selam <- best_cv$lambda.1se


# fit elastic net model
enet <- glmnet(x_train, y_train,
               alpha = best_alpha,
               family = "binomial")      # standardize = TRUE is the default 

# predicted probabilities with best lam from test set, type = response for pred prob (not log odds)
enet_prob <- predict(enet, s = bestlam, newx = x_test, type = "response")
# 1se lam
enet_prob_se <- predict(enet, s = selam, newx = x_test, type = "response")

# convert to vector
enet_prob_v <- as.vector(enet_prob)
enet_prob_se_v <- as.vector(enet_prob_se)

df_pred_c2 <- data.frame(enet_prob_se_v) %>% mutate(ID = mfpca_cond2_test$ID, .before = 1) %>% mutate(schizophrenia = as.factor(mfpca_cond2_test$group)) %>% mutate(trial = mfpca_cond2_test$trial)

df_pred_c2 %>% 
  ggplot(aes(x = as.factor(ID), y = enet_prob_se_v, group = as.factor(ID)))+
  geom_boxplot(aes(fill = schizophrenia), outlier.shape = NA)+
  geom_jitter(aes(colour = schizophrenia, alpha = 0.5))+
  geom_hline(
        yintercept = 0.33,
        linetype = 'dashed', 
        color = "red"
    )

# calibration plots
valProbggplot(enet_prob_v, y_test)
valProbggplot(enet_prob_se_v, y_test)


# Brier score on test set - 1se lam
enet_brier <- BrierScore(y_test, enet_prob_se_v)

# ROC + AUC on test set
enet_roc_obj <- roc(y_test, enet_prob_se_v)
enet_auc <- auc(enet_roc_obj)
plot(enet_roc_obj, print.auc = TRUE)
coords(enet_roc_obj, "best")   # 0.33 gives a good balance between sensitivity and specificity

# Plots 
plot(enet, xvar = "lambda")
abline(v = -log(bestlam), col = "darkgreen", lty = 2, lwd = 2)
abline(v = -log(selam), col = "orange", lty = 2, lwd = 2)
legend("topleft",
       legend = c("Lambda.min", "Lambda.1se"),
       col = c("darkgreen", "orange"),
       lty = 2,
       lwd = 2,
       bty = "n")
title("Coefficients - Elastic Net", line = 3)

# CV plot
plot(best_cv, pch = 19, main = "")   # remove default title so no overlap
title("Tuning Lambda - Elastic Net", line = 3)

# plot - best alpha 
plot(alpha_grid, cv_error, type = "b", pch = 19,
     xlab = expression(alpha),
     ylab = "Cross-validated binomial deviance",
     main = "Tuning Alpha - Elastic Net")


as.matrix(coef(enet, s = selam))[as.matrix(coef(enet, s = selam)) != 0, , drop = FALSE]

# To see which predictor was shrunk to zero (selam)
rownames(coef(enet, s = selam))[as.vector(coef(enet, s = selam)) == 0]

coef_selam <- coef(enet, s = selam)

coef_df <- data.frame(predictor = rownames(coef_selam),
                      coefficient = as.vector(coef_selam))

non_zero_coeffs_cond2 <- coef_df %>% filter(coefficient != 0)
non_zero_coeffs_cond2

zero_coeffs_cond2 <- coef_df %>% filter(coefficient == 0)

```



```{r Evaluating cond 3 performance}

x_train <- model.matrix(group ~ .-ID, data = mfpca_cond3)[, -1] 
y_train <- mfpca_cond3$group

x_test <- model.matrix(group ~ .-ID, data = mfpca_cond3_test)[, -1] 
y_test <- mfpca_cond3_test$group


# elastic net: 0 < alpha < 1
# 0 ridge, 1 lasso
# cv to find best value of lambda and alpha 
alpha_grid <- seq(0.1, 0.9, by = 0.1)

cv_list <- lapply(alpha_grid, function(a) {
  cv.glmnet(x_train, y_train, 
            alpha = a, 
            family = "binomial")})

# extract CV error for eachg alpha 
cv_error <- sapply(cv_list, function(m) min(m$cvm))

# best model
best_index <- which.min(cv_error)
best_cv <- cv_list[[best_index]]
best_alpha <- alpha_grid[best_index]

# best lambda corresponding to that alpha
bestlam <- best_cv$lambda.min
selam <- best_cv$lambda.1se


# fit elastic net model
enet <- glmnet(x_train, y_train,
               alpha = best_alpha,
               family = "binomial")      # standardize = TRUE is the default 

# predicted probabilities with best lam from test set, type = response for pred prob (not log odds)
enet_prob <- predict(enet, s = bestlam, newx = x_test, type = "response")
# 1se lam
enet_prob_se <- predict(enet, s = selam, newx = x_test, type = "response")

# convert to vector
enet_prob_v <- as.vector(enet_prob)
enet_prob_se_v <- as.vector(enet_prob_se)

##plot predicted probabilities for each subject
df_pred_c3 <- data.frame(enet_prob_se_v) %>% mutate(ID = mfpca_cond3_test$ID, .before = 1) %>% mutate(schizophrenia = as.factor(mfpca_cond3_test$group)) %>% mutate(trial = mfpca_cond3_test$trial)

df_pred_c3 %>% 
  ggplot(aes(x = as.factor(ID), y = enet_prob_se_v, group = as.factor(ID)))+
  geom_boxplot(aes(fill = schizophrenia), outlier.shape = NA)+
  geom_jitter(aes(colour = schizophrenia, alpha = 0.5))+
  geom_hline(
        yintercept = 0.33,
        linetype = 'dashed', 
        color = "red"
    )

# calibration plots
valProbggplot(enet_prob_v, y_test)
valProbggplot(enet_prob_se_v, y_test)


# Brier score on test set - 1se lam
enet_brier <- BrierScore(y_test, enet_prob_se_v)

# ROC + AUC on test set
enet_roc_obj <- roc(y_test, enet_prob_se_v)
enet_auc <- auc(enet_roc_obj)
plot(enet_roc_obj, print.auc = TRUE)
coords(enet_roc_obj, "best")   # 0.33 gives a good balance between sensitivity and specificity

# Plots 
plot(enet, xvar = "lambda")
abline(v = -log(bestlam), col = "darkgreen", lty = 2, lwd = 2)
abline(v = -log(selam), col = "orange", lty = 2, lwd = 2)
legend("topleft",
       legend = c("Lambda.min", "Lambda.1se"),
       col = c("darkgreen", "orange"),
       lty = 2,
       lwd = 2,
       bty = "n")
title("Coefficients - Elastic Net", line = 3)

# CV plot
plot(best_cv, pch = 19, main = "")   # remove default title so no overlap
title("Tuning Lambda - Elastic Net", line = 3)

# plot - best alpha 
plot(alpha_grid, cv_error, type = "b", pch = 19,
     xlab = expression(alpha),
     ylab = "Cross-validated binomial deviance",
     main = "Tuning Alpha - Elastic Net")


as.matrix(coef(enet, s = selam))[as.matrix(coef(enet, s = selam)) != 0, , drop = FALSE]

# To see which predictor was shrunk to zero (selam)
rownames(coef(enet, s = selam))[as.vector(coef(enet, s = selam)) == 0]

coef_selam <- coef(enet, s = selam)

coef_df <- data.frame(predictor = rownames(coef_selam),
                      coefficient = as.vector(coef_selam))

non_zero_coeffs_cond3 <- coef_df %>% filter(coefficient != 0)
non_zero_coeffs_cond3

zero_coeffs_cond3 <- coef_df %>% filter(coefficient == 0)

```


# Cluster L1 for mFPCA 

Clustering L1 scores from mFPCA to compare with FPCA. 



```{r mFPCA level1 clustering}

##KMEANS CLUSTERING - LEVEL 1 mFPCA
set.seed(1)

level1_scores <- as.data.frame(mlfpc$score$level1)

level1_scores <- level1_scores %>%
  mutate(ID = unique(df_wide$ID), .before = 1)

rownames(level1_scores)

level1_scores_demo <- merge(level1_scores, demographic, by = "ID")

target_y <- level1_scores_demo$group
features_x <- level1_scores_demo[c(2:5)]
```

```{r, identify the optimal number of clusters}
# Decide how many clusters to look at
n_clusters <- 10

# Initialize total within sum of squares error: wss
wss <- numeric(n_clusters)

set.seed(1)

# Look over 1 to n possible clusters
for (i in 1:n_clusters) {
  # Fit the model: km.out
  km.out <- kmeans(features_x, centers = i, nstart = 20)
  # Save the within cluster sum of squares
  wss[i] <- km.out$tot.withinss
}

# Produce a scree plot
wss_df <- tibble(clusters = 1:n_clusters, wss = wss)
 
scree_plot <- ggplot(wss_df, aes(x = clusters, y = wss, group = 1)) +
    geom_point(size = 4)+
    geom_line() +
    scale_x_continuous(breaks = c(2, 4, 6, 8, 10)) +
    xlab('Number of clusters')
scree_plot

scree_plot +
    geom_hline(
        yintercept = wss, 
        linetype = 'dashed', 
        col = c(rep('#000000',4),'#FF0000', rep('#000000', 5))
    )
```

```{r}
set.seed(1)
kmeans_c1 <- kmeans(features_x, centers = 4, nstart = 10)

#CONFUSION MATRIX
table(level1_scores_demo$group, kmeans_c1$cluster)

#Visualise Clusters
fviz_cluster(kmeans_c1, data = features_x, 
             main = "K-Means Clusters - Level 1 mFPCA")
```

```{r plotting mfpca1 clusters}
##PLOTTING THE CLUSTERS
km <- data.frame(ID = level1_scores_demo$ID, 
                 cluster = kmeans_c1$cluster)

df_cluster <- merge(km, dfresponse_trial, by= "ID")

df_cluster %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = interaction(ID, trial), color = schizophrenia)) +
  geom_line()+
  facet_wrap(~ cluster) +
  labs(title = "EEG signals by Level 1 mFPCA cluster, coloured by schizophrenia") 

```

