---
title: "Condition Pipeline"
author: "Hollie Keating, Sarah Conway & Maeve Samali"
date: "2026-03-10"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document contains the pipeline which can be applied to any of the conditions and nodes. Included:

 - Smoothing 
 - Generate FPCA scores
 - Plot top 5 and bottom 5 of FPCA scores 1, 2, and 3 
 
```{r include = FALSE, warning=FALSE, message=FALSE}
library(data.table)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(plotly)
library(fda)
library(gridExtra)
library(fdapace)
library(tidyr)
library(table1)
library(factoextra)
library(cluster)
library(sjPlot)

#setwd("C:/Users/holli/OneDrive - National University of Ireland, Galway/Modules/Research Project")
setwd("C:/Users/35387/Documents/College/MSc Health Data Science/HDS Project/Rscript/Data")
```

```{r}
c = 1 # condition 1 
t = 10 # 10 trials 
t1 = 1 # selecting one trial
# r_len = 1638 # timepoints in condition c
# c_len = 32 # subjects 
node <- "fp1"

```

 
```{r}
# Loop which reads in separate excel file for each patient and compiles all data into a dataframe
# Merges time data and demogrpahic data which contains whether patients are controls or cases
missing = c(5, 8, 9, 26, 31, 35, 38)

#setwd("C:/Users/holli/OneDrive - Natramional University of Ireland, Galway/Modules/Research Project")
setwd("C:/Users/35387/Documents/College/MSc Health Data Science/HDS Project/Rscript/Data")

times = fread("archive/time.csv")
demographic <- fread("archive/demographic.csv")
  
filenames = list.files("archive/patients/")

p_list <- list()

for(i in 1:length(filenames)){
  
  pathtomyfile = paste0("archive/patients/", filenames[i])
  temp <- fread(pathtomyfile, sep = ",", header = "auto")
  temp <- temp[ ,1:12]
  colnames(temp)<- c("subject", "trial", "condition", "sample", "fp1", "af7", "af3", "f1", "f3", "f5", "f7", "ft7")
  temp <- merge(temp, times, by ="sample")
  temp <- merge(temp, demographic, by = "subject")
  temp<- temp%>%
    filter (time_ms > -100)%>%   # time from -100ms to remove unnecessary data while including some pre beep curve info (100ms is where the beep occurs
    # ifelse(subject %in% missing, trial==15, trial==1) %>% # Only one trial out of 100 choosen. Trial 1 so patient hasn't learnt the test/timing/ keep up concealment
    filter(trial<=t)
  temp<- rename(temp, schizophrenia = group)

  p_list[[i]] <- temp
  
}

p_df <- do.call(rbind, p_list)

p_df<- p_df%>%
  mutate(
    subject = as.factor(subject),
    schizophrenia = as.factor(schizophrenia)
  )
```

 
```{r}
## Smooth the data 
## Create a B-spline basis (mybasis) - building blocks to represent the curves
K = 100 # basis functions. More = more flexible
d = 3 # basis degree e.g. 3 = cubic

mybasis = create.bspline.basis(rangeval = c(0,1638), # x-axis range for curves 0-1638 timepoints
                               nbasis = K, 
                               norder = d+1) # cubic is an order 4 polynomial 


## Smoothing parameter object

## create an fd object - functional data object

penalty_order = 2 # 2nd derivative ie penalizes wigglier
penalty = 0.05 #amount of smoothing to be applied. Larger penalty = smoother.
fdParObj <- fdPar(fd(matrix(0, K, 1), mybasis), # fd object with k=100 coeff, 1 curve
                  Lfdobj = penalty_order, 
                  lambda = penalty)
```

 
```{r}
# Smooth the data 
cond <- p_df %>%
  filter(condition == c) %>%
  filter(trial == t1)

#the subjects are not ordered correctly to fix this
cond$subject <- as.numeric(as.character(cond$subject))
cond <- cond %>% arrange(subject)

r_len <- nrow(cond %>% filter(subject == 1))
c_len <- length(unique(cond$subject))

matrix_sm_all <- matrix(cond[[node]], nrow = r_len, ncol = c_len, byrow= FALSE)

## smooth the data - cond
# smooth.basis(x, y, fdParObj)
sm_cond <- smooth.basis(seq(1, r_len, length.out = r_len), # x
                            matrix_sm_all, # y
                            fdParObj)# $fd     # extracts the smoothed functional data (fd) object ie
plot(sm_cond,
     main = "Smoothed Curves")

```




```{r}
# Plots 
# dfresponse - raw eeg, smoothed eeg, merged demographic 
 
# sm_cond$y is raw eeg signals
# smoothed is predict(sm_cond) 
# fd is basis and coeff

dfresponse <- data.frame(ID = as.numeric(as.character(cond$subject)),
                         time = cond$time_ms,
                         eeg_raw = as.numeric(sm_cond$y),
                         eeg_smooth = as.numeric(predict(sm_cond))) %>%   # smoothed predictions
                          arrange(ID)


demographic <- demographic %>%
  rename(ID = subject) %>%
  mutate(
    group = recode(group, `0` = "Control", `1` = "Case"),
    gender = recode(gender, " M" = "Male", " F" = "Female"),
    group = factor(group, levels = c("Control", "Case")),
    gender = factor(gender, levels = c("Male", "Female")))

# table(demographic$group) # 25 controls, 11 cases
# table(demographic$gender) # 30 males. 6 females

myIDS = unique((dfresponse$ID))
#test = as.numeric(predict(sm_cond)) # smoothed predictions
#dfresponse$eeg_smooth = test

dfresponse_merged <- dfresponse %>% left_join(demographic, by = "ID")
```

```{r, fig.width=9, fig.asp=1.5}

# plot color = cases vs controls 
# group = ID: draw one seperate line for each patient ID
group_plot <- dfresponse_merged %>%
  ggplot(aes(x = time, y = eeg_smooth, group = ID, color = group)) +
  geom_line(alpha = 0.6) +
  scale_color_manual(values = c("Control" = "orange", "Case" = "darkgreen")) +
  theme_bw()+
  labs(title = "Smoothed Curves Case vs Controls",
       x = "Time",
       y = "EEG Signals",
       color = "")

#group_plot


gender_plot <- dfresponse_merged %>%
  ggplot(aes(x = time, y = eeg_smooth, group = ID, color = group)) +
  geom_line(alpha = 0.6) +
  scale_color_manual(values = c("Control" = "orange", "Case" = "darkgreen")) +
  theme_bw()+
  labs(title = "Smoothed Curves Case vs Controls",
       x = "Time",
       y = "EEG Signals",
       color = "")+facet_wrap(~gender)

#gender_plot


# plot by age > 40 or < 40     its not the age they werre diagnosed at its just the age they are now.................
age_plot <- dfresponse_merged %>% 
  mutate(age_grouped = ifelse(age < 40, "<40", ">=40")) %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = ID, color = group)) +
  geom_line(alpha = 0.6) +
  scale_color_manual(values = c("Control" = "orange", "Case" = "darkgreen")) +
  theme_bw()+
  labs(title = "Smoothed Curves by Age",
       x = "Time",
       y = "EEG Signals",
       color = "")+facet_wrap(~age_grouped)

#age_plot

# years in education 
# primary plus secondary is ~14 years 
edu_plot <- dfresponse_merged %>% 
  mutate(education_grouped = ifelse(education < 15, "<15", ">=15")) %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = ID, color = group)) +
  geom_line(alpha = 0.6) +
  scale_color_manual(values = c("Control" = "orange", "Case" = "darkgreen")) +
  theme_bw()+
  labs(title = "Smoothed Curves by Years in Education",
       x = "Time",
       y = "EEG Signals",
       color = "")+facet_wrap(~education_grouped)

#edu_plot

grid.arrange(group_plot, gender_plot, age_plot, edu_plot, ncol = 1)

# table_data <- dfresponse_merged %>%
#   distinct(ID, group, gender)      # one per patient
# 
# table1(~ gender | group,
#        data = table_data)


# demographic only has up to patient 77


dfresponse %>%
  filter(ID==1) %>%
  ggplot(aes(x=time, y = eeg_raw)) + geom_line() + geom_line(lwd = 1.5, aes(y=eeg_smooth, color = "smooth"))


```

 
 
## FPCA Scores
 
```{r}
## FPCA through conditional expectation (PACE) - irregular data allowed
# FPCA via PACE 
# applied to sm_cond: condition c smoothed data 

# make a list containing dataframes with rows grouped by ID
list_format <- split(dfresponse, dfresponse$ID)

# create lists of time and response (eeg)
list_t <- lapply(list_format,
                 dplyr::select, time)
list_eeg <- lapply(list_format,
                   dplyr::select, eeg_raw)

# unlist
list_t <- lapply(list_t,
                 unlist, use.names = F)
list_eeg <- lapply(list_eeg,
                   unlist, use.names = F)

# apply PACE (with fdapace)
resMEAN <- FPCA(Ly = list_eeg, Lt = list_t,
                optns = list(dataType = 'Sparse')) #Dense or p >> n (time pts > n patients)

##is the data dense?
plot(FPCA(Ly = list_eeg, Lt = list_t,
                optns = list(dataType = 'Dense')))
# FPC results
plot(resMEAN)
#resMEAN$xiEst #FPCA scores
```
 
 
# FPCA 1
 
```{r}
# FPCA scores

# FPC result
# resMEAN$xiEst #FPCA scores

fpca_scores <- as.data.frame(resMEAN$xiEst) %>% 
  mutate(ID =myIDS, .before = 1)             # add an ID row, first column

colnames(fpca_scores) <- c("ID", "fpca_1", "fpca_2", "fpca_3")


dfresponse <- dfresponse %>% mutate(ID = as.numeric(ID))


# Plots - merge FCP scores with df response 
fpca_scores_merged <- dfresponse_merged %>% left_join(fpca_scores, by = "ID")

```

```{r}
# Extract top 5 fpca 1 scores
# Top 5: 1638*5= 8190
top_5_fpca1 <- fpca_scores_merged %>%
  arrange(desc(fpca_1)) %>%
  slice(1:8190) %>%
  pull(ID)


# Extract bottom 5 fpca 1 scores
bottom_5_fpca1 <- fpca_scores_merged %>%
  arrange(fpca_1) %>%              # default ascending
  slice(1:8190) %>%
  pull(ID)

# df of top & bottom 
group_ids_fpca1 <- bind_rows(
  data.frame(ID = top_5_fpca1, group = "Top 5"),
  data.frame(ID = bottom_5_fpca1, group = "Bottom 5")
)


plot_data_fpca1 <- fpca_scores_merged %>%
  semi_join(group_ids_fpca1, by = "ID") %>%      # keeps rows that IDs are in both dfs
  left_join(group_ids_fpca1, by = "ID")          # adds the group label

plot_data_fpca1 %>% ggplot(aes(x = time, y = eeg_raw, group = ID, colour = group.x)) +
  geom_line(linewidth = 0.5) +
  theme_bw() +
  labs(title = "Top vs Bottom 5 Curves FPCA 1 Scores",
       colour = "Group")
  




```
 
# FPCA 2 

```{r}
# Extract top 5 fpca 1 scores
# Top 5: 1638*5= 8190
top_5_fpca2 <- fpca_scores_merged %>%
  arrange(desc(fpca_2)) %>%
  slice(1:8190) %>%
  pull(ID)


# Extract bottom 5 fpca 1 scores
bottom_5_fpca2 <- fpca_scores_merged %>%
  arrange(fpca_2) %>%              # default ascending
  slice(1:8190) %>%
  pull(ID)

# df of top & bottom 
group_ids_fpca2 <- bind_rows(
  data.frame(ID = top_5_fpca2, group = "Top 5"),
  data.frame(ID = bottom_5_fpca2, group = "Bottom 5")
)


plot_data_fpca2 <- fpca_scores_merged %>%
  semi_join(group_ids_fpca2, by = "ID") %>%      # keeps rows that IDs are in both dfs
  left_join(group_ids_fpca2, by = "ID")          # adds the group label

plot_data_fpca2 %>% ggplot(aes(x = time, y = eeg_raw, group = ID, colour = group.x)) +
  geom_line(lwd = 0.5) +
  theme_bw() +
  labs(title = "Top vs Bottom 5 Curves FPCA 2 Scores",
       colour = "Group")
  

```


# FPCA 3 

```{r}
# Extract top 5 fpca 1 scores
# Top 5: 1638*5= 8190
top_5_fpca3 <- fpca_scores_merged %>%
  arrange(desc(fpca_3)) %>%
  slice(1:8190) %>%
  pull(ID)


# Extract bottom 5 fpca 1 scores
bottom_5_fpca3 <- fpca_scores_merged %>%
  arrange(fpca_3) %>%              # default ascending
  slice(1:8190) %>%
  pull(ID)

# df of top & bottom 
group_ids_fpca3 <- bind_rows(
  data.frame(ID = top_5_fpca3, group = "Top 5"),
  data.frame(ID = bottom_5_fpca3, group = "Bottom 5")
)


plot_data_fpca3 <- fpca_scores_merged %>%
  semi_join(group_ids_fpca3, by = "ID") %>%      # keeps rows that IDs are in both dfs
  left_join(group_ids_fpca3, by = "ID")          # adds the group label

plot_data_fpca3 %>% ggplot(aes(x = time, y = eeg_raw, group = ID, colour = group.x)) +
  geom_line(linewidth = 0.5) +
  theme_bw() +
  labs(title = "Top vs Bottom 5 Curves FPCA 3 Scores",
       colour = "Group")

```

 
 
```{r}
##KMEANS CLUSTERING
set.seed(1)

#scores <- na.omit(scores)
rownames(fpca_scores)


fpca_scores_demo <- merge(fpca_scores, demographic, by = "ID")

target_y <- fpca_scores_demo$group

features_x <- fpca_scores_demo[c(2,3,4)]

kmeans_c1 <- kmeans(features_x, centers = 4, nstart = 10)

#CONFUSION MATRIX
table(fpca_scores_demo$group, kmeans_c1$cluster)

#mean(scores$group == kmeans_c1$cluster)
fviz_cluster(kmeans_c1, data = features_x, 
             main = "K-Means Clusters")
```

```{r}
##PLOTTING THE CLUSTERS
km <- data.frame(ID = fpca_scores_demo$ID, 
                 cluster = kmeans_c1$cluster)

#smoothed_all <- merge(smooth_long, cond1,  by = "subject")

df_cluster <- merge(km, fpca_scores_merged, by= "ID") #%>% 
  #arrange(subject, time_ms)

#plotting the cluster by schizophrenia
df_cluster %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = ID, colour = group)) +
  geom_line(lwd = 0.5)+
  facet_wrap(~ cluster) +
  labs(title = "EEG signmals by cluster, coloured by schizophrenia")
```


 
 
 
```{r}
subject_df <- fpca_scores_merged %>%
  distinct(ID, group, gender, age, education, fpca_1, fpca_2, fpca_3)

# subject_df$group <- factor(subject_df$group)
# subject_df$gender <- factor(subject_df$gender)


#=================
# Gender Model 
# ================

log_gender <- glm(group ~ gender + age + education + fpca_1 + fpca_2 + fpca_3,
             data = subject_df,
             family = binomial)

summary(log_gender)


## converting coefficients into OR

exp(coef(log_gender))


# from summary, FPCA1 shows most significance at p = 0.06.., genderM significance = 0.2131
# gender coef = 0.1485062




# ====================================
# Education Model
# ====================================

model_edu <- glm(group ~ education,
                 data = subject_df,
                 family = binomial)

summary(model_edu)
exp(coef(model_edu))

## for each additional year of education, the odds of being in the schizophrenia group decrease by about 43%

## creating a boxplot 

ggplot(subject_df, aes(x = factor(group), y = education)) +
  geom_boxplot() +
  labs(x = "Group", y = "Education")

# =====================
# FPCA Model 
# =====================

model_eeg <- glm(group ~ fpca_1 + fpca_2 + fpca_3,
                 data = subject_df,
                 family = binomial)

summary(model_eeg)

exp(coef(model_eeg))

# 1st PC most significant, OR = 0.99.. so only marginally reduces odds of being in schizophrenia group with increase of PC1

ggplot(subject_df, aes(x = factor(group), y = fpca_1)) +
  geom_boxplot() +
  labs(x = "Group", y = "FPCA 1")

```

### Education Model

Education was significantly associated with schizophrenia diagnosis, with higher education linked to lower log-odds (OR = 0.57, p = 0.018), indicating a decrease in odds of approximately 43% per additional year of education.

By analysing the grouped bar chart for education, individuals with schizophrenia tended to have lower levels of education compared to controls - however this association may reflect the impact of earlier illness onset or other cofounding factors.


### Gender Model

Gender was not significant within association with schizophrenia presence (OR = 0.15, p = 0.213). 
The distribution of gender was similar across the control and schizophrenia groups, with a higher proportion of males in both groups which supports our above result. However, the sample was predominantly male which may limit our ability to detect gender differences. 



### FPCA Model

FPCA 1 was significantly associated with schizophrenia diagnosis (p = 0.039), although the the odds ratio was quite small (OR \approx 1.00). FPCA 2 showed no association, and FPCA 3 showed a weak non-significant trend.  
 
# MFPCA Model 

So now we want to take into account that we have more than 1 function per person. People completed each sensory condition up to 100 times. Including more of these trials will improve our ability to predict whether an EEG is typical of schizophrenia or not.
 
## Smoothing all 10 trials

As before we need to smooth the data, but this time for all the trials.

```{r}
##smoothed data for 10 trials:
cond_trials <- p_df %>% 
  filter(condition == c) %>% 
  arrange(trial)

#the subjects are not ordered correctly to fix this
cond_trials$subject <- as.numeric(as.character(cond_trials$subject))
cond_trials <- cond_trials %>% arrange(subject)

##number of time points for each curve
r_lent <- nrow(cond_trials %>% filter(subject == 1) %>% filter(trial == 1))
#number of individual curves
c_lent <- nrow(cond_trials %>% distinct(subject, trial))

##all trials for all people for fp1
matrix_sm_trial <- matrix(cond_trials[[node]], nrow = r_lent, ncol = c_lent, byrow= FALSE)

## smooth the data - cond
# smooth.basis(x, y, fdParObj)
sm_trials <- smooth.basis(seq(1, r_lent, length.out = r_lent), # x
                            matrix_sm_trial, # y
                            fdParObj)# $fd     # extracts the smoothed functional data (fd) object ie
plot(sm_trials,
     main = "Smoothed Curves")
```

### Data in long format & plots

We create a dataframe of all the relevant measurements and we have plotted the smoothed data for each trial for each person below.

```{r}
##create a response df in long format
dfresponse10 <- data.frame(ID = as.numeric(as.character(cond_trials$subject)),
                         time = cond_trials$time_ms,
                         trial = cond_trials$trial,
                         eeg_raw = as.numeric(sm_trials$y),
                         eeg_smooth = as.numeric(predict(sm_trials)),
                         schizophrenia = cond_trials$schizophrenia) %>%   # smoothed predictions
                         arrange(ID)

##Plot all trials for each subject 
dfresponse10  %>% filter(ID <= 10) %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = factor(trial), color = factor(trial))) +
  geom_line() +
  facet_wrap(~ ID, nrow = 2) +
  ggtitle("Multilevel functional observations")

dfresponse10  %>% filter(ID > 10 & ID <=20) %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = factor(trial), color = factor(trial))) +
  geom_line() +
  facet_wrap(~ ID, nrow = 2) +
  ggtitle("Multilevel functional observations")

dfresponse10  %>% filter(ID > 20 & ID <= 71) %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = factor(trial), color = factor(trial))) +
  geom_line() +
  facet_wrap(~ ID, nrow = 2) +
  ggtitle("Multilevel functional observations")

dfresponse10  %>% filter(ID > 71) %>% 
  ggplot(aes(x = time, y = eeg_smooth, group = factor(trial), color = factor(trial))) +
  geom_line() +
  facet_wrap(~ ID, nrow = 2) +
  ggtitle("Multilevel functional observations")

```
## For multinode modelling

```{r}
#Reshape data for multi node glm 
nodes <- c("fp1", "af7", "af3", "f1", "f3", "f5", "f7", "ft7")

nodes_df <- p_df %>% pivot_longer(
  cols = all_of(nodes),
  names_to = "node",
  values_to = "EEG"
) %>% dplyr::select(subject, time_ms, schizophrenia, node, EEG) %>% 
  arrange(subject, node, time_ms)

```


## mFPCA

Now we perform multi-level modelling, where the first level is the individual subject level and the second level is the individual trial level. So we get a score for each person and for each curve (trial).

```{r}
# df: Data in long format with columns, subject Id = Id, stride_id (wave_id) = visit, arg = time (seq(0, 1, length = 101), obs = y values
#Output: each row is a curve with subject Id and wave id
df_wide <- pivot_wider(dfresponse10[,-4], names_from = time, values_from = eeg_smooth)
Y <- as.matrix(df_wide[, c(4:1641)]) # excluding column id, trial, schizophrenia

#mfpca
mlfpc <- mfpca.face(Y = Y, # matrix- each row is a curves
                    id = df_wide$ID, # subject id: c(1,1,2,2,2,3,3 ...)
                    visit = df_wide$trial, # trial id c(1,2,3,.., 1, 2,3, 4, ...)
                    twoway = TRUE,
                    #pve2 = 0.75, # pve for level 2
                    pve = 0.95
                    )

```

## Multlevel FPCs

-   Level 1

```{r}
t <- unique(dfresponse10$time)
egf1 <- data.frame(mlfpc$efunctions$level1)
ggplot(egf1) +
  geom_line(lwd =0.55, aes(x = t, y = egf1[, 1], color = " 1")) +
  geom_line(lwd =0.55,aes(x = t, y = egf1[, 2], color = " 2")) +
  geom_line(lwd =0.55,aes(x = t, y = egf1[, 3], color = " 3")) +
  geom_line(lwd =0.55,aes(x = t, y = egf1[, 4], color = " 4")) +
  scale_color_manual(values = c(" 1" = "blue", " 2" = "green", " 3" = "red", " 4" = "purple")) +
  labs(x = "Time", y = "value", color = "mFPC") +
  theme_bw()

```


-   Level 2

```{r}
t <- unique(dfresponse10$time)
egf2 <- data.frame(mlfpc$efunctions$level2)
ggplot(egf2) +
  geom_line(aes(x = t, y = egf2[, 1], color = " 1")) +
  geom_line(aes(x = t, y = egf2[, 2], color = " 2")) +
  geom_line(aes(x = t, y = egf2[, 3], color = " 3")) +
  geom_line(aes(x = t, y = egf2[, 4], color = " 4")) +
  scale_color_manual(values = c(" 1" = "blue", " 2" = "green", " 3" = "red", " 4" = "purple")) +
  labs(x = "Time (proportion)", y = "value", color = "mFPC") +
  theme_bw()

```


## Nature of the variability

Level 1

```{r}
# add and subtract eigenfunction to the overall mean
egf_l1 <- egf1 %>%  mutate(mu = mlfpc$mu,
                     p1 = mlfpc$mu  + X1*sqrt(mlfpc$evalues$level1[1]),
                     n1 = mlfpc$mu - X1*sqrt(mlfpc$evalues$level1[1]),
                     
                     p2 = mlfpc$mu  + X2*sqrt(mlfpc$evalues$level1[2]),
                     n2 = mlfpc$mu - X2*sqrt(mlfpc$evalues$level1[2]),
                     
                     p3 = mlfpc$mu  + X3*sqrt(mlfpc$evalues$level1[3]),
                     n3 = mlfpc$mu - X3*sqrt(mlfpc$evalues$level1[3]),
                     
                     p4 = mlfpc$mu  + X4*sqrt(mlfpc$evalues$level1[4]),
                     n4 = mlfpc$mu - X4*sqrt(mlfpc$evalues$level1[4]))

```



# First and second modes
```{r}
# first mode
ggplot(egf_l1, aes(x = t, y = mu)) + geom_line() +
geom_point(aes(x = t, y = p1), shape = "+", size = 3) + 
geom_point(aes(x = t, y = n1), shape = "-", size = 3)  +
  labs(y ="Y", x = "Time (proportion)") +ggtitle("PC 1 (..%)") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),plot.title.position = "panel"
  )

# Second mode
ggplot(egf_l1, aes(x = t, y = mu)) + geom_line() +
geom_point(aes(x = t, y = p2), shape = "+", size = 3) + 
geom_point(aes(x = t, y = n2), shape = "-", size = 3)  +
  labs(y ="Y", x = "Time (proportion)") + ggtitle("PC 2 (...%)") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),plot.title.position = "panel"
  )

# third mode
ggplot(egf_l1, aes(x = t, y = mu)) + geom_line() +
geom_point(aes(x = t, y = p3), shape = "+", size = 3) + 
geom_point(aes(x = t, y = n3), shape = "-", size = 3)  +
  labs(y ="Y", x = "Time (proportion)") + ggtitle("PC 3 (...%)") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),plot.title.position = "panel"
  )
```


 level 2

First and second modes

```{r}
egf2 <- egf2 %>%  mutate(mu = mlfpc$mu,
                       p1 = mlfpc$mu  + X1*sqrt(mlfpc$evalues$level2[1]),
                       n1 = mlfpc$mu - X1*sqrt(mlfpc$evalues$level2[1]),
                       
                       p2 = mlfpc$mu  + X2*sqrt(mlfpc$evalues$level2[2]),
                       n2 = mlfpc$mu - X2*sqrt(mlfpc$evalues$level2[2]),
                       
                       p3 = mlfpc$mu  + X3*sqrt(mlfpc$evalues$level2[3]),
                       n3 = mlfpc$mu - X3*sqrt(mlfpc$evalues$level2[3]),
                       
                       p4 = mlfpc$mu  + X4*sqrt(mlfpc$evalues$level2[4]),
                       n4 = mlfpc$mu - X4*sqrt(mlfpc$evalues$level2[4]))
```


```{r}
# First mode
ggplot(egf2, aes(x = t, y = mu)) + geom_line() +
geom_point(aes(x = t, y = p1), shape = "+", size = 3) + 
geom_point(aes(x = t, y = n1), shape = "-", size = 3)  +
  labs(y ="Y", x = "Time (proportion)") +
   ggtitle("PC 1 (....%)") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),plot.title.position = "panel"
  )

# Second mode
ggplot(egf2, aes(x = t, y = mu)) + geom_line() +
geom_point(aes(x = t, y = p2), shape = "+", size = 3) + 
geom_point(aes(x = t, y = n2), shape = "-", size = 3)  +
  labs(y ="Y", x = "Time (proportion)") +
   ggtitle("PC 2 (....%)") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),plot.title.position = "panel"
  )

# Third mode
ggplot(egf2, aes(x = t, y = mu)) + geom_line() +
geom_point(aes(x = t, y = p3), shape = "+", size = 3) + 
geom_point(aes(x = t, y = n3), shape = "-", size = 3)  +
  labs(y ="Y", x = "Time (proportion)") +
   ggtitle("PC 3 (....%)") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),plot.title.position = "panel"
  )
```


### Multilevel FPC scores

#### Level 1

Patient Level scores - 1 per person

```{r}
level1_scores <- cbind(data.frame(mlfpc$scores$level1), subj_id = as.factor(unique(df_wide$ID)) )# subject level scores
ggplot(level1_scores,aes(x = X1, y = X2, colour = subj_id)) + geom_point() +
  labs(x = "mFPC1", y = "mFPC2", title = "..") + theme_bw()

```

#### Level 2

Individual Trial scores (i.e. 1 score per function/trial) -multiple per person

```{r}
level2_scores <- cbind(data.frame(mlfpc$scores$level2), subj_id = factor(df_wide$ID), trial_id = df_wide$trial)

ggplot(level2_scores, aes(x = trial_id, y = X1, group = subj_id, colour = subj_id)) + geom_line() + theme_bw() +
  labs(y = "mFPC1", x = "trials order - longitudinal", title = "") 
 
ggplot(level2_scores, aes(x = subj_id, y = X1, group = subj_id)) + geom_boxplot() + theme_bw() +
  labs(y = "mFPC1", x = "Subject Id", title = "Score distribution") 
 

```






