---
title: Exploring the EEG Signals of Schizophrenia
authors:
  - name: Maeve Samali
    department: School of Mathematics, Applied Mathematics & Statistics
    affiliation: University of Galway
    location: Galway, Ireland
    email: M.samali1@universityofgalway.ie
  - name: Sarah Conway
    department: School of Mathematics, Applied Mathematics & Statistics
    affiliation: University of Galway
    location: Galway, Ireland
    email: S.Conway29@universityofgalway.ie
  - name: Hollie Keating 
    department: School of Mathematics, Applied Mathematics & Statistics
    affiliation: University of Galway
    location: Galway, Ireland
    email: H.Keating3@universityofgalway.ie
abstract: |
  Enter the text of your abstract here.
keywords:
  - Electroencephelogram
  - Schizophrenia
  - Elastic Net
  - Functional Data Analysis
bibliography: references.bib
biblio-style: unsrt
output: rticles::arxiv_article
header-includes:
  - \newcommand{\pandocbounded}[1]{#1}
---

# Introduction
\label{sec:intro}

Electroencephelogram (EEG) is a test used to measure electrical activity in the brain by attaching 64 sensor nodes to the scalp. Each sensor measures a single curve so each test produces 64 curves. 

- Functional data and how FDA can be used to analyse this

Schizophrenia is a mental health condition which is associated with unusual activity in the frontal and pre-frontal sections of the brain.

Sensory gating is the process by which the brain filters expected or repeated stimuli. This process is thought to be impaired in individuals with schizophrenia.


- Introduce EEG, how it relates to schizophrenia, how FDA can be used to analyse curves
- Current Literature - FDA for EEG signals or for schizophrenia has been used before? findings? etc.

## Data

Our data were taken from the National Institute of Mental Health (NIMH project number R01MH058262), available from [Kaggle](www.kaggle.com/datasets/broach/button-tone-sz/data)

In our data, 40 subjects underwent three sensory conditions (25 controls, 15 cases) in which:

1.  Subjects press a button to immediately produce a sound
2.  Subjects passively listen to the same sound without pressing a button
3.  Subjects press a button, and no sound is produced.


- Explain the study background trials, conditions etc.


# Methods
\label{sec:methods}

We used several methods: spline smoothing, permutation test, FPCA & mFPCA

## Smoothing

Due to the inherently noisy nature of EEG signals, smoothing was applied to the functional observations. When smoothing functions, it is important to balance the trade-off between variance and bias. Variance refers to the sensitivity of an estimated function to random fluctuations in the observed data, while bias refers to systematic deviation of the estimated function from the true underlying signal. High variance may lead to overfitting and poor generalisability to new data, whereas high bias may over smooth the data and obscure important signal characteristics.

Spline smoothing was used to reduce noise while preserving the underlying brain activity patterns. Specifically, 100 cubic B-spline basis functions were defined over the domain of each EEG signal, (0-1638ms). Using a relatively large number of basis functions provides sufficient flexibility to represent complex functional features while allowing smoothness to be controlled separately through penalisation.

A second-derivative roughness penalty was then applied. In penalised smoothing approaches, smoothness is controlled by incorporating a roughness penalty into the objective function used to fit the data, rather than by restricting the number of basis functions available (Ramsay and Silverman, 2005). Penalising the second derivative discourages rapid changes in curvature, allowing gradual variation in the fitted function while reducing excessive wiggliness. In practice, this encourages neighbouring B-spline coefficients to vary smoothly and can be viewed as approximating penalties on finite differences between adjacent coefficients. The roughness penalty pools information across neighbouring regions of the function, resulting in more stable estimates and reduced sensitivity to noise. Consequently, variance is reduced at the cost of a small increase in bias, yielding smoother and more interpretable functional representations of the EEG signals (Ramsay and Silverman, 2005; Gertheiss et al., 2024).


## Permutation test

Permutation t-tests were used to assess differences in functional EEG trajectories between cases and controls. Permutation-based inference is well suited to functional data and has been widely applied in the FDA literature (Bugni and Horowitz, 2021). The test evaluates the null hypothesis that the mean functions of the two groups are equal across the domain of the EEG signal.

Under the null hypothesis, group labels are exchangeable, meaning that randomly reallocating observations between groups should not systematically alter the value of the test statistic. The permutation procedure repeatedly shuffles the case-control labels and recalculates the test statistic for each permutation, generating an empirical null distribution. The observed test statistic is then compared with this distribution to determine statistical significance. The resulting p-value represents the probability of obtaining a test statistic at least as extreme as that observed if the null hypothesis were true.

In this study, five thousand permutation t-tests were applied to the FP1 electrode on a condition-by-condition basis. FP1 region was selected due to previous evidence linking frontal EEG abnormalities, particularly gamma-band dysfunction, to schizophrenia and associated cognitive deficits (Senkowski and Gallinat, 2015; Jeon and Polich, 2009; Shin et al., 2011).


## Functional Principal Component Analysis

Following smoothing, each EEG recording was represented as a continuous function observed over 1,638 time points. Although, as mentioned above, smoothing reduces noise and improves interpretability, functional observations remain high-dimensional. As discussed by J.O Ramsay et al., this data may display a small number of dominant modes of variation - to identify and explore these, principcal component analysis was applied to functional data. 

In functional PCA, eigenvalues of the bivariate variance-covariance function $\nu(s,t)$ denote the importance of each principal component. An eigenfunction is associated with each eigenvalue, and these describe major variational components. 

Due to the inherently high-dimensional nature of EEG data, directly analysing observations can be challenging. FPCA addresses this issue by representing each function as a linear combination of a small number of eigenfunctions, this allows the dominant sources of variation to be captured through a reduced set of principal component scores.  


*FPCA via PACE*

Functional Principal Component Analysis was implemented using the Principal Analysis by Conditional Estimation (PACE) method, developed by Yao et al. (2005). This pacakge is useful for the analysis of data that have been generated alongside noise or irregular time points. A key advantage of the PACE framework is that it can be used for sparsely or densely sampled random trajectories - the EEG recordings analysed in this study can be considered as densely sampled data, as each EEG curve was observed across a large number of time points on the same time grid. Here, PACE was used to estimate the principal modes of variation in the EEG curves and obtain subject-specific FPCA scores. 

Each participant contributed a single densely sampled EEG curve corresponding to Trial 1 at the FP1 electrode under Condition 1. 

### FPCA Formula

Each EEG curve is decomposed into: (1) a mean function; (2) eigenfunctions describing major modes of variation, and; (3) subject-specific FPCA scores.

The formula is represented as follows:

$$
X_i (t) = \mu (t) + \sum _{k=1}^K \xi_{ik} \Phi_k (t)
$$
where $\xi_{ik} = \int_0^1 (X_i(t) - \mu(t))\Phi_k(t)  dt$ are the FPC scores.

Intuitively, the first 3 PC scores summarise much of the variation present across the observed EEG curves, and are used to explain the original high-dimensional functional observations. 

*Exploring FPCA scores via extreme curve and then k-means  clustering*


## Logistic Regression Model



## Multilevel Functional Principal Component Analysis

*Motivation - standard FPCA assumes independent functional observations*
*Arranging multilevel functional data*

*Multilevel FPCA model - retaining components to 95% variability explained*

- why mfpca can be applied to our data - multiple trials

$$
X_ij = \mu(t) + 
$$

*Multilevel FPC scores - describe level 1 and 2*

## Elastic Net

# Results
\label{sec:results}

- discuss FPCA components look like (top & bottom 5 scores), eigenfunctions
- discuss mfpca components - what do they look like 
- elastic net how many predictors included in final model

# Discussion
\label{sec:discussion}

- what does the permutation test tell us?
- what di the fpcs tell us and what do the scores tell us about how much an individual eeg aligns with the eigenfunction?
- what do the mfpcs tell us? at each level?
- why do they differ at level 1? why is mfpca better/worse?
- what can we conclude from elastic net - more less shrinkage? too few subjects?
- How does our model perform - calibration curve, auc, misclassification
- Anything that could further this research

# Examples of citations, figures, tables, references
\label{sec:others}

You can insert references. Here is some text [@kour2014real; @kour2014fast] and see @hadash2018estimate.

The documentation for \verb+natbib+ may be found at

You can use custom blocks with LaTeX support from **rmarkdown** to create environment.

::: {.center latex=true}
  <http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}>
:::

Of note is the command \verb+\citet+, which produces citations
appropriate for use in inline text.  

You can insert LaTeX environment directly too.

\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}

produces

\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}


## Figures

You can insert figure using LaTeX directly. 

See Figure \ref{fig:fig1}. Here is how you add footnotes. [^Sample of the first footnote.]

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{4cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
  \label{fig:fig1}
\end{figure}

But you can also do that using R.

```{r fig2, fig.cap = "Another sample figure"}
plot(mtcars$mpg)
```

You can use **bookdown** to allow references for Tables and Figures.


## Tables

Below we can see how to use tables. 

See awesome Table~\ref{tab:table} which is written directly in LaTeX in source Rmd file.

\begin{table}
 \caption{Sample table title}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
  \label{tab:table}
\end{table}

You can also use R code for that.

```{r}
knitr::kable(head(mtcars), caption = "Head of mtcars table")
```


## Lists

- Item 1
- Item 2 
- Item 3

## References

Bugni, F.A. and Horowitz, J.L., 2021. Permutation tests for equality of distributions of functional data. Journal of Applied Econometrics, 36(7), pp.861–877. 

Gertheiss, J., Rügamer, D., Liew, B.X.W. and Greven, S., 2024. Functional data analysis: An introduction and recent developments. Biometrical Journal, 66(6), e202300363.

Jeon, Y.-W. and Polich, J., 2003. The P300 event-related potential in psychosis: a meta-analysis. American Journal of Psychiatry, 166(10), pp.1115–1123.

Ramsay, J.O. and Silverman, B.W., 2005. Functional Data Analysis. 2nd ed. New York: Springer. ISBN 0-387-40080-X.

<<<<<<< HEAD
Germán Aneiros-Pérez, Ricardo Cao, Juan M. Vilar-Fernández. (2011). Functional methods for time series prediction: a nonparametric approach. Journal of Forecasting. 30(4), pp.377-392. [Online]. Available at: https://doi.org/10.1002/for.1169 [Accessed 11 June 2026].

J.O Ramsay, Giles Hooker, Spencer Graves. (2009). Functional Data Analysis with R and MATLAB. NY: Springer New York.

Unknown. (2019). Principal Analysis by Conditional Estimation for Functional Data Analysis and Empirical Dynamics in Matlab.. [Online]. stat. Last Updated: 2019. Available at: https://anson.ucdavis.edu/~mueller/data/pace.html [Accessed 11 June 2026].

=======
Senkowski, D. and Gallinat, J., 2015. Dysfunctional prefrontal gamma-band oscillations reflect working memory and other cognitive deficits in schizophrenia. Biological Psychiatry, 77(12), pp.1010–1019.

Shin, Y.W., Kwon, J.S., Ha, T.H., Park, H.J., Hong, K.S., Kim, Y.Y. and Cho, H.S., 2011. Gamma oscillation in schizophrenia. Psychiatry Investigation, 8(4), pp.288–296.
>>>>>>> b720f887b39546aca4786026cf2d25f5312bc228
