---
title: Exploring the EEG Signals of Schizophrenia
authors:
  - name: Maeve Samali
    department: School of Mathematics, Applied Mathematics & Statistics
    affiliation: University of Galway
    location: Galway, Ireland
    email: M.samali1@universityofgalway.ie
  - name: Sarah Conway
    department: School of Mathematics, Applied Mathematics & Statistics
    affiliation: University of Galway
    location: Galway, Ireland
    email: S.Conway29@universityofgalway.ie
  - name: Hollie Keating 
    department: School of Mathematics, Applied Mathematics & Statistics
    affiliation: University of Galway
    location: Galway, Ireland
    email: H.Keating3@universityofgalway.ie
abstract: |
  Enter the text of your abstract here.
keywords:
  - Electroencephelogram
  - Schizophrenia
  - Elastic Net
  - Functional Data Analysis
bibliography: references.bib
biblio-style: unsrt
output: rticles::arxiv_article
header-includes:
  - \newcommand{\pandocbounded}[1]{#1}
---

# Introduction
\label{sec:intro}

Electroencephelogram (EEG) is a test used to measure electrical activity in the brain by attaching 64 sensor nodes to the scalp. Each sensor measures a single curve so each test produces 64 curves. 

- Functional data and how FDA can be used to analyse this

Schizophrenia is a mental health condition which is associated with unusual activity in the frontal and pre-frontal sections of the brain.

Sensory gating is the process by which the brain filters expected or repeated stimuli. This process is thought to be impaired in individuals with schizophrenia.


- Introduce EEG, how it relates to schizophrenia, how FDA can be used to analyse curves
- Current Literature - FDA for EEG signals or for schizophrenia has been used before? findings? etc.

## Data

Our data were taken from the National Institute of Mental Health (NIMH project number R01MH058262), available from [Kaggle](www.kaggle.com/datasets/broach/button-tone-sz/data)

In our data, 40 subjects underwent three sensory conditions (25 controls, 15 cases) in which:

1.  Subjects press a button to immediately produce a sound
2.  Subjects passively listen to the same sound without pressing a button
3.  Subjects press a button, and no sound is produced.


- Explain the study background trials, conditions etc.


# Methods
\label{sec:methods}

We used several methods: spline smoothing, permutation test, FPCA & mFPCA

## Smoothing

Due to the inherently noisy nature of EEG signals, smoothing was applied to the functional observations. When smoothing functions, it is important to balance the trade-off between variance and bias. Variance refers to the sensitivity of an estimated function to random fluctuations in the observed data, while bias refers to systematic deviation of the estimated function from the true underlying signal. High variance may lead to overfitting and poor generalisability to new data, whereas high bias may over smooth the data and obscure important signal characteristics.

Spline smoothing was used to reduce noise while preserving the underlying brain activity patterns. Specifically, 100 cubic B-spline basis functions were defined over the domain of each EEG signal, (0-1638ms). Using a relatively large number of basis functions provides sufficient flexibility to represent complex functional features while allowing smoothness to be controlled separately through penalisation.

A second-derivative roughness penalty was then applied. In penalised smoothing approaches, smoothness is controlled by incorporating a roughness penalty into the objective function used to fit the data, rather than by restricting the number of basis functions available (Ramsay and Silverman, 2005). Penalising the second derivative discourages rapid changes in curvature, allowing gradual variation in the fitted function while reducing excessive wiggliness. In practice, this encourages neighbouring B-spline coefficients to vary smoothly and can be viewed as approximating penalties on finite differences between adjacent coefficients. The roughness penalty pools information across neighbouring regions of the function, resulting in more stable estimates and reduced sensitivity to noise. Consequently, variance is reduced at the cost of a small increase in bias, yielding smoother and more interpretable functional representations of the EEG signals (Ramsay and Silverman, 2005; Gertheiss et al., 2024).

**insert plot of raw data and smoothed means
**insert formula 




## Permutation test

Permutation t-tests were used to assess differences in functional EEG trajectories between cases and controls. Permutation-based inference is well suited to functional data and has been widely applied in the FDA literature (Bugni and Horowitz, 2021). The test evaluates the null hypothesis that the mean functions of the two groups are equal across the domain of the EEG signal.

Under the null hypothesis, group labels are exchangeable, meaning that randomly reallocating observations between groups should not systematically alter the value of the test statistic. The permutation procedure repeatedly shuffles the case-control labels and recalculates the test statistic for each permutation, generating an empirical null distribution. The observed test statistic is then compared with this distribution to determine statistical significance. The resulting p-value represents the probability of obtaining a test statistic at least as extreme as that observed if the null hypothesis were true.

In this study, five thousand permutation t-tests were applied to the FP1 electrode on a condition-by-condition basis. FP1 region was selected due to previous evidence linking frontal EEG abnormalities, particularly gamma-band dysfunction, to schizophrenia and associated cognitive deficits (Senkowski and Gallinat, 2015; Jeon and Polich, 2009; Shin et al., 2011).


## Functional Principal Component Analysis

Following smoothing, each EEG recording was represented as a continuous function observed over 1,638 time points. Although, as mentioned above, smoothing reduces noise and improves interpretability, functional observations remain high-dimensional. As discussed by J.O Ramsay et al., this data may display a small number of dominant modes of variation - to identify and explore these, principcal component analysis was applied to functional data. 

In functional PCA, eigenvalues of the bivariate variance-covariance function $\nu(s,t)$ denote the importance of each principal component. An eigenfunction is associated with each eigenvalue, and these describe major variational components. 

Due to the inherently high-dimensional nature of EEG data, directly analysing observations can be challenging. FPCA addresses this issue by representing each function as a linear combination of a small number of eigenfunctions, this allows the dominant sources of variation to be captured through a reduced set of principal component scores.  


*FPCA via PACE*

Functional Principal Component Analysis was implemented using the Principal Analysis by Conditional Estimation (PACE) method, developed by Yao et al. (2005). This pacakge is useful for the analysis of data that have been generated alongside noise or irregular time points. A key advantage of the PACE framework is that it can be used for sparsely or densely sampled random trajectories - the EEG recordings analysed in this study can be considered as densely sampled data, as each EEG curve was observed across a large number of time points on the same time grid. Here, PACE was used to estimate the principal modes of variation in the EEG curves and obtain subject-specific FPCA scores (FPCs). 

Each participant contributed a single densely sampled EEG curve corresponding to Trial 1 at the FP1 electrode under Condition 1. 

### FPCA Formula

Each EEG curve is decomposed into: (1) a mean function; (2) eigenfunctions describing major modes of variation, and; (3) subject-specific FPCA scores.

The formula is represented as follows:

$$
X_i (t) = \mu (t) + \sum _{k=1}^K \xi_{ik} \Phi_k (t)
$$
where $\xi_{ik} = \int_0^1 (X_i(t) - \mu(t))\Phi_k(t)  dt$ are the FPC scores.

Intuitively, the first 3 PC scores summarise much of the variation present across the observed EEG curves, 95%, and are used to explain the original high-dimensional functional observations. 

*Exploring FPCA scores via extreme curve and then k-means  clustering*

To help interpret the functional principal components, five subjects with the highest and lowest scores for each of the first three FCPs were chosen for visualisation. The corresponding smoothed EEG curves were then plotted to examine how variation in each score related to the change in shape of the EEG signal. This aided visual understanding of the first three FPCs.  

Following this investigation, the first three FPCA scores were then used as inputs for the k-means clustering analysis. 

As quoted from Fortuna F. "cluster analysis aims to group a set of data such that units within clusters are more similar than across clusters with respect to a metric. In the FDA context, the most popular technique is the k-means clustering algorithm" (Fortuna F., Maturo F, 2019). 

To establish an appropriate number of clusters for our k-means analysis, several values of $k$ were explored and a four-cluster solution was selected. This number was chosen as it provided a reasonable balance between cluster separation and interpretability. Given the relatively small sample size, large numbers of clusters would have resulted in very small cluster memberships. The aim of this is to investigate whether subjects with similar FPCA score patterns group together naturally. The resulting clusters were compared with schizophrenia status using a confusion matrix, and smoothed EEG curves were then plotted within each cluster. This analysis was used to assess whether FPCA scores captured patterns in the EEG data that could be relevant for differentiating between schizophrenia cases and controls. 

## Logistic Regression Model

According to Lan N. Bui, "Logistic regression is commonly utilized in clinical and educational research to examine relationships between risk factors and binary outcomes" (Lan N. Bui, Qian Ding, 2025). Following clustering analysis, logistic regression models were fitted to examine whether FPCA scores were associated with the classification of schizophrenia (group = 1). 

The purpose of this analysis was to assess whetherthe selected three FPCA scores contained significant information to discriminate between schizophrenia cases and controls. Odds ratios were also calculated to aid interpretation of the resulting fitted coefficients. 

## Multilevel Functional Principal Component Analysis

At this stage in analysis, the EEG curves of subjects within one trial and one condition were investigated. To extend this, and include several of the repeated trials per subject and conditions one could no longer use standard FPCA. 

An important limitation of standard FPCA methods is that they are not designed for multilevel analyses, due to their assumption of independence between observations. Introduced in 2009 by Di CZ. et al., multilevel FPCA was designed to address the volume and importance of EEG data specifically. This methodology is designed to extract core intra- and inter- subject components of multilevel functional data. 

*Arranging multilevel functional data*

*Multilevel FPCA model - retaining components to 95% variability explained*

- why mfpca can be applied to our data - multiple trials

$$
X_ij = \mu(t) + 
$$

*Multilevel FPC scores - describe level 1 and 2*

## Elastic Net

Elastic net is a regularization and variable selection method that combines the penalties used in ridge and lasso regression. The method is controlled by a mixing parameter, $\alpha$, where $\alpha = 0$ corresponds to ridge regression and $\alpha = 1$ corresponds to lasso regression. Values of $\alpha$ between 0 and 1 provide a balance between the two approaches, (Zou and Hastie, 2005).

In ordinary least squares (OLS) regression, coefficients are estimated by minimizing the residual sum of squares (RSS), which represents the sum of the squared differences between the observed and predicted values. Consequently, the fitted regression line is chosen to produce predictions that are as close as possible to the observed data. Ridge regression, introduced by Hoerl and Kennard (1970), extends OLS by adding an (L_2) penalty to the objective function. This penalty constrains the magnitude of the regression coefficients by minimizing:

$$
\mathrm{RSS} + \lambda \sum_{j=1}^{p}\beta_j^2
$$

where $\lambda$ is a non-negative tuning parameter controlling the degree of shrinkage. By shrinking coefficients towards zero, ridge regression reduces model variance and can improve predictive performance in the presence of multicollinearity. However, coefficients are not reduced exactly to zero, meaning all predictors remain in the model. Lasso regression applies an (L_1) penalty to the regression coefficients and minimizes:

$$
\mathrm{RSS} + \lambda \sum_{j=1}^{p}|\beta_j|
$$

This penalty shrinks coefficients towards zero and can set some coefficients exactly equal to zero. Consequently, lasso performs automatic variable selection, removing less informative predictors and producing a more parsimonious model, (Tibshirani, 1996). Elastic net combines the (L_1) and (L_2) penalties and minimizes:

$$
\mathrm{RSS} + \lambda \left[(1-\alpha)\sum_{j=1}^{p}\beta_j^2 + \alpha\sum_{j=1}^{p}|\beta_j|\right]
$$

This approach balances the sparsity-inducing properties of lasso with the ability of ridge regression to handle correlated predictors. The mixing parameter $\alpha$ was tuned over a grid of values between 0 and 1. For each value of $\alpha$, an elastic net model was fitted and evaluated using 10-fold cross-validation. The value of $\alpha$ that produced the lowest cross-validated error was selected, together with its corresponding optimal penalty parameter, $\lambda$. Elastic net is particularly useful when the number of predictors exceeds the number of observations and when predictors are highly correlated. In such settings, it often provides more stable and accurate predictions than lasso regression alone (Zou and Hastie, 2005).

# Performance Evaluation

Model performance was evaluated using the Receiver Operating Characteristic (ROC) curve, the Area Under the ROC Curve (AUC), accuracy, misclassification error rate, and calibration. The ROC curve assesses the ability of a classification model to distinguish between two classes, (Fawcett, 2006), cases and controls in this application. It plots the true positive rate (sensitivity) against the false positive rate (specificity) across a range of classification thresholds. These thresholds determine the predicted probability above which an observation is classified as a case (1) rather than a control (0). The discriminatory ability of the model was quantified using the Area Under the ROC Curve (AUC), (Fawcett, 2006). The AUC can be interpreted as the probability that a randomly selected case receives a higher predicted probability than a randomly selected control. An AUC of 0.5 indicates no discriminatory ability, whereas an AUC of 1 indicates perfect discrimination. Classification performance was further evaluated using a confusion matrix. Accuracy was calculated as the proportion of correctly classified observations, while the misclassification error rate was calculated as the proportion of incorrectly classified observations. Calibration was assessed using calibration plots, which compare observed outcome frequencies with predicted probabilities, (Collins et al., 2014). A well-calibrated model produces predicted probabilities that closely reflect the observed outcomes. Perfect calibration is represented by a curve lying on the diagonal reference line, corresponding to a calibration slope of 1 and an intercept of 0.

# Results
\label{sec:results}

- discuss FPCA components look like (top & bottom 5 scores), eigenfunctions
- discuss mfpca components - what do they look like 
- elastic net how many predictors included in final model

# Discussion
\label{sec:discussion}

- what does the permutation test tell us?
- what di the fpcs tell us and what do the scores tell us about how much an individual eeg aligns with the eigenfunction?
- what do the mfpcs tell us? at each level?
- why do they differ at level 1? why is mfpca better/worse?
- what can we conclude from elastic net - more less shrinkage? too few subjects?
- How does our model perform - calibration curve, auc, misclassification
- Anything that could further this research

# Examples of citations, figures, tables, references
\label{sec:others}

You can insert references. Here is some text [@kour2014real; @kour2014fast] and see @hadash2018estimate.

The documentation for \verb+natbib+ may be found at

You can use custom blocks with LaTeX support from **rmarkdown** to create environment.

::: {.center latex=true}
  <http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}>
:::

Of note is the command \verb+\citet+, which produces citations
appropriate for use in inline text.  

You can insert LaTeX environment directly too.

\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}

produces

\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}


## Figures

You can insert figure using LaTeX directly. 

See Figure \ref{fig:fig1}. Here is how you add footnotes. [^Sample of the first footnote.]

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{4cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
  \label{fig:fig1}
\end{figure}

But you can also do that using R.

```{r fig2, fig.cap = "Another sample figure"}
plot(mtcars$mpg)
```

You can use **bookdown** to allow references for Tables and Figures.


## Tables

Below we can see how to use tables. 

See awesome Table~\ref{tab:table} which is written directly in LaTeX in source Rmd file.

\begin{table}
 \caption{Sample table title}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
  \label{tab:table}
\end{table}

You can also use R code for that.

```{r}
knitr::kable(head(mtcars), caption = "Head of mtcars table")
```


## Lists

- Item 1
- Item 2 
- Item 3

## References

Bugni, F.A. and Horowitz, J.L., 2021. Permutation tests for equality of distributions of functional data. Journal of Applied Econometrics, 36(7), pp.861–877. 

Collins, G.S., de Groot, J.A.H., Dutton, S., Omar, O., Shanyinde, M., Tajar, A., Voysey, M., Wharton, R., Yu, L.M., Moons, K.G.M. and Altman, D.G. (2014) ‘External validation of multivariable prediction models: a systematic review of methodological conduct and reporting’, BMC Medical Research Methodology, 14(1), p. 40.

Fawcett, T. (2006) ‘An introduction to ROC analysis’, Pattern Recognition Letters, 27(8), pp. 861–874.

Gertheiss, J., Rügamer, D., Liew, B.X.W. and Greven, S., 2024. Functional data analysis: An introduction and recent developments. Biometrical Journal, 66(6), e202300363.

Hoerl, A.E. and Kennard, R.W. (1970) ‘Ridge Regression: Biased Estimation for Nonorthogonal Problems’, Technometrics, 12(1), pp. 55–67.

Jeon, Y.-W. and Polich, J., 2003. The P300 event-related potential in psychosis: a meta-analysis. American Journal of Psychiatry, 166(10), pp.1115–1123.

Ramsay, J.O. and Silverman, B.W., 2005. Functional Data Analysis. 2nd ed. New York: Springer. ISBN 0-387-40080-X.

Robert Tibshirani (1996) ‘Regression shrinkage and selection via the lasso’, Journal of the Royal Statistical Society: Series B (Methodological), 58(1), pp. 267–288.

Germán Aneiros-Pérez, Ricardo Cao, Juan M. Vilar-Fernández. (2011). Functional methods for time series prediction: a nonparametric approach. Journal of Forecasting. 30(4), pp.377-392. [Online]. Available at: https://doi.org/10.1002/for.1169 [Accessed 11 June 2026].

J.O Ramsay, Giles Hooker, Spencer Graves. (2009). Functional Data Analysis with R and MATLAB. NY: Springer New York.

Unknown. (2019). Principal Analysis by Conditional Estimation for Functional Data Analysis and Empirical Dynamics in Matlab.. [Online]. stat. Last Updated: 2019. Available at: https://anson.ucdavis.edu/~mueller/data/pace.html [Accessed 11 June 2026].

=======
Senkowski, D. and Gallinat, J., 2015. Dysfunctional prefrontal gamma-band oscillations reflect working memory and other cognitive deficits in schizophrenia. Biological Psychiatry, 77(12), pp.1010–1019.

Shin, Y.W., Kwon, J.S., Ha, T.H., Park, H.J., Hong, K.S., Kim, Y.Y. and Cho, H.S., 2011. Gamma oscillation in schizophrenia. Psychiatry Investigation, 8(4), pp.288–296.
b720f887b39546aca4786026cf2d25f5312bc228

Fortuna, F., Maturo, F. K-means clustering of item characteristic curves and item information curves via functional principal component analysis. Qual Quant 53, 2291–2304 (2019). https://doi.org/10.1007/s11135-018-0724-7

Lan N. Bui, Qian Ding,
Logistic regression modeling: methodological insights and roadmap, Currents in Pharmacy Teaching and Learning, Volume 17, Issue 12, 2025, 102460, ISSN 1877-1297, https://doi.org/10.1016/j.cptl.2025.102460.

Di CZ, Crainiceanu CM, Caffo BS, Punjabi NM. MULTILEVEL FUNCTIONAL PRINCIPAL COMPONENT ANALYSIS. Ann Appl Stat. 2009 Mar 1;3(1):458-488. doi: 10.1214/08-AOAS206SUPP. PMID: 20221415; PMCID: PMC2835171.

Zou, H. and Hastie, T. (2005) ‘Regularization and variable selection via the elastic net’, Journal of the Royal Statistical Society: Series B (Statistical Methodology), 67(2), pp. 301–320.

