---
title: Exploring the EEG Signals of Schizophrenia
authors:
  - name: Maeve Samali
    department: School of Mathematics, Applied Mathematics & Statistics
    affiliation: University of Galway
    location: Galway, Ireland
    email: M.samali1@universityofgalway.ie
  - name: Sarah Conway
    department: School of Mathematics, Applied Mathematics & Statistics
    affiliation: University of Galway
    location: Galway, Ireland
    email: stariate@ee.mount-sheikh.edu
  - name: Hollie Keating 
    department: School of Mathematics, Applied Mathematics & Statistics
    affiliation: University of Galway
    location: Galway, Ireland
    email: stariate@ee.mount-sheikh.edu
abstract: |
  Enter the text of your abstract here.
keywords:
  - Electroencephelogram
  - Schizophrenia
  - Elastic Net
  - Functional Data Analysis
bibliography: references.bib
biblio-style: unsrt
output: rticles::arxiv_article
header-includes:
  - \newcommand{\pandocbounded}[1]{#1}
---

# Introduction
\label{sec:intro}

- Introduce EEG, how it relates to schizophrenia, how FDA can be used to analyse curves
- Current Literature - FDA for EEG signals or for schizophrenia has been used before? findings? etc.

##Data

- Explain the study background trials, conditions etc.


# Methods
\label{sec:methods}

We used several methods: spline smoothing, permutation test, FPCA & mFPCA

## Smoothing

- explain mathematically 
- Cubic b-spline, 2nd derivative penalty, why?

## Permutation test

## Functional Principal Component Analysis

- single trial, single node, single condition

## Multilevel Functional Principal Component Analysis

- why mfpca can be applied to our data - multiple trials

# Results
\label{sec:results}

- discuss FPCA components look like (top & bottom 5 scores), eigenfunctions
- discuss mfpca components - what do they look like 
- elastic net how many predictors included in final model

# Discussion
\label{sec:discussion}

- what does the permutation test tell us?
- what di the fpcs tell us and what do the scores tell us about how much an individual eeg aligns with the eigenfunction?
- what do the mfpcs tell us? at each level?
- why do they differ at level 1? why is mfpca better/worse?
- what can we conclude from elastic net - more less shrinkage? too few subjects?
- How does our model perform - calibration curve, auc, misclassification
- Anything that could further this research

# Examples of citations, figures, tables, references
\label{sec:others}

You can insert references. Here is some text [@kour2014real; @kour2014fast] and see @hadash2018estimate.

The documentation for \verb+natbib+ may be found at

You can use custom blocks with LaTeX support from **rmarkdown** to create environment.

::: {.center latex=true}
  <http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}>
:::

Of note is the command \verb+\citet+, which produces citations
appropriate for use in inline text.  

You can insert LaTeX environment directly too.

\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}

produces

\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}


## Figures

You can insert figure using LaTeX directly. 

See Figure \ref{fig:fig1}. Here is how you add footnotes. [^Sample of the first footnote.]

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{4cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
  \label{fig:fig1}
\end{figure}

But you can also do that using R.

```{r fig2, fig.cap = "Another sample figure"}
plot(mtcars$mpg)
```

You can use **bookdown** to allow references for Tables and Figures.


## Tables

Below we can see how to use tables. 

See awesome Table~\ref{tab:table} which is written directly in LaTeX in source Rmd file.

\begin{table}
 \caption{Sample table title}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
  \label{tab:table}
\end{table}

You can also use R code for that.

```{r}
knitr::kable(head(mtcars), caption = "Head of mtcars table")
```


## Lists

- Item 1
- Item 2 
- Item 3
