---
title: Exploring the EEG Signals of Schizophrenia
authors:
  - name: Maeve Samali
    department: School of Mathematics, Applied Mathematics & Statistics
    affiliation: University of Galway
    location: Galway, Ireland
    email: M.samali1@universityofgalway.ie
  - name: Sarah Conway
    department: School of Mathematics, Applied Mathematics & Statistics
    affiliation: University of Galway
    location: Galway, Ireland
    email: stariate@ee.mount-sheikh.edu
  - name: Hollie Keating 
    department: School of Mathematics, Applied Mathematics & Statistics
    affiliation: University of Galway
    location: Galway, Ireland
    email: stariate@ee.mount-sheikh.edu
abstract: |
  Enter the text of your abstract here.
keywords:
  - Electroencephelogram
  - Schizophrenia
  - Elastic Net
  - Functional Data Analysis
bibliography: references.bib
biblio-style: unsrt
output: rticles::arxiv_article
header-includes:
  - \newcommand{\pandocbounded}[1]{#1}
---

# Introduction
\label{sec:intro}

- Introduce EEG, how it relates to schizophrenia, how FDA can be used to analyse curves
- Current Literature - FDA for EEG signals or for schizophrenia has been used before? findings? etc.

##Data

- Explain the study background trials, conditions etc.


# Methods
\label{sec:methods}

We used several methods: spline smoothing, permutation test, FPCA & mFPCA

## Smoothing

Due to the inherently noisy nature of EEG signals, smoothing was applied to the functional observations. When smoothing functions, it is important to balance the trade-off between variance and bias. Variance refers to the sensitivity of an estimated function to random fluctuations in the observed data, while bias refers to systematic deviation of the estimated function from the true underlying signal. High variance may lead to overfitting and poor generalisability to new data, whereas high bias may over smooth the data and obscure important signal characteristics.

Spline smoothing was used to reduce noise while preserving the underlying brain activity patterns. Specifically, 100 cubic B-spline basis functions were defined over the domain of each EEG signal, (0-1638ms). Using a relatively large number of basis functions provides sufficient flexibility to represent complex functional features while allowing smoothness to be controlled separately through penalisation.

A second-derivative roughness penalty was then applied. In penalised smoothing approaches, smoothness is controlled by incorporating a roughness penalty into the objective function used to fit the data, rather than by restricting the number of basis functions available (Ramsay and Silverman, 2005). Penalising the second derivative discourages rapid changes in curvature, allowing gradual variation in the fitted function while reducing excessive wiggliness. In practice, this encourages neighbouring B-spline coefficients to vary smoothly and can be viewed as approximating penalties on finite differences between adjacent coefficients. The roughness penalty pools information across neighbouring regions of the function, resulting in more stable estimates and reduced sensitivity to noise. Consequently, variance is reduced at the cost of a small increase in bias, yielding smoother and more interpretable functional representations of the EEG signals (Ramsay and Silverman, 2005; Gertheiss et al., 2024).


## Permutation test

## Functional Principal Component Analysis

- single trial, single node, single condition

## Multilevel Functional Principal Component Analysis

- why mfpca can be applied to our data - multiple trials

# Results
\label{sec:results}

- discuss FPCA components look like (top & bottom 5 scores), eigenfunctions
- discuss mfpca components - what do they look like 
- elastic net how many predictors included in final model

# Discussion
\label{sec:discussion}

- what does the permutation test tell us?
- what di the fpcs tell us and what do the scores tell us about how much an individual eeg aligns with the eigenfunction?
- what do the mfpcs tell us? at each level?
- why do they differ at level 1? why is mfpca better/worse?
- what can we conclude from elastic net - more less shrinkage? too few subjects?
- How does our model perform - calibration curve, auc, misclassification
- Anything that could further this research

# Examples of citations, figures, tables, references
\label{sec:others}

You can insert references. Here is some text [@kour2014real; @kour2014fast] and see @hadash2018estimate.

The documentation for \verb+natbib+ may be found at

You can use custom blocks with LaTeX support from **rmarkdown** to create environment.

::: {.center latex=true}
  <http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}>
:::

Of note is the command \verb+\citet+, which produces citations
appropriate for use in inline text.  

You can insert LaTeX environment directly too.

\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}

produces

\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}


## Figures

You can insert figure using LaTeX directly. 

See Figure \ref{fig:fig1}. Here is how you add footnotes. [^Sample of the first footnote.]

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{4cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
  \label{fig:fig1}
\end{figure}

But you can also do that using R.

```{r fig2, fig.cap = "Another sample figure"}
plot(mtcars$mpg)
```

You can use **bookdown** to allow references for Tables and Figures.


## Tables

Below we can see how to use tables. 

See awesome Table~\ref{tab:table} which is written directly in LaTeX in source Rmd file.

\begin{table}
 \caption{Sample table title}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
  \label{tab:table}
\end{table}

You can also use R code for that.

```{r}
knitr::kable(head(mtcars), caption = "Head of mtcars table")
```


## Lists

- Item 1
- Item 2 
- Item 3
